### 2015-06-12 tex/manuscript.tex
Added a figure and started the manuscript

\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}

% use proper unicode fonts
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\usepackage{amsmath} % for better display of equations

\usepackage{setspace}
\usepackage{natbib}

\usepackage{titling} % controls the way the title information is displayed
\pretitle{\begin{flushleft}\Large}
\posttitle{\end{flushleft}}
\predate{}
\postdate{}
\preauthor{\begin{flushleft}}
\postauthor{\end{flushleft}}
\setlength{\droptitle}{-3em}

\usepackage{authblk} % adds some nice options for displaying the author list
\renewcommand\Authsep{\protect\\}
\renewcommand\Authands{\protect\\}

%% graphics packages
\usepackage{graphicx}
%\usepackage[nomarkers, tablesfirst]{endfloat} % for final
\usepackage{caption}
%\captionsetup{labelsep=none,textformat=empty} % for final
\captionsetup{labelformat=simple} % for drafts
\usepackage{booktabs}

%% ----------------------------------
%
%     Title and authorship information
%
%% ----------------------------------

\title{Colonization-extinction dynamics determine tree species range limits in northeastern North America}
\date{}
\author[1,4]{Matthew V. Talluto (mtalluto@gmail.com)}
\author[1]{Isabelle Boulangeat (isabelle.boulangeat@gmail.com)}
\author[1]{Steve Vissault (steve.vissault@uqar.ca)}
\author[2,3]{Wilfried Thuiller (wilfried.thuiller@ujf-grenoble.fr)}
\author[1]{Dominique Gravel (dominique\_gravel@uqar.ca)}
\affil[1]{Département de biologie, Université du Québec à Rimouski, Rimouski, Quebec, Canada}
\affil[2]{Université Grenoble Alpes, Laboratoire d’Ecologie Alpine (LECA), F-38000 Grenoble, France}
\affil[3]{CNRS, Laboratoire d’Ecologie Alpine (LECA), F-38000 Grenoble, France}
\affil[4]{Author for correspondance. Address: Departament de Biologie, chimie, et geographie, 300, Allée des Ursulines, Rimouski, Quebec G5L 3A1, Canada}

%% ----------------------------------
%
%     END PREAMBLE
%
%% ----------------------------------

\begin{document}
%\doublespacing
%% ----------------------------------
%
%     TITLE PAGE
%
%% ----------------------------------

%TC:ignore
\begin{titlingpage}
	\maketitle

	\begin{flushleft}

	\textbf{Short title:} Colonization-extinction dynamics of eastern North American trees

	\textbf{Keywords:}
	\end{flushleft}
\end{titlingpage}
%TC:endignore

%TC:break abstract
\begin{abstract}
\noindent

 \end{abstract}

%TC:break manuscript
%% ----------------------------------
%
%     INTRODUCTION
%
%% ----------------------------------

%\section*{Introduction}

%\section*{Methods}

%\section*{Results}

%\section*{Discussion}

\subsection*{Conclusions}
\begin{itemize}
	\item Range limits of [some/many/most/all] [temperate/boreal] tree species in northeastern north America are determined by colonization-extinction dynamics
	\item Moisture and energy (i.e., mean annual temperature and precipitation) are associated with colonization and extinction rates for many species
	\item For some species, the equilibrium state of the model did not match the predicted distribution of the species, suggesting that these species are either out of equilibrium with the environment, or that other processes (e.g., large-scale disturbances) contribute to their range dynamics.
	\item Resilience (represented by the dominant eigenvalue (c - e) near the transition point) varies widely among species, suggesting greatly differential rates of response to environmental change.
	\item Validation? Model Selection?
%		\begin{itemize}
%			\item use DIC with a few models for each species (for each combo of c and e)
%			\item $t^3 + p^3$
%			\item $t^2 + p^2$
%			\item $t^3$
%			\item $p^3$
%			\item $t^2$
%			\item $p^2$
%			\item intercept only
%		\end{itemize}
\end{itemize}

\subsection*{List of figures}
\begin{itemize}
	\item Fig 1: theoretical/introductory figure or diagram
	\item Fig 2: posterior response curves for c and e w.r.t climate
	\item Fig 3: maps for selected species (with remaining in appendix) - 4 panels - c; e; predicted range; SDM
	\item Fig 4: eigenvalue maps
\end{itemize}

\subsection*{Tables}
\input{table}

\end{document}

### 2015-06-12 tex/table.tex
Added a figure and started the manuscript

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
\begin{table}[h]
\label{tab:species_list}
\caption{Colonization-extinction model selected and model performance for each species studied}
\begin{tabular}{@{}llllll@{}}
\toprule
Scientific name             & Common name    & Model & $\Delta$DIC & AUC? & Occurrences \\ \midrule
{\it Abies balsamea}        & Balsam fir     & $T^3 + P^3$ &  & & 18961 \\
{\it Acer rubrum}           & Red maple      &  &  & & 36726 \\
{\it Acer saccharum}        & Sugar maple    &  &  & & 18634 \\
{\it Betula alleghaniensis} & Yellow birch   &  &  & & 10371 \\
{\it Betula papyrifera}     & Paper birch    &  &  & & 15531 \\
{\it Fagus grandifolia}     & American beech &  &  & & 9619 \\
{\it Fraxinus americana}    & White ash      &  &  & & 10385 \\
{\it Picea glauca}          & White spruce   &  &  & & 8526 \\
{\it Pinus banksiana}       & Jack pine      &  &  & & 3023 \\
{\it Populus tremuloides}   & Quaking aspen  &  &  & & 12736 \\ 
{\it Quercus alba}          & White oak      &  &  & & 17280 \\
{\it Ulmus americana}       & American elm   &  &  & & 11597 \\ \midrule
\textbf{Species not started} & & & & & \\
{\it Pinus taeda} & & & & & 16351 \\
{\it Liquidambar styraciflua} & & & & & 14778 \\
{\it Quercus rubra} & & & & & 14337 \\
{\it Liriodendron tulipifera} & & & & & 11045 \\
{\it Quercus velutina} & & & & & 9997 \\
{\it Nyssa sylvatica} & & & & & 9209 \\
{\it Quercus nigra} & & & & & 8291 \\
{\it Pinus strobus} & & & & & 8182 \\
{\it Fraxinus pennsylvanica} & & & & & 7662 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[h]
\label{tab:species_list}
\caption{Parameterized equations for colonization and extinction rate}
\begin{tabular}{lll}
\toprule
Species    & Colonization & Extinction  \\ \midrule
{\it Abies balsamea} & & \\ 
\bottomrule
\end{tabular}
\end{table}

### 2015-06-12 scr/7_make_figs.r

@@ -5,12 +5,22 @@ library(rgdal)
 library(fields)
 library(reshape2)
 library(raster)
+library(coda)
 
 # handle command line arguments
 parser = ArgumentParser()
 parser$add_argument("-s", "--species", default="28731-ACE-SAC", help="desired species code")
 argList = parser$parse_args()
 spName = argList$species
+
+
+# plot traces
+posterior = readRDS(paste("results/", spName, "/", spName, "_posterior.rds", sep=""))
+pdf(width=30, height = 10, file=paste("img/", spName, "/", spName, "_mcmc_trace.pdf", sep=""))
+par(mfcol=c(2,7), mar=c(2,2,2,1), oma=c(0,0,0,1))
+plot(posterior, ask=FALSE, auto.layout=FALSE)
+dev.off()
+
 
 # set up projections
 P4S.latlon <- CRS("+proj=longlat +datum=WGS84")

### 2015-06-12 scr/4_fit_stm.r
Added most of the code for SA model selection

@@ -2,42 +2,48 @@
 library(argparse)
 parser = ArgumentParser()
 parser$add_argument("-s", "--species", default="28731-ACE-SAC", help="desired species code")
-parser$add_argument("-c", "--cubic", action="store_true", default=FALSE, help="use cubic terms in model (default is square only)")
 parser$add_argument("-r", "--rf", action="store_true", default=FALSE, help="random forest for prevalence (default: GLM)")
 parser$add_argument("-m", "--gam", action="store_true", default=FALSE, help="use GAM for prevalence (default: GLM)")
-parser$add_argument("-f", "--fraction", default=0.1, type="double", help="proportion of data to use for model fitting")
+parser$add_argument("-f", "--fraction", default=0.5, type="double", help="proportion of data to use for model fitting")
 parser$add_argument("-i", "--interval", default=5, type="double", help="how many years should the parameterization interval be")
+parser$add_argument("-t", "--tempvar", default="annual_mean_temp", help="Temperature variable to use")
+parser$add_argument("-p", "--precipvar", default="tot_annual_pp", help="Precipitation variable to use")
+parser$add_argument("-c", "--coldesign", default="1111111", help="7-character design vector for colonization model")
+parser$add_argument("-e", "--extdesign", default="1111111", help="7-character design vector for extinction model")
+
+
 argList = parser$parse_args()
 spName = argList$species
 targetInterval = argList$interval
+colDesign = sapply(1:nchar(argList$coldesign), function(i) as.integer(substr(argList$coldesign,i,i)))
+extDesign = sapply(1:nchar(argList$extdesign), function(i) as.integer(substr(argList$extdesign,i,i)))
 
-minus_log_likelihood = function(p, dat, useCubic)
+design_err = function(x)
 {
-	inv_logit = function(logitVal)
-	{
-		val = numeric(length(logitVal))
-		val[logitVal > 0] = 1.0 / (1.0 + exp(-logitVal[logitVal > 0]))
-		val[logitVal <= 0] = exp(logitVal[logitVal <= 0]) / (1.0 + exp(logitVal[logitVal <= 0]))
-		
[…]

### 2015-06-19 tex/manuscript.tex
Commit everything before I start cleaning up for ms production

@@
 \usepackage{booktabs}
+\usepackage[flushleft]{threeparttable}
 
@@
 \date{}
-\author[1,4]{Matthew V. Talluto (mtalluto@gmail.com)}
+\author[1,2,3,4]{Matthew V. Talluto (mtalluto@gmail.com)}
 \author[1]{Isabelle Boulangeat (isabelle.boulangeat@gmail.com)}
@@
 
-%\section*{Methods}
+\section*{Methods}
+
+\subsection*{Model description}
+
+\subsection*{Prevalence model}
+We used the RF package in R to build models of regional prevalence for each species. 
+For predictors, we initially evaluated a set of xx environmental variables. 
+From these, we selected xx variables that were relatively uncorrelated to each other (pairwise Pearson correlation < xx) and that had high loadings in a principal components analysis (PCA) of all variables.
+In this way, we reduced the risk of estimation problems due to correlations among the predictors while retaining variables that were representative of the available climatic space.
+
+\subsection*{Model selection}
+For each species, we evaluated a set of 145 colonization-extinction models. 
+Each model consisted of at most two environmental variables, one associated with temperature and the other precipitation\. 
+The model set was constructed from eight possible model forms (Table?) applied to all possible two-variable combinations of the four energy and four moisture variables (i.e., 16 possible variable combinations), along with two additional single-variable forms for each variable and an intercept-only model.
+Because of the large number of models, fitting all models using MCMC was infeasible; therefore, we performed model selection by computing the Akaike Information Criterion (AIC) on maximum likelihood estimates for each model, fit using simulated annealing (package GenSA in R).
+We then selected the model that minimized AIC for final parameterization.
+
+\subsection*{Parameterization}
+We used a Markov Chain Monte Carlo (MCMC) sampler employing a Metropolis-Hastings scheme to estimate posterior distributions for all parameters.
+This method has a number of advantages over the simulated annealing approach used for preliminary model selection.
+First, MCMC results in full joint posterior distributions of the parameters, allowing for a robust understanding of the uncertainty both in model parameters and predictions.
+Second, the use of a Bayesian framework allows the inclusion of prior information.
+As our parameters were all estimated on the logit scale, we used the weakly informative Cauchy priors suggested by (cite Gelman paper), with a mean of 0 and (variance?) of 10 for intercept parameters or 2.5 for slope parameters.
+Finally, having posterior distributions of the predictors allows for the information produced by our model to be incorporated in future meta-modeling efforts that can further improve our characterization of species' ranges and our understanding of model uncertainty (cite GEB paper).
+
+For each species, we obtained xxx posterior samples...
+We assessed convergence...
 
@@
 	\item Validation? Model Selection?
-%		\begin{itemize}
-%			\item use DIC with a few models for each species (for each combo of c and e)
-%			\item $t^3 + p^3$
-%			\item $t^2 + p^2$
-%			\item $t^3$
-%			\item $p^3$
-%			\item $t^2$
-%			\item $p^2$
-%			\item intercept only
-%		\end{itemize}
 \end{itemize}

### 2015-06-19 tex/table.tex
Commit everything before I start cleaning up for ms production

@@
 % \usepackage{booktabs}
-\begin{table}[h]
+\begin{table}[tb]
 \label{tab:species_list}
-\caption{Colonization-extinction model selected and model performance for each species studied}
-\begin{tabular}{@{}llllll@{}}
+\caption{List of species studied and the number of observations in the database}
+\begin{tabular}{llll}
 \toprule
-Scientific name             & Common name    & Model & $\Delta$DIC & AUC? & Occurrences \\ \midrule
-{\it Abies balsamea}        & Balsam fir     & $T^3 + P^3$ &  & & 18961 \\
-{\it Acer rubrum}           & Red maple      &  &  & & 36726 \\
-{\it Acer saccharum}        & Sugar maple    &  &  & & 18634 \\
-{\it Betula alleghaniensis} & Yellow birch   &  &  & & 10371 \\
-{\it Betula papyrifera}     & Paper birch    &  &  & & 15531 \\
-{\it Fagus grandifolia}     & American beech &  &  & & 9619 \\
-{\it Fraxinus americana}    & White ash      &  &  & & 10385 \\
-{\it Picea glauca}          & White spruce   &  &  & & 8526 \\
-{\it Pinus banksiana}       & Jack pine      &  &  & & 3023 \\
-{\it Populus tremuloides}   & Quaking aspen  &  &  & & 12736 \\ 
-{\it Quercus alba}          & White oak      &  &  & & 17280 \\
-{\it Ulmus americana}       & American elm   &  &  & & 11597 \\ \midrule
-\textbf{Species not started} & & & & & \\
-{\it Pinus taeda} & & & & & 16351 \\
-{\it Liquidambar styraciflua} & & & & & 14778 \\
-{\it Quercus rubra} & & & & & 14337 \\
-{\it Liriodendron tulipifera} & & & & & 11045 \\
-{\it Quercus velutina} & & & & & 9997 \\
-{\it Nyssa sylvatica} & & & & & 9209 \\
-{\it Quercus nigra} & & & & & 8291 \\
-{\it Pinus strobus} & & & & & 8182 \\
-{\it Fraxinus pennsylvanica} & & & & & 7662 \\
+Scientific name               & English name   & Occurrences & Transitions \\ 
+\midrule
+{\it Abies balsamea}          & Balsam fir     & 18961       &             \\
+{\it Acer rubrum}             & Red maple      & 36726       &             \\
+{\it Acer saccharum}          & Sugar maple    & 18634       &             \\
+{\it Betula alleghaniensis}   & Yellow birch   & 10371       &             \\
+{\it Betula papyrifera}       & Paper birch    & 15531       &             \\
+{\it Fagus grandifolia}       & American beech & 9619        &             \\
+{\it Fraxinus americana}      & White ash      & 10385       &             \\
+{\it Fraxinus pennsylvanica}  &                & 7662        &             \\
+{\it Liquidambar styraciflua} &                & 14778       &             \\
+{\it Liriodendron tulipifera} &                & 11045       &             \\
+{\it Nyssa sylvatica}         &                & 9209        &             \\
+{\it Picea glauca}            & White spruce   & 8526        &             \\
+{\it Pinus banksiana}         & Jack pine      & 3023        &             \\
+{\it Pinus strobus}           &                & 8182        &             \\
+{\it Pinus taeda}             &                & 16351       &             \\
+{\it Populus tremuloides}     & Quaking aspen  & 12736       &             \\ 
+{\it Quercus alba}            & White oak      & 17280       &             \\
+{\it Quercus nigra}           &                & 8291        &             \\
+{\it Quercus rubra}           &                & 14337       &             \\
+{\it Quercus velutina}        &                & 9997        &             \\
+{\it Ulmus americana}         & American elm   & 11597       &             \\ 
 \bottomrule
@@
 
-\begin{table}[h]
-\label{tab:species_list}
+\begin{table}[tb]
+\begin{threeparttable}
+\label{tab:model_selection}
+\caption{Colonization-extinction model selected for each species studied}
+\begin{tabular}{llllll}
+\toprule
+Scientific name               & Temperature Variable & Precipitation Variable & Model\tnote{1} & $\Delta$BIC\tnote{2} & BIC$_w$\tnote{3} \\ 
+\midrule
+{\it Abies balsamea}          & env1                 & env2                   & $T^3 + P^3$    & xxx                  &                  \\
+{\it Acer rubrum}             &   &   & & \\
+{\it Acer saccharum}          &   &   & & \\
+{\it Betula alleghaniensis}   &   &   & & \\
+{\it Betula papyrifera}       &   &   & & \\
+{\it Fagus grandifolia}       &   &   & & \\
+{\it Fraxinus americana}      &   &   & & \\
+{\it Fraxinus pennsylvanica}  &   &   & & \\
+{\it Liquidambar styraciflua} &   &   & & \\
+{\it Liriodendron tulipifera} &   &   & & \\
+{\it Nyssa sylvatica}         &   &   & & \\
+{\it Picea glauca}            &   &   & & \\
+{\it Pinus banksiana}         &   &   & & \\
+{\it Pinus strobus}           &   &   & & \\
+{\it Pinus taeda}             &   &   & & \\
+{\it Populus tremuloides}     &   &   & & \\ 
+{\it Quercus alba}            &   &   & & \\
+{\it Quercus nigra}           &   &   & & \\
+{\it Quercus rubra}           &   &   & & \\
+{\it Quercus velutina}        &   &   & & \\
+{\it Ulmus americana}         &   &   & & \\ 
+\bottomrule
+\end{tabular}
+\begin{tablenotes}
+\item [1] Model specification shows whether temperature (T) and precipitation (P) variables were included as well as the order of polynomial used.
+\item [2] $\Delta$BIC refers to absolute BIC difference between the selected model and the next best model in the set.
+\item [3] BIC$_w$ can be interpreted as the probability that this model is the best model in the set (cite Burnham and Anderson).
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+
+\begin{table}[tb]
+\label{tab:parameterized_equations}
 \caption{Parameterized equations for colonization and extinction rate}

### 2015-06-19 scr/1_prep_data.r

@@ -38,7 +38,8 @@ scalefunc = function(sc_mean, sc_sd) {
 		means = sc_mean,
 		sds = sc_sd))
 }
-clim.scale = scalefunc(attr(stateData.clim.scaled, "scaled:center"), attr(stateData.clim.scaled, "scaled:scale"))
+clim.scale = scalefunc(attr(stateData.clim.scaled, "scaled:center"), 
+attr(stateData.clim.scaled, "scaled:scale"))
 
 stateData.scaled = cbind(stateData[,-climVars], stateData.clim.scaled)
 transitionData.scaled = clim.scale$scale(transitionData)

### 2015-06-19 scr/3_project_sdm.r

@@ -5,6 +5,7 @@ library(gam)
 library(randomForest)
 parser = ArgumentParser()
 parser$add_argument("-s", "--species", default="28731-ACE-SAC", help="desired species code")
+parser$add_argument("-f", "--fraction", default=0.5, type="double", help="proportion of data to use for model fitting")
 argList = parser$parse_args()
 spName = argList$species
 
@@ -20,6 +21,19 @@ transitionData.projected$expectedGLM = predict(glm.mod, newdata = transitionData
 transitionData.projected$expectedGAM = predict(gam.mod, newdata = transitionData.scaled, type='response')
 transitionData.projected$expectedRF = predict(rf.mod, newdata = transitionData.scaled, type='prob')[,2]
 saveRDS(transitionData.projected, paste("dat/", spName, "/", spName, "_transitions_projected.rds", sep=""))
+
+
+# subset the transition data for use in the annealing
+if(argList$fraction < 1)
+{
+	sel = sample(nrow(transitionData.projected), 
+			as.integer(argList$fraction*nrow(transitionData.projected)))
+} else {
+	sel = 1:nrow(transitionData.projected)
+}
+transitionData.subsample = transitionData.projected[sel,]
+saveRDS(transitionData.subsample, paste("dat/", spName, "/", spName, "_transitions_projected_subsample.rds", sep=""))
+
 
 
 # project the SDMs to the climate grid

### 2015-06-19 scr/4_fit_stm.r

@@ -1,26 +1,20 @@
 #!/usr/bin/Rscript
-library(argparse)
-parser = ArgumentParser()
-parser$add_argument("-s", "--species", default="28731-ACE-SAC", help="desired species code")
-parser$add_argument("-r", "--rf", action="store_true", default=FALSE, help="random forest for prevalence (default: GLM)")
-parser$add_argument("-m", "--gam", action="store_true", default=FALSE, help="use GAM for prevalence (default: GLM)")
-parser$add_argument("-f", "--fraction", default=0.5, type="double", help="proportion of data to use for model fitting")
-parser$add_argument("-i", "--interval", default=5, type="double", help="how many years should the parameterization interval be")
-parser$add_argument("-t", "--tempvar", default="annual_mean_temp", help="Temperature variable to use")
-parser$add_argument("-p", "--precipvar", default="tot_annual_pp", help="Precipitation variable to use")
-parser$add_argument("-c", "--coldesign", default="1111111", help="7-character design vector for colonization model")
-parser$add_argument("-e", "--extdesign", default="1111111", help="7-character design vector for extinction model")
+library(GenSA)
+targetInterval = 5
+modType = "rf"
 
-
-argList = parser$parse_args()
-spName = argList$species
-targetInterval = argList$interval
-colDesign = sapply(1:nchar(argList$coldesign), function(i) as.integer(substr(argList$coldesign,i,i)))
-extDesign = sapply(1:nchar(argList$extdesign), function(i) as.integer(substr(argList$extdesign,i,i)))
+parameters = commandArgs(trailingOnly = TRUE)
+#parameters = scan("inp.txt", what=character())
+spName = parameters[1]
+tempVar = parameters[2]
+precipVar = parameters[3]
+# adding a one to the beginning for the intercept
+colDesign = c(1, sapply(1:nchar(parameters[4]), function(i) as.integer(substr(parameters[4],i,i))))
+extDesign = c(1, sapply(1:nchar(parameters[5]), function(i) as.integer(substr(parameters[5],i,i))))
 
 design_err = function(x)
 {
-	err = length(x) != 7 | setequal(x, c(0,1))
+	err = length(x) != 7 | (!setequal(x, c(0,1)) & !setequal(x, 1))
 }
 if(design_err(colDesign) | design_err(extDesign)) 
[…]

### 2015-06-19 scr/1_prep_data.r
Set up a directory setup script

@@ -1,13 +1,6 @@
 #!/usr/bin/Rscript
 
-library(argparse)
-# handle command line arguments
-parser = ArgumentParser()
-parser$add_argument("-s", "--species", default="28731-ACE-SAC", help="desired species code")
-argList = parser$parse_args()
-spName = argList$species
-# set seed - drawn from sample(1:1e6, 1)
-## set.seed(588533)
+
 
 #-------------------
 #  Load and prepare data

### 2015-06-22 scr/1_prep_data.r
reorganized the first script

@@ -1,64 +1,84 @@
 #!/usr/bin/Rscript
+library(ade4)
+library(raster)
 
 
-
-#-------------------
-#  Load and prepare data
-#-------------------
-
-infile = paste("dat/", spName, "/transition_twostate_", spName, ".rdata", sep="")
-load(infile)
-
-# get and scale the climate variables
-colnames(stateData)[which(colnames(stateData) == spName)] = "presence"
-climVars = which(!(colnames(stateData) %in% c('presence', 'plot_id', 'year_measured', 'lat', 'lon')))
-climVarNames = colnames(stateData)[climVars]
-stateData.clim.scaled = scale(stateData[,climVars])
-
-# save the transformation as a function
-scalefunc = function(sc_mean, sc_sd) {
-	return(list(
-		scale = function(dat)
-		{
-			for(nm in names(sc_mean)) dat[,nm] = (dat[,nm] - sc_mean[nm])/sc_sd[nm]
-			return(dat)
-		},
-		unscale = function(dat)
-		{
-			for(nm in names(sc_mean)) dat[,nm] = dat[,nm]*sc_sd[nm] + sc_mean[nm]
-			return(dat)
-		},
-		means = sc_mean,
-		sds = sc_sd))
-}
-clim.scale = scalefunc(attr(stateData.clim.scaled, "scaled:center"), 
-attr(stateData.clim.scaled, "scaled:scale"))
-
-stateData.scaled = cbind(stateData[,-climVars], stateData.clim.scaled)
-transitionData.scaled = clim.scale$scale(transitionData)
-
-# remove extremely long sampling intervals from the dataset 
-# (to avoid the possibility of 2 transitions in the space we missed)
-intervals = transitionData.scaled$year2 - transitionData.scaled$year1
-indices = which(intervals <= 15)
-transitionData.scaled = transitionData.scaled[indices,]
+# set up map projections
+P4S.latlon = CRS("+proj=longlat +datum=WGS84")
+stmMapProjection = CRS("+init=epsg:6350") # albers equal area conic NAD-83 north america
+save(P4S.latlon, stmMapProjection, file="dat/map_projections.rdata")
 
 
-# subset the data for the SDM
-# select only a single row for each plot (to avoid too much spatial duplication)
-rows = sapply(unique(stateData.scaled$plot_id), function(i) {
[…]

### 2015-06-22 scr/1_prep_data.r
Removed raster processing from the first step

@@ -63,20 +63,6 @@ trClim.scaled = cbind(transitionClimDat[,1:3], trClim.scaled)
 climGrid.unscaled = cbind(climGrid.raw[,1:2], climGrid.unscaled)
 climGrid.scaled = cbind(climGrid.raw[,1:2], climGrid.scaled)
 
-# make the climGrids into projected raster brick objects
-make_brick = function(x)
-{
-	coordinates(x) = c('lon', 'lat')
-	gridded(x) = TRUE
-	x.ras = brick(x)
-	setMinMax(x.ras)
-	proj4string(x.ras) = P4S.latlon
-	projectRaster(x.ras, crs=stmMapProjection)
-}
-climGrid.unscaled = make_brick(climGrid.unscaled)
-climGrid.scaled = make_brick(climGrid.scaled)
-
-
 ## save climate variables
 saveRDS(climVars.scaled, "dat/plotClimate_scaled.rds")
 saveRDS(trClim.scaled, "dat/transitionClimate_scaled.rds")

### 2015-06-22 scr/2_fit_sdm.r
Reorganized sdm scripts

@@ -1,90 +1,99 @@
 #!/usr/bin/Rscript
-
-library(argparse)
-# handle command line arguments
-parser = ArgumentParser()
-parser$add_argument("-s", "--species", default="28731-ACE-SAC", help="desired species code")
-parser$add_argument("-p", "--power", default=2, type="integer", help="max exponent for model fitting")
-argList = parser$parse_args()
-spName = argList$species
-
-# set seed - drawn from sample(1:1e6, 1)
-## set.seed(588533)
-infile = paste("dat/", spName, "/", spName, "_processed.rdata", sep="")
-load(infile)
-
-
-# --------------------
-#  variable selection 
-# --------------------
-
-# select vars to include in the SDM
-# mean annual temp and total precip are included by default because we are using them as  
-# the environmental variables in the CE model
-
-# PCA
-library(ade4)
-var.pca = dudi.pca(stateData.subset[,climVarNames], scannf=FALSE, nf = 5)
-var.pca$eig / sum(var.pca$eig)
-varCor = cor(stateData.subset[,climVarNames])
-
-contrib = inertia.dudi(var.pca, row = FALSE, col = TRUE)$col.abs
-
-# procedure:
-# examine nonCorVars at each step
-# choose one with high inertia (in contrib) for the next step
-# repeat until all non-correlated variables are included in selectedVars (below)
-nonCorVars = intersect(names(varCor[which(abs(varCor[,"annual_mean_temp"])<0.7),
-		"annual_mean_temp"]), names(varCor[which(abs(varCor[,"tot_annual_pp"])<0.7),
-		"tot_annual_pp"]))
-contrib[nonCorVars,]
-
-nonCorVars = intersect(nonCorVars, names(varCor[which(abs(varCor[,"pp_seasonality"])<0.7),"pp_seasonality"]))
-contrib[nonCorVars,]
-nonCorVars = intersect(nonCorVars, names(varCor[which(abs(varCor[,"pp_driest_period"])<0.7),"pp_driest_period"]))
-contrib[nonCorVars,]
-nonCorVars = intersect(nonCorVars, names(varCor[which(abs(varCor[,"pp_warmest_quarter"])<0.7),"pp_warmest_quarter"]))
-contrib[nonCorVars,]
[…]

### 2015-06-22 scr/3_project_sdm.r

@@ -1,137 +1,0 @@
-#!/usr/bin/Rscript
-
-library(argparse)
-library(gam)
-library(randomForest)
-parser = ArgumentParser()
-parser$add_argument("-s", "--species", default="28731-ACE-SAC", help="desired species code")
-parser$add_argument("-f", "--fraction", default=0.5, type="double", help="proportion of data to use for model fitting")
-argList = parser$parse_args()
-spName = argList$species
-
-load(paste("dat/", spName, "/", spName, "_processed.rdata", sep=""))
-load(paste("results/", spName, "/", spName, "_sdm_models.rdata", sep=""))
-climGrid = readRDS(paste("dat/", spName, "/", spName, "_climGrid_scaled.rds", sep=""))
-# load(paste("dat/", spName, "/", spName, "_sdmData.rdata", sep=""))
-
-
-# project the SDM prob of presence to the plots
-transitionData.projected = transitionData.scaled
-transitionData.projected$expectedGLM = predict(glm.mod, newdata = transitionData.scaled, type='response')
-transitionData.projected$expectedGAM = predict(gam.mod, newdata = transitionData.scaled, type='response')
-transitionData.projected$expectedRF = predict(rf.mod, newdata = transitionData.scaled, type='prob')[,2]
-saveRDS(transitionData.projected, paste("dat/", spName, "/", spName, "_transitions_projected.rds", sep=""))
-
-
-# subset the transition data for use in the annealing
-if(argList$fraction < 1)
-{
-	sel = sample(nrow(transitionData.projected), 
-			as.integer(argList$fraction*nrow(transitionData.projected)))
-} else {
-	sel = 1:nrow(transitionData.projected)
-}
-transitionData.subsample = transitionData.projected[sel,]
-saveRDS(transitionData.subsample, paste("dat/", spName, "/", spName, "_transitions_projected_subsample.rds", sep=""))
-
-
-
-# project the SDMs to the climate grid
-climGrid$glm.predict = predict(glm.mod, newdata=climGrid[,selectedVars], type='response')
[…]

### 2015-06-22 scr/4_fit_stm.r

@@ -1,3 +1,16 @@
+
+## need to work this in to this script somewhere
+# remove extremely long sampling intervals from the dataset 
+# (to avoid the possibility of 2 transitions in the space we missed)
+intervals = transitionData.scaled$year2 - transitionData.scaled$year1
+indices = which(intervals <= 15)
+transitionData.scaled = transitionData.scaled[indices,]
+
+
+
+
+
+
 #!/usr/bin/Rscript
 library(GenSA)
 targetInterval = 5

### 2015-06-22 scr/2_fit_sdm.r
Made a few revisions to speed up the SDM

@@ -4,6 +4,7 @@ library(pROC)
 library(raster)
 library(rgdal)
 
+print("Setting up variables and reading files")
 annealFraction = 0.5
 
 speciesList = readRDS('dat/speciesList.rds')
@@ -15,8 +16,19 @@ ocean = spTransform(ocean, stmMapProjection)
 
 climDat = readRDS("dat/plotClimate_scaled.rds")
 
+
+# subset the data for the SDM
+# select only a single row for each plot (to avoid too much spatial duplication)
+# note that all species get the same subset (will be applied during the species loop)
+rows = sapply(unique(sdmDat$plot_id), function(i) {
+	candidates = which(sdmDat$plot_id == i)
+	if(length(candidates) == 1) candidates else sample(candidates, 1)
+	})
+
+
 for(spName in speciesList)
 {
+	print(paste("Starting species", spName))
 	baseDir = file.path('species', spName)
 
 	spPath = file.path(baseDir, 'dat', paste(spName, 'presence.rds', sep='_'))
@@ -24,31 +36,29 @@ for(spName in speciesList)
 	colnames(spData)[3] = 'presence'
 	sdmDat = merge(spData, climDat)
 
-	# subset the data for the SDM
-	# select only a single row for each plot (to avoid too much spatial duplication)
-	# drop the plot ID and year measured columns also
-	rows = sapply(unique(sdmDat$plot_id), function(i) {
-		candidates = which(sdmDat$plot_id == i)
-		if(length(candidates) == 1) candidates else sample(candidates, 1)
-		})
+	# apply the subset, and drop the plot ID and year measured columns also
 	sdmDat.subset = sdmDat[rows,-c(1,2)]
 	sdmDat.unsubset = sdmDat[-rows,-c(1,2)]
 
+	print("  Fitting RF model")
 	rf.mod = randomForest(as.factor(presence) ~ . , data = sdmDat.subset, ntree = 500)
 	saveRDS(rf.mod, file.path(baseDir, 'res', paste(spName, 'sdm.rds', sep='_'))
 
 	# see how well the rf fits
+	print("  Computing ROC")
 	ubsubset.predictions = predict(rf.mod, newdata = sdmDat.unsubset, type='prob')[,2]
 	rf.roc = roc(response=sdmDat.unsubset$presence, predictor=ubsubset.predictions)
[…]

### 2015-06-22 scr/1_prep_data.r
Updated the CRS and fixed a small bug

@@ -1,11 +1,11 @@
 #!/usr/bin/Rscript
 library(ade4)
 library(raster)
-
+library(rgdal)
 
 # set up map projections
 P4S.latlon = CRS("+proj=longlat +datum=WGS84")
-stmMapProjection = CRS("+init=epsg:6350") # albers equal area conic NAD-83 north america
+stmMapProjection = CRS("+init=epsg:5070") # albers equal area conic NAD-83 north america
 save(P4S.latlon, stmMapProjection, file="dat/map_projections.rdata")
 
 

### 2015-06-22 scr/2_fit_sdm.r

@@ -20,8 +20,8 @@ climDat = readRDS("dat/plotClimate_scaled.rds")
 # subset the data for the SDM
 # select only a single row for each plot (to avoid too much spatial duplication)
 # note that all species get the same subset (will be applied during the species loop)
-rows = sapply(unique(sdmDat$plot_id), function(i) {
-	candidates = which(sdmDat$plot_id == i)
+rows = sapply(unique(climDat$plot_id), function(i) {
+	candidates = which(climDat$plot_id == i)
 	if(length(candidates) == 1) candidates else sample(candidates, 1)
 	})
 

### 2015-06-22 scr/2_fit_sdm.r
Missed a parenthesis

@@ -42,7 +42,7 @@ for(spName in speciesList)
 
 	print("  Fitting RF model")
 	rf.mod = randomForest(as.factor(presence) ~ . , data = sdmDat.subset, ntree = 500)
-	saveRDS(rf.mod, file.path(baseDir, 'res', paste(spName, 'sdm.rds', sep='_'))
+	saveRDS(rf.mod, file.path(baseDir, 'res', paste(spName, 'sdm.rds', sep='_')))
 
 	# see how well the rf fits
 	print("  Computing ROC")

### 2015-06-22 scr/2_fit_sdm.r
Moved some more preprocessing to script 2

@@ -66,7 +66,15 @@ for(spName in speciesList)
 
 
 	# subset the transition data for use in the annealing
-	sel = sample(nrow(stmData), as.integer(annealFraction*nrow(stmData)))
+	# first, drop all intervals greater than 15 years
+	# then take half (or whatever fraction) of the remaining observed transitions
+	# along with the same fraction of non-transitions
+	intervals = stmData$year2 - stmData$year1
+	indices = which(intervals <= 15)
+	transitions = with(stmData[indices,], which(state1 != state2))
+	notTransitions = with(stmData[indices,], which(state1 == state2))
+	sel = c(sample(transitions, as.integer(annealFraction*length(transitions))),
+			sample(notTransitions, as.integer(annealFraction*length(notTransitions))))
 	stmData.subset = stmData[sel,]
 	stmData.unsubset = stmData[-sel,]
 	saveRDS(stmData.subset, file.path(baseDir, 'dat', paste(spName, 'stm', 'calib.rds', sep='_')))

### 2015-06-24 scr/3_anneal_queue/abibal.sh
Revised the anneal model selection procedure to exclude cubic models and to vary the design between col and ext

@@ -1,10 +1,10 @@
 #!/bin/sh
 #PBS -q default
 #PBS -l walltime=24:00:00
-#PBS -l nodes=1:ppn=30
+#PBS -l nodes=1:ppn=40
 #PBS -r n
-#PBS -N querub
+#PBS -N abibal
 
 DIR=~/STModel-Two-State
-SPECIES=19408-QUE-RUB
+SPECIES=18032-ABI-BAL
 source $DIR/scr/3a_stm_model_selection.sh

### 2015-06-24 scr/3_anneal_queue/acerub.sh

@@ -1,10 +1,10 @@
 #!/bin/sh
 #PBS -q default
 #PBS -l walltime=24:00:00
-#PBS -l nodes=1:ppn=30
+#PBS -l nodes=1:ppn=40
 #PBS -r n
-#PBS -N frapen
+#PBS -N acerub
 
 DIR=~/STModel-Two-State
-SPECIES=32929-FRA-PEN
+SPECIES=28728-ACE-RUB
 source $DIR/scr/3a_stm_model_selection.sh

### 2015-06-24 scr/3_anneal_queue/acesac.sh

@@ -1,10 +1,10 @@
 #!/bin/sh
 #PBS -q default
 #PBS -l walltime=24:00:00
-#PBS -l nodes=1:ppn=30
+#PBS -l nodes=1:ppn=40
 #PBS -r n
-#PBS -N picmar
+#PBS -N acesac
 
 DIR=~/STModel-Two-State
-SPECIES=183302-PIC-MAR
+SPECIES=28731-ACE-SAC
 source $DIR/scr/3a_stm_model_selection.sh

### 2015-06-24 scr/3_anneal_queue/betall.sh

@@ -1,10 +1,10 @@
 #!/bin/sh
 #PBS -q default
 #PBS -l walltime=24:00:00
-#PBS -l nodes=1:ppn=30
+#PBS -l nodes=1:ppn=40
 #PBS -r n
-#PBS -N quealb
+#PBS -N betall
 
 DIR=~/STModel-Two-State
-SPECIES=19290-QUE-ALB
+SPECIES=19481-BET-ALL
 source $DIR/scr/3a_stm_model_selection.sh

### 2015-06-24 scr/3_anneal_queue/betpap.sh

@@ -1,10 +1,10 @@
 #!/bin/sh
 #PBS -q default
 #PBS -l walltime=24:00:00
-#PBS -l nodes=1:ppn=30
+#PBS -l nodes=1:ppn=40
 #PBS -r n
-#PBS -N abibal
+#PBS -N betpap
 
 DIR=~/STModel-Two-State
-SPECIES=18032-ABI-BAL
+SPECIES=19489-BET-PAP
 source $DIR/scr/3a_stm_model_selection.sh

### 2015-06-24 scr/3_anneal_queue/faggra.sh

@@ -1,10 +1,10 @@
 #!/bin/sh
 #PBS -q default
 #PBS -l walltime=24:00:00
-#PBS -l nodes=1:ppn=30
+#PBS -l nodes=1:ppn=40
 #PBS -r n
-#PBS -N betpap
+#PBS -N faggra
 
 DIR=~/STModel-Two-State
-SPECIES=19489-BET-PAP
+SPECIES=19462-FAG-GRA
 source $DIR/scr/3a_stm_model_selection.sh

### 2015-06-24 scr/3_anneal_queue/fraame.sh

@@ -1,10 +1,10 @@
 #!/bin/sh
 #PBS -q default
 #PBS -l walltime=24:00:00
-#PBS -l nodes=1:ppn=30
+#PBS -l nodes=1:ppn=40
 #PBS -r n
-#PBS -N lirtul
+#PBS -N fraame
 
 DIR=~/STModel-Two-State
-SPECIES=18086-LIR-TUL
+SPECIES=32931-FRA-AME
 source $DIR/scr/3a_stm_model_selection.sh

### 2015-06-24 scr/3_anneal_queue/frapen.sh

@@ -1,10 +1,10 @@
 #!/bin/sh
 #PBS -q default
 #PBS -l walltime=24:00:00
-#PBS -l nodes=1:ppn=30
+#PBS -l nodes=1:ppn=40
 #PBS -r n
-#PBS -N acesac
+#PBS -N frapen
 
 DIR=~/STModel-Two-State
-SPECIES=28731-ACE-SAC
+SPECIES=32929-FRA-PEN
 source $DIR/scr/3a_stm_model_selection.sh

### 2015-06-24 scr/3_anneal_queue/lirtul.sh

@@ -1,10 +1,10 @@
 #!/bin/sh
 #PBS -q default
 #PBS -l walltime=24:00:00
-#PBS -l nodes=1:ppn=30
+#PBS -l nodes=1:ppn=40
 #PBS -r n
-#PBS -N pintae
+#PBS -N lirtul
 
 DIR=~/STModel-Two-State
-SPECIES=18037-PIN-TAE
+SPECIES=18086-LIR-TUL
 source $DIR/scr/3a_stm_model_selection.sh

### 2015-06-24 scr/3_anneal_queue/nyssyl.sh

@@ -1,10 +1,10 @@
 #!/bin/sh
 #PBS -q default
 #PBS -l walltime=24:00:00
-#PBS -l nodes=1:ppn=30
+#PBS -l nodes=1:ppn=40
 #PBS -r n
-#PBS -N ulmame
+#PBS -N nyssyl
 
 DIR=~/STModel-Two-State
-SPECIES=19049-ULM-AME
+SPECIES=27821-NYS-SYL
 source $DIR/scr/3a_stm_model_selection.sh

### 2015-06-24 scr/3_anneal_queue/picgla.sh

@@ -1,7 +1,7 @@
 #!/bin/sh
 #PBS -q default
 #PBS -l walltime=24:00:00
-#PBS -l nodes=1:ppn=30
+#PBS -l nodes=1:ppn=40
 #PBS -r n
 #PBS -N picgla
 

### 2015-06-24 scr/3_anneal_queue/pinstr.sh

@@ -1,10 +1,10 @@
 #!/bin/sh
 #PBS -q default
 #PBS -l walltime=24:00:00
-#PBS -l nodes=1:ppn=30
+#PBS -l nodes=1:ppn=40
 #PBS -r n
-#PBS -N faggra
+#PBS -N pinstr
 
 DIR=~/STModel-Two-State
-SPECIES=19462-FAG-GRA
+SPECIES=183385-PIN-STR
 source $DIR/scr/3a_stm_model_selection.sh

### 2015-06-24 scr/3_anneal_queue/poptre.sh

@@ -1,10 +1,10 @@
 #!/bin/sh
 #PBS -q default
 #PBS -l walltime=24:00:00
-#PBS -l nodes=1:ppn=30
+#PBS -l nodes=1:ppn=40
 #PBS -r n
-#PBS -N quevel
+#PBS -N poptre
 
 DIR=~/STModel-Two-State
-SPECIES=19447-QUE-VEL
+SPECIES=195773-POP-TRE
 source $DIR/scr/3a_stm_model_selection.sh

### 2015-06-24 scr/3_anneal_queue/quenig.sh

@@ -1,10 +1,10 @@
 #!/bin/sh
 #PBS -q default
 #PBS -l walltime=24:00:00
-#PBS -l nodes=1:ppn=30
+#PBS -l nodes=1:ppn=40
 #PBS -r n
-#PBS -N nyssyl
+#PBS -N quenig
 
 DIR=~/STModel-Two-State
-SPECIES=27821-NYS-SYL
+SPECIES=19280-QUE-NIG
 source $DIR/scr/3a_stm_model_selection.sh

### 2015-06-24 scr/3_anneal_queue/quevel.sh

@@ -1,10 +1,10 @@
 #!/bin/sh
 #PBS -q default
 #PBS -l walltime=24:00:00
-#PBS -l nodes=1:ppn=30
+#PBS -l nodes=1:ppn=40
 #PBS -r n
-#PBS -N quenig
+#PBS -N quevel
 
 DIR=~/STModel-Two-State
-SPECIES=19280-QUE-NIG
+SPECIES=19447-QUE-VEL
 source $DIR/scr/3a_stm_model_selection.sh

### 2015-06-24 scr/3_anneal_queue/ulmame.sh

@@ -1,10 +1,10 @@
 #!/bin/sh
 #PBS -q default
 #PBS -l walltime=24:00:00
-#PBS -l nodes=1:ppn=30
+#PBS -l nodes=1:ppn=40
 #PBS -r n
-#PBS -N fraame
+#PBS -N ulmame
 
 DIR=~/STModel-Two-State
-SPECIES=32931-FRA-AME
+SPECIES=19049-ULM-AME
 source $DIR/scr/3a_stm_model_selection.sh

### 2015-06-24 scr/3_qsub_jobs1.sh

@@ -4,5 +4,5 @@ declare -a ScriptList=(abibal acesac picgla picmar ulmame fraame quealb querub f
 
 for scr in ${ScriptList[@]}
 do
-	cd $DIR; qsub ./scr/queue/"$scr".sh
+	cd $DIR; qsub ./scr/3_anneal_queue/"$scr".sh
 done

### 2015-06-24 scr/3_qsub_jobs2.sh

@@ -4,5 +4,5 @@ declare -a ScriptList=(nyssyl acerub lirtul quevel quenig pintae frapen pinban p
 
 for scr in ${ScriptList[@]}
 do
-	cd $DIR; qsub ./scr/queue/"$scr".sh
+	cd $DIR; qsub ./scr/3_anneal_queue/"$scr".sh
 done

### 2015-06-24 scr/3a_stm_model_selection.sh

@@ -1,6 +1,6 @@
 if [ -z "$SPECIES" ]; then
-    echo "Error: species not set for model selection"
-    exit 1
+  echo "Error: species not set for model selection"
+  exit 1
 fi
 
 DIR=~/STModel-Two-State
@@ -8,14 +8,14 @@ DIR=~/STModel-Two-State
 declare -a TempVars=(annual_mean_temp mean_diurnal_range mean_temp_wettest_quarter)
 declare -a PrecipVars=(tot_annual_pp pp_seasonality pp_warmest_quarter)
 declare -a AllVars=( ${TempVars[@]} ${PrecipVars[@]} )
-declare -a DesignStrings=(111111 111110 111100 110111 110110 110100 100111 100110)
-declare -a SingleVar=(110000 111000)
+declare -a DesignStrings=(110110 110100 100110)
+SingleVar=(110000 100000)
 
 RUNNING=0
 DONE=0
-MAXCORES=30
+MAXCORES=40
 
-# start with an intercept-only model
+# start with an intercept-only model - 1 model
 TEMP=NA
 PRECIP=NA
 COL=000000
@@ -23,30 +23,38 @@ cd $DIR; Rscript ./scr/3b_fit_stm.r $SPECIES $TEMP $PRECIP $COL $COL &
 RUNNING=$((RUNNING + 1)) 
 
 # loop through all single variable models
+# 6 variables * 2 col * 2 ext models each = 24 models
 for TEMP in ${AllVars[@]}
 do
+  for COL in ${SingleVar[@]}
+  do
     for COL in ${SingleVar[@]}
     do
-        cd $DIR; Rscript ./scr/3b_fit_stm.r $SPECIES $TEMP $PRECIP $COL $COL &
-        RUNNING=$((RUNNING + 1))
-    done
+      cd $DIR; Rscript ./scr/3b_fit_stm.r $SPECIES $TEMP $PRECIP $COL $EXT &
+      RUNNING=$((RUNNING + 1))
+  done
 done
 
+# 3 temp * 3 precip * 3 col models * 3 ext models = 81 models
+# with previous, we have 106 total models
 for TEMP in ${TempVars[@]}
 do
-    for PRECIP in ${PrecipVars[@]}
+  for PRECIP in ${PrecipVars[@]}
+  do
+    for COL in ${DesignStrings[@]}
     do
-        for COL in ${DesignStrings[@]}
-        do
[…]

### 2015-06-24 scr/3a_stm_model_selection.sh
Fixed a bug in the model selection launch script

@@ -28,10 +28,11 @@ for TEMP in ${AllVars[@]}
 do
   for COL in ${SingleVar[@]}
   do
-    for COL in ${SingleVar[@]}
+    for EXT in ${SingleVar[@]}
     do
       cd $DIR; Rscript ./scr/3b_fit_stm.r $SPECIES $TEMP $PRECIP $COL $EXT &
       RUNNING=$((RUNNING + 1))
+    done
   done
 done
 

### 2015-06-24 tex/manuscript.tex
Reorganized sdm scripts
Added some text to the ms and polished up script 4

@@
 
-%\section*{Introduction}
+\section*{Introduction}
+\subsection*{Objectives}
+\begin{itemize}
+	\item Describe the range limits at equilibrium of ecologically important trees in eastern North America as predicted by metapopulation theory
+	\item Evaluate the relationship between macroclimatic variables and colonization and extinction rates
+	\item Test quantitatively whether the equilibrium prediction matches the contemporary observed range
+	\item Compare among species the predicted resilience to environmental change 
+\end{itemize}
+
+Present and future rapidly changing global environmental conditions are likely to push many species to the limits of their ecological niches.
+Biogeographical theory and empirical work suggests that species can respond to these changes by adapting \emph{in situ} to the new conditions, geographically tracking ideal conditions (resulting in range shifts), and/or experiencing local extinction where conditions no longer support presence.
+Species distribution models (SDMs) are commonly-applied to understand the latter two options.
+Classical SDMs often use a bioclimatic modeling approach that statistically relates presence or absence of the species to climate.
+Such models are quite powerful when describing the present realized niche of a species, however they suffer from several drawbacks when predicting the future distribution of species, and they provide no insight into the mechanisms underlying a species' response to the environment.
+An alternative more mechanistic approach is to examine the colonization rate $c$ of unoccupied patches and the extinction rate $e$ of occupied patches.
+(here maybe briefly review the literature on the subject)
+In such a stochastic system, the long-term expectation is that presence will only be possible where $c > e$.
+It follows that, if these rates vary along an environmental gradient, the equilibrium position of the range limit on the gradient will be the point at which $c = e$.
+
+We designed a colonization-extinction model with a dynamic parameterization that allowed the colonization and extinction rates to vary as a function of climate.
+A major challenge to employing these models is the lack of time series data needed to estimate the parameters.
+We used a large database of permanent forest plots, containing repeated observations of xxx plots in eastern North America (n=xxx paired observations). 
+We selected the 21 most abundant species (based on the number of observations in the database) as the focus of this study.
+We performed model selection, using 6 macroclimatic variables, then parameterized the model using Markov Chain Monte Carlo (MCMC) sampling and addressed the following questions for each selected species:
+\begin{enumerate}
+	\item Are range limits driven by colonization-extinction dynamics?
+	\item To what extent are the species studied presently out of equilibrium with the environment?
+	\item How resilient are range limits to perturbations in the environment?
+	\item Which climatic variables best describe colonization and extinction rates?
+\end{enumerate}
+
+%% here expand a bit on each question above and perhaps provide a testable hypothesis.
+We evaluated the model by reserving half of the observed sample pairs.
+We parameterized the model and performed model selection on the calibration half, then used the parameters to predict the second state as a function of the climate and the first state in the evaluation dataset.
+We then compared the predicted to the observed second state and computed TSS and ROC (explanation and citation?).
+
+%\section*{Results}
+
+%\section*{Discussion}
+
+% can this apply to other organisms? animals prob not - dispersal is just too high
+% this deserves a paragraph here
+
+\subsection*{Conclusions}
+\begin{itemize}
+	\item Range limits of [some/many/most/all] [temperate/boreal] tree species in northeastern north America are determined by colonization-extinction dynamics
+	\item Moisture and energy (i.e., mean annual temperature and precipitation) are associated with colonization and extinction rates for many species
+	\item For some species, the equilibrium state of the model did not match the predicted distribution of the species, suggesting that these species are either out of equilibrium with the environment, or that other processes (e.g., large-scale disturbances) contribute to their range dynamics.
+	\item Resilience (represented by the dominant eigenvalue (c - e) near the transition point) varies widely among species, suggesting greatly differential rates of response to environmental change.
+	\item Validation? Model Selection?
+\end{itemize}
+
+\subsection*{List of figures}
+\begin{itemize}
+	\item Fig 1: theoretical/introductory figure or diagram
+	\item Fig 2: posterior response curves for c and e w.r.t climate
+	\item Fig 3: maps for selected species (with remaining in appendix) - 4 panels - c; e; predicted range; SDM
+	\item Fig 4: eigenvalue maps
+\end{itemize}
+
+\subsection*{Tables}
+\input{table}
 
@@
 
-%\section*{Results}
-
-%\section*{Discussion}
-
-\subsection*{Conclusions}
-\begin{itemize}
-	\item Range limits of [some/many/most/all] [temperate/boreal] tree species in northeastern north America are determined by colonization-extinction dynamics
-	\item Moisture and energy (i.e., mean annual temperature and precipitation) are associated with colonization and extinction rates for many species
-	\item For some species, the equilibrium state of the model did not match the predicted distribution of the species, suggesting that these species are either out of equilibrium with the environment, or that other processes (e.g., large-scale disturbances) contribute to their range dynamics.
-	\item Resilience (represented by the dominant eigenvalue (c - e) near the transition point) varies widely among species, suggesting greatly differential rates of response to environmental change.
-	\item Validation? Model Selection?
-\end{itemize}
-
-\subsection*{List of figures}
-\begin{itemize}
-	\item Fig 1: theoretical/introductory figure or diagram
-	\item Fig 2: posterior response curves for c and e w.r.t climate
-	\item Fig 3: maps for selected species (with remaining in appendix) - 4 panels - c; e; predicted range; SDM
-	\item Fig 4: eigenvalue maps
-\end{itemize}
-
-\subsection*{Tables}
-\input{table}
-
 \end{document}

### 2015-06-24 scr/4_prep_mcmc.r

@@ -19,7 +19,6 @@ compute_c = function(p, env1, env2)
 spList = readRDS('dat/speciesList.rds')
 for(spName in spList)
 {
-spName = '18032-ABI-BAL'
 	cat(paste("Starting species", spName, "\n"))
 	baseDir = file.path('species', spName)
 	annealDir = file.path(baseDir, 'res', 'anneal')
@@ -31,57 +30,63 @@ spName = '18032-ABI-BAL'
 	# compute AUC
 	calibrationSet = readRDS(file.path(baseDir, 'dat', paste(spName, 'stm_calib.rds', sep='_')))
 	validationSet = readRDS(file.path(baseDir, 'dat', paste(spName, 'stm_valid.rds', sep='_')))
+
+	prep_prediction = function(x, p, e1, e2)
+	{
+		if(e1 == "NA" | is.na(e1)) {
+			env1 = rep(0, nrow(x))
+		} else {
+			env1 = x[,e1]
+		}
+		if(e2 == "NA" | is.na(e2)) {
+			env2 = rep(0, nrow(x))
+		} else {
+			env2 = x[,e2]
+		}
+		
+		interval = x$year2 - x$year1
+		c.rate = compute_c(p, env1, env2)
+		e.rate = compute_e(p, env1, env2)
+		
+		# get the interval rates (duh!)
+		# the parameters were estimated for a 5 year interval
+		c.rate = 1 - (1 - c.rate)^(interval/5)
+		e.rate = 1 - (1 - e.rate)^(interval/5)
+
+		response = x$state2
+		prediction = rep(0, length(response))
+		
+		prediction[x$state1 == 0] = (c.rate * x$prevalence)[x$state1 == 0]
+		prediction[x$state1 == 1] = (1 - e.rate)[x$state1 == 1]
+		prediction = 1000*prediction # biomod expects data from 1 to 1000
+		
+		data.frame(response, prediction)
+	}
+
 	compute_valid = function(design, params, env1, env2)
 	{
-		if(env1 == "NA" | is.na(env1))
-		{
-			ce1= ve1 = rep(0, nrow(validationSet))
-		} else 
-		{
-			ce1 = calibrationSet[,env1]
-			ve1 = validationSet[,env1]
-		}
-		if(env2 == "NA" | is.na(env2)) {
-			ce2 = ve2 = rep(0, length(ce1))
-		} else {
-			ce2 = calibrationSet[,env2]
-			ve2 = validationSet[,env2]
[…]

### 2015-06-25 tex/manuscript.tex
Updated figs 1 and 2 and added some ms text

@@
 We then compared the predicted to the observed second state and computed TSS and ROC (explanation and citation?).
+This evaluation can be interpreted as a test of the model iteself, and thus of the hypothesis that range limits are determined by colonization-extinction dynamics.
 
-%\section*{Results}
+To test to what extent species were out of equilibrium with the environment, we employed an independent dataset from the same database.
+This dataset consisted of plots that were only sampled once, and thus were not useful in parameterizing the CE model.
+We used the posterior predictions from the CE model to compute the probability of presence at equilibrium; we then compared this with observed presences from the dataset and computed TSS and ROC.
+For comparison with a highly fit model, we also built an SDM using the Random Forests algorithm and computed the same statistics.
+Species that are in equilibrium with their environment will show high coherence between the two models, and thus should have high validation scores in for both the CE and SDM.
+Species out of equilibrium should perform poorly for the CE model and better for the SDM.
 
-%\section*{Discussion}
+\section*{Results}
+\begin{itemize}
+	\item response of C-E to env variables
+	\item validation of models w.r.t. reserved data (how good is the model)
+	\item ability of the model to predict the range (equilibrium)
+	\item rate of response to disequilibrium
+	\item ability of model to predict response to environmental change
+\end{itemize}
+\begin{figure}
+	\includegraphics[width=5in]{../img/response_curves.pdf}
+	\caption{
+		Response of colonization (blue) and extinction (red) rates of selected species to the environment.
+		Solid lines are the median predictions of xx posterior samples, and shaded regions show 95\% Bayesian credible intervals.
+		Species not shown here can be found in supporting information.
+	}
+	\label{fig:response_curves}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=6.5in]{../img/posterior_maps.pdf}
+	\caption{
+		Maps showing the predicted colonization rate (blue, left column), extinction rate(red, second column), posterior probability of presence from the CE model (third column), and predicted probability of presence from a conventional SDM (right column).
+		Posterior probability of presence is computed as the proportion of posterior samples where colonization exceeds extinction for a given location.
+		Species not shown here can be found in supporting information.
+	}
+	\label{fig:response_curves}
+\end{figure}
+
+\section*{Discussion}
+
+This model enables us to more precisely define the niche in terms of the fundamental processes of colonization and extinction (fig..). 
+For example, the response of \emph(Acer saccharum) to temperature suggests that the southern range limit is a consequence of elevated extinction rates, whereas the northern limit is likely determined by low rates of colonization.
+The consequence of these two processes, when overlaid, is a range covering intermediate temperatures, bounded by low colonization in the north and high extinction in the south (fig - map).
 
@@
 
-\subsection*{List of figures}
+\subsection*{List of figures to make}
 \begin{itemize}
-	\item Fig 1: theoretical/introductory figure or diagram
-	\item Fig 2: posterior response curves for c and e w.r.t climate
 	\item Fig 3: maps for selected species (with remaining in appendix) - 4 panels - c; e; predicted range; SDM
-	\item Fig 4: eigenvalue maps
+	\item Fig 3: eigenvalue maps
 \end{itemize}

### 2015-06-25 scr/fig/fig1_response_curves.r

@@ -1,0 +1,109 @@
+library(coda)
+
+# temporary
+spName = '18032-ABI-BAL'
+env1 = 'annual_mean_temp'
+env2 = 'tot_annual_pp'
+
+compute_e = function(p, env1, env2)
+{
+	plogis(p[8] + env1*p[9] + env2*p[10] + env1^2*p[11] + env2^2*p[12] + env1^3*p[13] + env2^3*p[14])
+}
+
+
+compute_c = function(p, env1, env2)
+{
+	plogis(p[1] + env1*p[2] + env2*p[3] + env1^2*p[4] + env2^2*p[5] + env1^3*p[6] + env2^3*p[7])
+}
+
+
+posterior = readRDS(file.path('species', spName, 'res', paste(spName, 'posterior.rds', sep='_')))
+climGrid = readRDS('dat/climateGrid_scaled.rds')
+
+
+x1 = seq(min(climGrid[,env1]), max(climGrid[,env1]), length.out=100)
+x2 = seq(min(climGrid[,env2]), max(climGrid[,env2]), length.out=100)
+
+# now compute the y values for each x value for every posterior sample
+# first we have to choose a value at which to fix the second environmental variable
+# do this with the mean values of the parameters
+meanParams = colMeans(posterior)
+meanParams = c(meanParams[1:5], 0, 0, meanParams[6:10], 0, 0)
+# find the point at which env1 maximizes lambda
+env1.yc = compute_c(meanParams, x1, rep(0, length(x1)))
+env1.ye = compute_e(meanParams, x1, rep(0, length(x1)))
+env1.lam = env1.yc - env1.ye
+env1.maxx = x1[which(env1.lam == max(env1.lam))[1]]
+
+# find the point at which env2 maximizes lambda with the max from the prev step
+env2.yc = compute_c(meanParams, rep(env1.maxx, length(x2)), x2)
+env2.ye = compute_e(meanParams, rep(env1.maxx, length(x2)), x2)
+env2.lam = env2.yc - env2.ye
+env2.maxx = x2[which(env2.lam == max(env2.lam))[1]]
+
+# re-do the computation of x1 with the new max from x2
[…]

### 2015-06-25 scr/fig/fig2_posterior_maps.r

@@ -1,0 +1,121 @@
+library(coda)
+library(sp)
+library(raster)
+library(rgdal)
+
+# temporary
+spName = '18032-ABI-BAL'
+spText = expression(italic(Abies~balsamifera))
+env1 = 'annual_mean_temp'
+env2 = 'tot_annual_pp'
+
+load('dat/map_projections.rdata')
+
+prj.ras = function(x)
+{
+	coordinates(x) = 1:2
+	gridded(x) = TRUE
+	x = raster(x)
+	proj4string(x) = P4S.latlon
+	projectRaster(x, crs=stmMapProjection)
+}
+
+ocean = readOGR(dsn="dat/ne_50m_ocean", layer="ne_50m_ocean")
+ocean = spTransform(ocean, stmMapProjection)
+lakes = readOGR(dsn="dat/ne_50m_lakes", layer="ne_50m_lakes")
+# grab specific lakes
+lkNames = c("Huron", "Michigan", "Superior", "Ontario", "Erie", "St. Clair")
+grLakes = lakes[as.integer(sapply(lkNames, grep, lakes$name)),]
+grLakes = spTransform(grLakes, stmMapProjection)
+
+plotbg = function(txt="", rangeMap=T)
+{
+	plot(ocean, col="white", add=T)
+	plot(grLakes, col="white", add=T)
+	mtext(txt, adj=0, cex = 1)
+	
+	# plot grid lines and labels
+	llgridlines(ocean, easts=seq(-90, -40, 10), norths=seq(20,50,10), ndiscr=100, 
+			side="ES", offset=1e6, col='#44444444')
+	par(xpd=TRUE)
+#	westText = c("30°N", "40°N", "50°N")
+	eastText = c("20°N", "30°N", "40°N")
+#	westCoords = list(rep(-3e5,3), c(0.75e6, 1.9e6, 3e6))
+	eastCoords = list(rep(3.5e6,3), c(0.55e6, 1.7e6, 2.9e6))
+	southText = c("90°W", "80°W", "70°W")
+	southCoords = list(c(0.65e6, 1.7e6, 2.8e6), rep(-0.7e6,3))
+	cex.ll = 0.8
+	text(eastCoords[[1]], eastCoords[[2]], eastText, cex=cex.ll)
+#	text(westCoords[[1]], westCoords[[2]], westText, cex=cex.ll)
+	text(southCoords[[1]], southCoords[[2]], southText, cex=cex.ll)
[…]

### 2015-06-29 tex/table.tex
Added first 2 mcmc species

@@
 \toprule
-Scientific name               & Temperature Variable & Precipitation Variable & Model\tnote{1} & $\Delta$BIC\tnote{2} & BIC$_w$\tnote{3} \\ 
+Scientific name               & Temperature Variable    & Precipitation Variable    & Model\tnote{1} & $\Delta$AIC\tnote{2} & AIC$_w$\tnote{3} \\ 
 \midrule
-{\it Abies balsamea}          & env1                 & env2                   & $T^3 + P^3$    & xxx                  &                  \\
+{\it Abies balsamea}          & Mean Annual Temperature & Mean Annual Precipitation & $T^2 + P^2$    & 2.15                 & 0.74             \\
 {\it Acer rubrum}             &   &   & & \\
-{\it Acer saccharum}          &   &   & & \\
+{\it Acer saccharum}          & Mean Annual Temperature & Precip. Seasonality       & $T^2 + P$      & 0.82                 & 0.36             \\
 {\it Betula alleghaniensis}   &   &   & & \\
@@
 \begin{table}[tb]
+\label{tab:model_selection}
+\caption{Model validation statistics}
+\begin{tabular}{lllll}
+\toprule
+Scientific name               & Calibration ROC & Validation ROC & Calibration TSS & Validation TSS \\ 
+\midrule
+{\it Abies balsamea}          & 0.9770          & 0.9740          & 0.8651         & 0.8545         \\
+{\it Acer rubrum}             &   &   &  \\
+{\it Acer saccharum}          & 0.9873          & 0.9876          & 0.9365         & 0.9377         \\
+{\it Betula alleghaniensis}   &   &   &  \\
+{\it Betula papyrifera}       &   &   &  \\
+{\it Fagus grandifolia}       &   &   &  \\
+{\it Fraxinus americana}      &   &   &  \\
+{\it Fraxinus pennsylvanica}  &   &   &  \\
+{\it Liquidambar styraciflua} &   &   &  \\
+{\it Liriodendron tulipifera} &   &   & & \\
+{\it Nyssa sylvatica}         &   &   & & \\
+{\it Picea glauca}            &   &   & & \\
+{\it Pinus banksiana}         &   &   & & \\
+{\it Pinus strobus}           &   &   & & \\
+{\it Pinus taeda}             &   &   & & \\
+{\it Populus tremuloides}     &   &   & & \\ 
+{\it Quercus alba}            &   &   & & \\
+{\it Quercus nigra}           &   &   & & \\
+{\it Quercus rubra}           &   &   & & \\
+{\it Quercus velutina}        &   &   & & \\
+{\it Ulmus americana}         &   &   & & \\ 
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}[tb]
 \label{tab:parameterized_equations}

### 2015-06-29 scr/5_mcmc_scripts/abibal.sh

@@ -13,9 +13,11 @@ module load openmpi/1.8.3
 SRC=~/STModel-MCMC/bin
 DIR=~/STModel-Two-State/species/$SPECIES
 
+cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
+
 mkdir $DIR/res/mcmc
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
 
 wait

### 2015-06-29 tex/table.tex
Added two more species

--- 2015-06-19 tex/table.tex
@@
 {\it Pinus taeda}             &   &   & & \\
-{\it Populus tremuloides}     &   &   & & \\ 
-{\it Quercus alba}            &   &   & & \\
+{\it Populus tremuloides}     & Mean Annual Temperature & Mean Annual Precipitation & $T^2 + P$      & 0.65                 & 0.12        \\ 
+{\it Quercus alba}            & Mean Annual Temperature & Precip. Seasonality       & $T^2 + P^2$    & 0.15                 & 0.27            \\
 {\it Quercus nigra}           &   &   & & \\
@@
 \begin{tablenotes}
-\item [1] Model specification shows whether temperature (T) and precipitation (P) variables were included as well as the order of polynomial used.
+\item [1] HEY PAY ATTENTION: this column will have to be split into two, one for C and one for E.
+Model specification shows whether temperature (T) and precipitation (P) variables were included as well as the order of polynomial used.
 \item [2] $\Delta$BIC refers to absolute BIC difference between the selected model and the next best model in the set.
@@
 \end{threeparttable}
+\end{table}
+
+\begin{table}[tb]
+\label{tab:model_selection}
+\caption{Model validation statistics}
+\begin{tabular}{lllll}
+\toprule
+Scientific name               & Calibration ROC & Validation ROC & Calibration TSS & Validation TSS \\ 
+\midrule
+{\it Abies balsamea}          & 0.9770          & 0.9740          & 0.8651         & 0.8545         \\
+{\it Acer rubrum}             &   &   &  \\
+{\it Acer saccharum}          & 0.9873          & 0.9876          & 0.9365         & 0.9377         \\
+{\it Betula alleghaniensis}   &   &   &  \\
+{\it Betula papyrifera}       &   &   &  \\
+{\it Fagus grandifolia}       &   &   &  \\
+{\it Fraxinus americana}      &   &   &  \\
+{\it Fraxinus pennsylvanica}  &   &   &  \\
+{\it Liquidambar styraciflua} &   &   &  \\
+{\it Liriodendron tulipifera} &   &   & & \\
+{\it Nyssa sylvatica}         &   &   & & \\
+{\it Picea glauca}            &   &   & & \\
+{\it Pinus banksiana}         &   &   & & \\
+{\it Pinus strobus}           &   &   & & \\
+{\it Pinus taeda}             &   &   & & \\
+{\it Populus tremuloides}     & 0.9788          & 0.9769          & 0.8888         & 0.8830         \\ 
+{\it Quercus alba}            & 0.9901          & 0.9913          & 0.9382         & 0.9454         \\
+{\it Quercus nigra}           &   &   & & \\
+{\it Quercus rubra}           &   &   & & \\
+{\it Quercus velutina}        &   &   & & \\
+{\it Ulmus americana}         &   &   & & \\ 
+\bottomrule
+\end{tabular}
 \end{table}

### 2015-06-29 scr/5_mcmc_scripts/poptre.sh

@@ -1,0 +1,23 @@
+#!/bin/sh
+#PBS -q default
+#PBS -l walltime=336:00:00
+#PBS -l nodes=3:ppn=40
+#PBS -r n
+#PBS -N poptre
+SPECIES=195773-POP-TRE
+
+module load gcc/4.9.2
+module load openmpi/1.8.3
+
+# written for version 1.2
+SRC=~/STModel-MCMC/bin
+DIR=~/STModel-Two-State/species/$SPECIES
+
+cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
+
+mkdir $DIR/res/mcmc
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+
+wait

### 2015-06-29 scr/5_mcmc_scripts/quealb.sh

@@ -1,0 +1,23 @@
+#!/bin/sh
+#PBS -q default
+#PBS -l walltime=336:00:00
+#PBS -l nodes=3:ppn=40
+#PBS -r n
+#PBS -N quealb
+SPECIES=19290-QUE-ALB
+
+module load gcc/4.9.2
+module load openmpi/1.8.3
+
+# written for version 1.2
+SRC=~/STModel-MCMC/bin
+DIR=~/STModel-Two-State/species/$SPECIES
+
+cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
+
+mkdir $DIR/res/mcmc
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+
+wait

### 2015-06-29 tex/table.tex
Added mcmc scripts for everyone

--- 2015-06-19 tex/table.tex
@@
 {\it Fraxinus americana}      & White ash         & 10385       &             \\
-{\it Fraxinus pennsylvanica}  &                & 7662        &             \\
-{\it Liquidambar styraciflua} &                & 14778       &             \\
-{\it Liriodendron tulipifera} &                & 11045       &             \\
-{\it Nyssa sylvatica}         &                & 9209        &             \\
+{\it Fraxinus pennsylvanica}  & Red ash           & 7662        &             \\
+{\it Liriodendron tulipifera} & American sweetgum & 11045       &             \\
+{\it Nyssa sylvatica}         & Blackgum          & 9209        &             \\
 {\it Picea glauca}            & White spruce      & 8526        &             \\
+{\it Picea mariana}           & Black spruce      &             &             \\
 {\it Pinus banksiana}         & Jack pine         & 3023        &             \\
-{\it Pinus strobus}           &                & 8182        &             \\
-{\it Pinus taeda}             &                & 16351       &             \\
+{\it Pinus strobus}           & White pine        & 8182        &             \\
+{\it Pinus taeda}             & Loblolly pine     & 16351       &             \\
 {\it Populus tremuloides}     & Quaking aspen     & 12736       &             \\ 
 {\it Quercus alba}            & White oak         & 17280       &             \\
-{\it Quercus nigra}           &                & 8291        &             \\
-{\it Quercus rubra}           &                & 14337       &             \\
-{\it Quercus velutina}        &                & 9997        &             \\
+{\it Quercus nigra}           & Water oak         & 8291        &             \\
+{\it Quercus rubra}           & Northern red oak  & 14337       &             \\
+{\it Quercus velutina}        & Black oak         & 9997        &             \\
 {\it Ulmus americana}         & American elm      & 11597       &             \\ 
@@
 \toprule
-Scientific name               & Temperature Variable & Precipitation Variable & Model\tnote{1} & $\Delta$BIC\tnote{2} & BIC$_w$\tnote{3} \\ 
+Scientific name               & Temperature Variable    & Precipitation Variable    & $\Delta$AIC\tnote{1} & AIC$_w$\tnote{2} \\ 
 \midrule
-{\it Abies balsamea}          & env1                 & env2                   & $T^3 + P^3$    & xxx                  &                  \\
-{\it Acer rubrum}             &   &   & & \\
-{\it Acer saccharum}          &   &   & & \\
-{\it Betula alleghaniensis}   &   &   & & \\
-{\it Betula papyrifera}       &   &   & & \\
-{\it Fagus grandifolia}       &   &   & & \\
-{\it Fraxinus americana}      &   &   & & \\
-{\it Fraxinus pennsylvanica}  &   &   & & \\
-{\it Liquidambar styraciflua} &   &   & & \\
-{\it Liriodendron tulipifera} &   &   & & \\
-{\it Nyssa sylvatica}         &   &   & & \\
-{\it Picea glauca}            &   &   & & \\
-{\it Pinus banksiana}         &   &   & & \\
-{\it Pinus strobus}           &   &   & & \\
-{\it Pinus taeda}             &   &   & & \\
-{\it Populus tremuloides}     &   &   & & \\ 
-{\it Quercus alba}            &   &   & & \\
-{\it Quercus nigra}           &   &   & & \\
-{\it Quercus rubra}           &   &   & & \\
-{\it Quercus velutina}        &   &   & & \\
-{\it Ulmus americana}         &   &   & & \\ 
+{\it Abies balsamea}          & Mean Annual Temperature & Mean Annual Precipitation & 2.15                 & 0.74             \\
+{\it Acer rubrum}             & Mean Annual Temperature & Precip. Warmest Quarter   & 0.36                 & 0.27             \\
+{\it Acer saccharum}          & Mean Annual Temperature & Precip. Seasonality       & 0.82                 & 0.36             \\
+{\it Betula alleghaniensis}   & Mean Annual Temperature & Mean Annual Precipitation & 1.45                 & 0.35             \\
+{\it Betula papyrifera}       & Mean Annual Temperature & Precip. Seasonality       & 0.90                 & 0.32             \\
+{\it Fagus grandifolia}       & Mean Annual Temperature & Precip. Warmest Quarter   & 0.52                 & 0.41             \\
+{\it Fraxinus americana}      & Mean Annual Temperature & Mean Annual Precipitation & 0.28                 & 0.23             \\
+{\it Fraxinus pennsylvanica}  & Mean Annual Temperature & Mean Annual Precipitation & 1.80                 & 0.40             \\
+{\it Liriodendron tulipifera} & Mean Annual Temperature &  Precip. Warmest Quarter  & 0.75                 & 0.30             \\
+{\it Nyssa sylvatica}         & Mean Temp Wettest Q.    & Precip. Seasonality       & 0.35                 & 0.18             \\
+{\it Picea glauca}            & Mean Diurnal Range      & Precip. Seasonality       & 0.64                 & 0.12             \\
+{\it Picea mariana}           & Mean Annual Temperature & Precip. Warmest Quarter   & 1.39                 & 0.34             \\
+{\it Pinus banksiana}         & Mean Annual Temperature & Mean Annual Precipitation & 0.76                 & 0.22             \\
+{\it Pinus strobus}           & Mean Annual Temperature & Precip. Warmest Quarter   & 0.032                & 0.14             \\
+{\it Pinus taeda}             & Mean Annual Temperature & Mean Annual Precipitation & 0.99                 & 0.30             \\
+{\it Populus tremuloides}     & Mean Annual Temperature & Mean Annual Precipitation & 0.65                 & 0.12             \\ 
+{\it Quercus alba}            & Mean Annual Temperature & Precip. Seasonality       & 0.15                 & 0.27             \\
+{\it Quercus nigra}           & Mean Diurnal Range      & Precip. Warmest Quarter   & 0.34                 & 0.092            \\
+{\it Quercus rubra}           & Mean Annual Temperature & Mean Annual Precipitation & 1.81                 & 0.50             \\
+{\it Quercus velutina}        & Mean Annual Temperature & Precip. Warmest Quarter   & 1.11                 & 0.41             \\
+{\it Ulmus americana}         & Mean Annual Temperature & Precip. Seasonality       & 0.79                 & 0.13             \\ 
 \bottomrule
@@
 \begin{tablenotes}
-\item [1] Model specification shows whether temperature (T) and precipitation (P) variables were included as well as the order of polynomial used.
-\item [2] $\Delta$BIC refers to absolute BIC difference between the selected model and the next best model in the set.
-\item [3] BIC$_w$ can be interpreted as the probability that this model is the best model in the set (cite Burnham and Anderson).
+\item [1] $\Delta$AIC refers to absolute AIC difference between the selected model and the next best model in the set.
+\item [2] AIC$_w$ can be interpreted as the probability that this model is the best model in the set (cite Burnham and Anderson).
 \end{tablenotes}
 \end{threeparttable}
+\end{table}
+
+\begin{table}[tb]
+\label{tab:model_selection}
+\caption{Model validation statistics}
+\begin{tabular}{lllll}
+\toprule
+Scientific name               & Calibration ROC & Validation ROC & Calibration TSS & Validation TSS \\ 
+\midrule
+{\it Abies balsamea}          & 0.9770          & 0.9740          & 0.8651         & 0.8545         \\
+{\it Acer rubrum}             & 0.9690          & 0.9684          & 0.8929         & 0.8854         \\
+{\it Acer saccharum}          & 0.9873          & 0.9876          & 0.9365         & 0.9377         \\
+{\it Betula alleghaniensis}   & 0.9886          & 0.9880          & 0.9193         & 0.9208         \\
+{\it Betula papyrifera}       & 0.9791          & 0.9780          & 0.8880         & 0.8856         \\
+{\it Fagus grandifolia}       & 0.9841          & 0.9865          & 0.9060         & 0.9135         \\
+{\it Fraxinus americana}      & 0.9787          & 0.9774          & 0.8817         & 0.8741         \\
+{\it Fraxinus pennsylvanica}  & 0.9691          & 0.9740          & 0.8255         & 0.8379         \\
+{\it Liriodendron tulipifera} & 0.9925          & 0.9930          & 0.9146         & 0.9162         \\
+{\it Nyssa sylvatica}         & 0.9824          & 0.9833          & 0.8674         & 0.8622         \\
+{\it Picea glauca}            & 0.9764          & 0.9748          & 0.8443         & 0.8405         \\
+{\it Picea mariana}           & 0.9906          & 0.9893          & 0.9208         & 0.9148         \\
+{\it Pinus banksiana}         & 0.9940          & 0.9952          & 0.9405         & 0.9466         \\
+{\it Pinus strobus}           & 0.9861          & 0.9844          & 0.9091         & 0.8978         \\
+{\it Pinus taeda}             & 0.9937          & 0.9937          & 0.9303         & 0.9278         \\
+{\it Populus tremuloides}     & 0.9788          & 0.9769          & 0.8888         & 0.8830         \\ 
+{\it Quercus alba}            & 0.9901          & 0.9913          & 0.9382         & 0.9454         \\
+{\it Quercus nigra}           & 0.9897          & 0.9907          & 0.8931         & 0.8927         \\
+{\it Quercus rubra}           & 0.9812          & 0.9807          & 0.9031         & 0.8980         \\
+{\it Quercus velutina}        & 0.9860          & 0.9860          & 0.9018         & 0.8940         \\
+{\it Ulmus americana}         & 0.9635          & 0.9646          & 0.8223         & 0.8168         \\
+\bottomrule
+\end{tabular}
 \end{table}
@@
 Species                       & Colonization                               & Extinction                                  \\ \midrule
-{\it Abies balsamea} & & \\ 
+{\it Abies balsamea}          & $x.xx + x.xxT + x.xxT^2 + x.xxP + x.xxP^2$ & $x.xx + x.xxT + x.xxT^2 + x.xxP + x.xxP^2$  \\
+{\it Acer rubrum}             & $x.xx + x.xxT + x.xxP + x.xxP^2$           & $x.xx + x.xxT + x.xxT^2 + x.xxP$            \\
+{\it Acer saccharum}          & $x.xx + x.xxT + x.xxT^2 + x.xxP$           & $x.xx + x.xxT + x.xxT^2 + x.xxP$            \\
+{\it Betula alleghaniensis}   & $x.xx + x.xxT + x.xxT^2 + x.xxP$           & $x.xx + x.xxT + x.xxP$                      \\
+{\it Betula papyrifera}       & $x.xx + x.xxT + x.xxT^2 + x.xxP + x.xxP^2$ & $x.xx + x.xxT + x.xxP$                      \\
+{\it Fagus grandifolia}       & $x.xx + x.xxT + x.xxT^2 + x.xxP$           & $x.xx + x.xxT + x.xxT^2 + x.xxP + x.xxP^2$  \\
+{\it Fraxinus americana}      & $x.xx + x.xxT + x.xxT^2 + x.xxP$           & $x.xx + x.xxT + x.xxT^2 + x.xxP + x.xxP^2$  \\
+{\it Fraxinus pennsylvanica}  & $x.xx + x.xxT + x.xxT^2 + x.xxP + x.xxP^2$ & $x.xx + x.xxT + x.xxT^2 + x.xxP + x.xxP^2$  \\
+{\it Liriodendron tulipifera} & $x.xx + x.xxT + x.xxT^2 + x.xxP$           & $x.xx + x.xxT + x.xxT^2 + x.xxP$            \\
+{\it Nyssa sylvatica}         & $x.xx + x.xxT + x.xxT^2 + x.xxP + x.xxP^2$ & $x.xx + x.xxT + x.xxP + x.xxP^2$            \\
+{\it Picea glauca}            & $x.xx + x.xxT + x.xxT^2 + x.xxP + x.xxP^2$ & $x.xx + x.xxT + x.xxP + x.xxP^2$            \\
+{\it Picea mariana}           & $x.xx + x.xxT + x.xxP$                     & $x.xx + x.xxT + x.xxT^2 + x.xxP$            \\
+{\it Pinus banksiana}         & $x.xx + x.xxT + x.xxP$                     & $x.xx + x.xxT + x.xxT^2 + x.xxP$            \\
+{\it Pinus strobus}           & $x.xx + x.xxT + x.xxP$                     & $x.xx + x.xxT + x.xxT^2 + x.xxP + x.xxP^2$  \\
+{\it Pinus taeda}             & $x.xx + x.xxT + x.xxT^2 + x.xxP$           & $x.xx + x.xxT + x.xxT^2 + x.xxP$            \\
+{\it Populus tremuloides}     & $x.xx + x.xxT + x.xxT^2 + x.xxP$           & $x.xx + x.xxT + x.xxP$                      \\ 
+{\it Quercus alba}            & $x.xx + x.xxT + x.xxT^2 + x.xxP$           & $x.xx + x.xxT + x.xxT^2 + x.xxP$            \\
+{\it Quercus nigra}           & $x.xx + x.xxT + x.xxP$                     & $x.xx + x.xxT + x.xxT^2 + x.xxP$            \\
+{\it Quercus rubra}           & $x.xx + x.xxT + x.xxT^2 + x.xxP + x.xxP^2$ & $x.xx + x.xxT + x.xxT^2 + x.xxP + x.xxP^2$  \\
+{\it Quercus velutina}        & $x.xx + x.xxT + x.xxT^2 + x.xxP + x.xxP^2$ & $x.xx + x.xxT + x.xxT^2 + x.xxP$            \\
+{\it Ulmus americana}         & $x.xx + x.xxT + x.xxP$                     & $x.xx + x.xxT + x.xxT^2 + x.xxP$            \\ 
 \bottomrule

### 2015-06-29 scr/5_mcmc_scripts/acerub.sh

@@ -1,0 +1,23 @@
+#!/bin/sh
+#PBS -q default
+#PBS -l walltime=336:00:00
+#PBS -l nodes=3:ppn=40
+#PBS -r n
+#PBS -N acerub
+SPECIES=28728-ACE-RUB
+
+module load gcc/4.9.2
+module load openmpi/1.8.3
+
+# written for version 1.2
+SRC=~/STModel-MCMC/bin
+DIR=~/STModel-Two-State/species/$SPECIES
+
+cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
+
+mkdir $DIR/res/mcmc
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+
+wait

### 2015-06-29 scr/5_mcmc_scripts/betall.sh

@@ -1,0 +1,23 @@
+#!/bin/sh
+#PBS -q default
+#PBS -l walltime=336:00:00
+#PBS -l nodes=3:ppn=40
+#PBS -r n
+#PBS -N betall
+SPECIES=19481-BET-ALL
+
+module load gcc/4.9.2
+module load openmpi/1.8.3
+
+# written for version 1.2
+SRC=~/STModel-MCMC/bin
+DIR=~/STModel-Two-State/species/$SPECIES
+
+cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
+
+mkdir $DIR/res/mcmc
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+
+wait

### 2015-06-29 scr/5_mcmc_scripts/betpap.sh

@@ -1,0 +1,23 @@
+#!/bin/sh
+#PBS -q default
+#PBS -l walltime=336:00:00
+#PBS -l nodes=3:ppn=40
+#PBS -r n
+#PBS -N betpap
+SPECIES=19489-BET-PAP
+
+module load gcc/4.9.2
+module load openmpi/1.8.3
+
+# written for version 1.2
+SRC=~/STModel-MCMC/bin
+DIR=~/STModel-Two-State/species/$SPECIES
+
+cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
+
+mkdir $DIR/res/mcmc
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+
+wait

### 2015-06-29 scr/5_mcmc_scripts/faggra.sh

@@ -1,0 +1,23 @@
+#!/bin/sh
+#PBS -q default
+#PBS -l walltime=336:00:00
+#PBS -l nodes=3:ppn=40
+#PBS -r n
+#PBS -N faggra
+SPECIES=19462-FAG-GRA
+
+module load gcc/4.9.2
+module load openmpi/1.8.3
+
+# written for version 1.2
+SRC=~/STModel-MCMC/bin
+DIR=~/STModel-Two-State/species/$SPECIES
+
+cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
+
+mkdir $DIR/res/mcmc
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+
+wait

### 2015-06-29 scr/5_mcmc_scripts/fraame.sh

@@ -1,0 +1,23 @@
+#!/bin/sh
+#PBS -q default
+#PBS -l walltime=336:00:00
+#PBS -l nodes=3:ppn=40
+#PBS -r n
+#PBS -N fraame
+SPECIES=32931-FRA-AME
+
+module load gcc/4.9.2
+module load openmpi/1.8.3
+
+# written for version 1.2
+SRC=~/STModel-MCMC/bin
+DIR=~/STModel-Two-State/species/$SPECIES
+
+cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
+
+mkdir $DIR/res/mcmc
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+
+wait

### 2015-06-29 scr/5_mcmc_scripts/frapen.sh

@@ -1,0 +1,23 @@
+#!/bin/sh
+#PBS -q default
+#PBS -l walltime=336:00:00
+#PBS -l nodes=3:ppn=40
+#PBS -r n
+#PBS -N frapen
+SPECIES=32929-FRA-PEN
+
+module load gcc/4.9.2
+module load openmpi/1.8.3
+
+# written for version 1.2
+SRC=~/STModel-MCMC/bin
+DIR=~/STModel-Two-State/species/$SPECIES
+
+cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
+
+mkdir $DIR/res/mcmc
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+
+wait

### 2015-06-29 scr/5_mcmc_scripts/lirtul.sh

@@ -1,0 +1,23 @@
+#!/bin/sh
+#PBS -q default
+#PBS -l walltime=336:00:00
+#PBS -l nodes=3:ppn=40
+#PBS -r n
+#PBS -N lirtul
+SPECIES=18086-LIR-TUL
+
+module load gcc/4.9.2
+module load openmpi/1.8.3
+
+# written for version 1.2
+SRC=~/STModel-MCMC/bin
+DIR=~/STModel-Two-State/species/$SPECIES
+
+cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
+
+mkdir $DIR/res/mcmc
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+
+wait

### 2015-06-29 scr/5_mcmc_scripts/nyssyl.sh

@@ -1,0 +1,23 @@
+#!/bin/sh
+#PBS -q default
+#PBS -l walltime=336:00:00
+#PBS -l nodes=3:ppn=40
+#PBS -r n
+#PBS -N nyssyl
+SPECIES=27821-NYS-SYL
+
+module load gcc/4.9.2
+module load openmpi/1.8.3
+
+# written for version 1.2
+SRC=~/STModel-MCMC/bin
+DIR=~/STModel-Two-State/species/$SPECIES
+
+cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
+
+mkdir $DIR/res/mcmc
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+
+wait

### 2015-06-29 scr/5_mcmc_scripts/picgla.sh

@@ -1,0 +1,23 @@
+#!/bin/sh
+#PBS -q default
+#PBS -l walltime=336:00:00
+#PBS -l nodes=3:ppn=40
+#PBS -r n
+#PBS -N picgla
+SPECIES=183295-PIC-GLA
+
+module load gcc/4.9.2
+module load openmpi/1.8.3
+
+# written for version 1.2
+SRC=~/STModel-MCMC/bin
+DIR=~/STModel-Two-State/species/$SPECIES
+
+cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
+
+mkdir $DIR/res/mcmc
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+
+wait

### 2015-06-29 scr/5_mcmc_scripts/picmar.sh

@@ -1,0 +1,23 @@
+#!/bin/sh
+#PBS -q default
+#PBS -l walltime=336:00:00
+#PBS -l nodes=3:ppn=40
+#PBS -r n
+#PBS -N picmar
+SPECIES=183302-PIC-MAR
+
+module load gcc/4.9.2
+module load openmpi/1.8.3
+
+# written for version 1.2
+SRC=~/STModel-MCMC/bin
+DIR=~/STModel-Two-State/species/$SPECIES
+
+cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
+
+mkdir $DIR/res/mcmc
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+
+wait

### 2015-06-29 scr/5_mcmc_scripts/pinban.sh

@@ -1,0 +1,23 @@
+#!/bin/sh
+#PBS -q default
+#PBS -l walltime=336:00:00
+#PBS -l nodes=3:ppn=40
+#PBS -r n
+#PBS -N pinban
+SPECIES=183319-PIN-BAN
+
+module load gcc/4.9.2
+module load openmpi/1.8.3
+
+# written for version 1.2
+SRC=~/STModel-MCMC/bin
+DIR=~/STModel-Two-State/species/$SPECIES
+
+cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
+
+mkdir $DIR/res/mcmc
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+
+wait

### 2015-06-29 scr/5_mcmc_scripts/pinstr.sh

@@ -1,0 +1,23 @@
+#!/bin/sh
+#PBS -q default
+#PBS -l walltime=336:00:00
+#PBS -l nodes=3:ppn=40
+#PBS -r n
+#PBS -N pinstr
+SPECIES=183385-PIN-STR
+
+module load gcc/4.9.2
+module load openmpi/1.8.3
+
+# written for version 1.2
+SRC=~/STModel-MCMC/bin
+DIR=~/STModel-Two-State/species/$SPECIES
+
+cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
+
+mkdir $DIR/res/mcmc
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+
+wait

### 2015-06-29 scr/5_mcmc_scripts/pintae.sh

@@ -1,0 +1,23 @@
+#!/bin/sh
+#PBS -q default
+#PBS -l walltime=336:00:00
+#PBS -l nodes=3:ppn=40
+#PBS -r n
+#PBS -N pintae
+SPECIES=18037-PIN-TAE
+
+module load gcc/4.9.2
+module load openmpi/1.8.3
+
+# written for version 1.2
+SRC=~/STModel-MCMC/bin
+DIR=~/STModel-Two-State/species/$SPECIES
+
+cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
+
+mkdir $DIR/res/mcmc
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+
+wait

### 2015-06-29 scr/5_mcmc_scripts/quenig.sh

@@ -1,0 +1,23 @@
+#!/bin/sh
+#PBS -q default
+#PBS -l walltime=336:00:00
+#PBS -l nodes=3:ppn=40
+#PBS -r n
+#PBS -N quenig
+SPECIES=19280-QUE-NIG
+
+module load gcc/4.9.2
+module load openmpi/1.8.3
+
+# written for version 1.2
+SRC=~/STModel-MCMC/bin
+DIR=~/STModel-Two-State/species/$SPECIES
+
+cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
+
+mkdir $DIR/res/mcmc
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+
+wait

### 2015-06-29 scr/5_mcmc_scripts/querub.sh

@@ -1,0 +1,23 @@
+#!/bin/sh
+#PBS -q default
+#PBS -l walltime=336:00:00
+#PBS -l nodes=3:ppn=40
+#PBS -r n
+#PBS -N querub
+SPECIES=19408-QUE-RUB
+
+module load gcc/4.9.2
+module load openmpi/1.8.3
+
+# written for version 1.2
+SRC=~/STModel-MCMC/bin
+DIR=~/STModel-Two-State/species/$SPECIES
+
+cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
+
+mkdir $DIR/res/mcmc
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+
+wait

### 2015-06-29 scr/5_mcmc_scripts/quevel.sh

@@ -1,0 +1,23 @@
+#!/bin/sh
+#PBS -q default
+#PBS -l walltime=336:00:00
+#PBS -l nodes=3:ppn=40
+#PBS -r n
+#PBS -N quevel
+SPECIES=19447-QUE-VEL
+
+module load gcc/4.9.2
+module load openmpi/1.8.3
+
+# written for version 1.2
+SRC=~/STModel-MCMC/bin
+DIR=~/STModel-Two-State/species/$SPECIES
+
+cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
+
+mkdir $DIR/res/mcmc
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+
+wait

### 2015-06-29 scr/5_mcmc_scripts/ulmame.sh

@@ -1,0 +1,23 @@
+#!/bin/sh
+#PBS -q default
+#PBS -l walltime=336:00:00
+#PBS -l nodes=3:ppn=40
+#PBS -r n
+#PBS -N ulmame
+SPECIES=19049-ULM-AME
+
+module load gcc/4.9.2
+module load openmpi/1.8.3
+
+# written for version 1.2
+SRC=~/STModel-MCMC/bin
+DIR=~/STModel-Two-State/species/$SPECIES
+
+cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
+
+mkdir $DIR/res/mcmc
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+
+wait

### 2015-07-01 tex/tab_si1_col_pars.tex
Prepped relaunch scripts for 3 more species and updated some tables

\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{pdflscape}

% use proper unicode fonts
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\usepackage{amsmath} % for better display of equations

\usepackage{caption}
%\captionsetup{labelsep=none,textformat=empty} % for final
\captionsetup{labelformat=simple} % for drafts
\usepackage{booktabs}
\usepackage{multirow}
\usepackage[flushleft]{threeparttable}

\begin{document}

%\newgeometry{margin=1cm}
%\begin{landscape}

\newcommand{\ts}{\scriptsize}
\begin{table}[tb]
\small
\begin{threeparttable}
\label{tab:model_selection}
\caption{Parameterized colonization model selected for each species studied}
\begin{tabular}{lccccccc}
\toprule
\multirow{2}{*}{Scientific name} & \multirow{2}{2cm}{\centering Temperature variable\tnote{*}} & 
\multirow{2}{2cm}{\centering Precipitation variable\tnote{\textdagger}} & \multicolumn{5}{c}{Parameters\tnote{\textdaggerdbl}} \\
& & & $c_0$ & $c_1$ & $c_2$ & $c_3$ & $c_4$ \\ 
\midrule

\multirow{2}{*}{\it Abies balsamea} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx\tnote{\textsection} & x.xx & x.xx & x.xx & x.xx \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} \\

\multirow{2}{*}{\it Acer rubrum} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx &  & x.xx \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  & {\ts (x.xx, x.xx)} \\

\multirow{2}{*}{\it Acer saccharum} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx & x.xx &  \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  \\

\multirow{2}{*}{\it Betula alleghaniensis} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx & x.xx & \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & \\

\multirow{2}{*}{\it Betula papyrifera} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx & x.xx & x.xx \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} \\

\multirow{2}{*}{\it Fagus grandifolia} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx & x.xx &  \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  \\

\multirow{2}{*}{\it Fraxinus americana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx & x.xx &  \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  \\

\multirow{2}{*}{\it Fraxinus pennsylvanica} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx & x.xx & x.xx \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} \\

\multirow{2}{*}{\it Liriodendron tulipifera} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx & x.xx &  \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & \\

\multirow{2}{*}{\it Nyssa sylvatica} & \multirow{2}{*}{MTWQ} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx & x.xx & x.xx \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} \\

\multirow{2}{*}{\it Picea glauca} & \multirow{2}{*}{MDR} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx & x.xx & x.xx \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} \\

\multirow{2}{*}{\it Picea mariana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx &  &  \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  &  \\

\multirow{2}{*}{\it Pinus banksiana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx &  &  \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  &  \\

\multirow{2}{*}{\it Pinus strobus} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx &  &  \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  &  \\

\multirow{2}{*}{\it Pinus taeda} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx & x.xx &  \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & \\

\multirow{2}{*}{\it Populus tremuloides} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx & x.xx &  \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  \\

\multirow{2}{*}{\it Quercus alba} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx & x.xx &  \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  \\

\multirow{2}{*}{\it Quercus nigra} & \multirow{2}{*}{MDR} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx & &  \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  &  \\

\multirow{2}{*}{\it Quercus rubra} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx & x.xx & x.xx \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} \\

\multirow{2}{*}{\it Quercus velutina} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx & x.xx & x.xx \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} \\

\multirow{2}{*}{\it Ulmus americana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx &  & \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & & \\
\bottomrule
\end{tabular}
\begin{tablenotes}
\item [*] {\ts Temperature variable abbreviations: MAT: Mean annual temperature;
		MTWQ: Mean temperature wettest quarter; MDR: Mean diurnal range}
\item [\textdagger] {\ts Precipitation variable abbreviations: MAP: Mean annual precipitation; 
MPWQ: Mean precipitation, warmest quarter; PS: Precipitation seasonality}
\item [\textdaggerdbl] {\ts The parameters fit the following equation for $\gamma$, the 
		colonization rate: $log \left( \frac{\gamma}{1-\gamma} \right) = c_0 + c_1T + c_2P + c_3T^2 + c_4P^2$;
		where $T$ and $P$ are the selected temperature and precipitation variables, 
		respectively, and where a blank cell indicates a parameter that was not included 
		in the best-supported model}
\item [\textsection] {\ts Values are the median values from the posterior distribution with 5 and 95\% quantiles}
\end{tablenotes}
\end{threeparttable}
\end{table}

%\end{landscape}
%\restoregeometry

\end{document}

### 2015-07-01 tex/tab_si2_ext_pars.tex
Prepped relaunch scripts for 3 more species and updated some tables

--- 2015-07-01 tex/tab_si1_col_pars.tex
@@
 \label{tab:model_selection}
-\caption{Parameterized colonization model selected for each species studied}
+\caption{Parameterized extinction model selected for each species studied}
 \begin{tabular}{lccccccc}
@@
 \multirow{2}{2cm}{\centering Precipitation variable\tnote{\textdagger}} & \multicolumn{5}{c}{Parameters\tnote{\textdaggerdbl}} \\
-& & & $c_0$ & $c_1$ & $c_2$ & $c_3$ & $c_4$ \\ 
+& & & $e_0$ & $e_1$ & $e_2$ & $e_3$ & $e_4$ \\ 
 \midrule
@@
 
-\multirow{2}{*}{\it Acer rubrum} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx &  & x.xx \\
- &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  & {\ts (x.xx, x.xx)} \\
+\multirow{2}{*}{\it Acer rubrum} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx & x.xx & \\
+&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  \\
 
@@
 
-\multirow{2}{*}{\it Betula alleghaniensis} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx & x.xx & \\
- &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & \\
+\multirow{2}{*}{\it Betula alleghaniensis} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx & & \\
+&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  & \\
 
-\multirow{2}{*}{\it Betula papyrifera} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx & x.xx & x.xx \\
+\multirow{2}{*}{\it Betula papyrifera} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx &  &  \\
+&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  &  \\
+
+\multirow{2}{*}{\it Fagus grandifolia} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx & x.xx & x.xx \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} \\
 
-\multirow{2}{*}{\it Fagus grandifolia} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx & x.xx &  \\
- &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  \\
-
-\multirow{2}{*}{\it Fraxinus americana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx & x.xx &  \\
- &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  \\
+\multirow{2}{*}{\it Fraxinus americana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx & x.xx & x.xx \\
+&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)}  \\
 
@@
 
-\multirow{2}{*}{\it Nyssa sylvatica} & \multirow{2}{*}{MTWQ} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx & x.xx & x.xx \\
+\multirow{2}{*}{\it Nyssa sylvatica} & \multirow{2}{*}{MTWQ} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx &  & x.xx \\
+&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  & {\ts (x.xx, x.xx)} \\
+
+\multirow{2}{*}{\it Picea glauca} & \multirow{2}{*}{MDR} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx &  & x.xx \\
+&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  & {\ts (x.xx, x.xx)} \\
+
+\multirow{2}{*}{\it Picea mariana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx & x.xx &  \\
+&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  \\
+
+\multirow{2}{*}{\it Pinus banksiana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx & x.xx &  \\
+&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  \\
+
+\multirow{2}{*}{\it Pinus strobus} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx & x.xx & x.xx \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} \\
-
-\multirow{2}{*}{\it Picea glauca} & \multirow{2}{*}{MDR} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx & x.xx & x.xx \\
- &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} \\
-
-\multirow{2}{*}{\it Picea mariana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx &  &  \\
- &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  &  \\
-
-\multirow{2}{*}{\it Pinus banksiana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx &  &  \\
- &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  &  \\
-
-\multirow{2}{*}{\it Pinus strobus} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx &  &  \\
- &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  &  \\
 
@@
 
-\multirow{2}{*}{\it Populus tremuloides} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx & x.xx &  \\
+\multirow{2}{*}{\it Populus tremuloides} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx &  &  \\
+&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  &  \\
+
+\multirow{2}{*}{\it Quercus alba} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx & x.xx & x.xx \\
+&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} \\
+
+\multirow{2}{*}{\it Quercus nigra} & \multirow{2}{*}{MDR} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx & x.xx &  \\
 &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  \\
-
-\multirow{2}{*}{\it Quercus alba} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx & x.xx &  \\
- &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  \\
-
-\multirow{2}{*}{\it Quercus nigra} & \multirow{2}{*}{MDR} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx & &  \\
- &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  &  \\
 
@@
 
-\multirow{2}{*}{\it Quercus velutina} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx & x.xx & x.xx \\
- &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} \\
+\multirow{2}{*}{\it Quercus velutina} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx & x.xx &  \\
+&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  \\
 
-\multirow{2}{*}{\it Ulmus americana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx &  & \\
- &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & & \\
+\multirow{2}{*}{\it Ulmus americana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx & x.xx & \\
+&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & \\
 \bottomrule
@@
 MPWQ: Mean precipitation, warmest quarter; PS: Precipitation seasonality}
-\item [\textdaggerdbl] {\ts The parameters fit the following equation for $\gamma$, the 
-		colonization rate: $log \left( \frac{\gamma}{1-\gamma} \right) = c_0 + c_1T + c_2P + c_3T^2 + c_4P^2$;
+\item [\textdaggerdbl] {\ts The parameters fit the following equation for $\epsilon$, the 
+		extinction rate: $log \left( \frac{\epsilon}{1-\epsilon} \right) = e_0 + e_1T + e_2P + e_3T^2 + e_4P^2$;
 		where $T$ and $P$ are the selected temperature and precipitation variables, 

### 2015-07-01 scr/5_mcmc_scripts/fraame.sh

@@ -1,22 +1,0 @@
-#!/bin/sh
-#PBS -q default
-#PBS -l walltime=336:00:00
-#PBS -l nodes=3:ppn=40
-#PBS -r n
-#PBS -N fraame
-SPECIES=32931-FRA-AME
-
-module load gcc/4.9.2
-module load openmpi/1.8.3
-
-# written for version 1.2
-SRC=~/STModel-MCMC/bin
-DIR=~/STModel-Two-State/species/$SPECIES
-
-cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
-
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
-
-wait

### 2015-07-01 scr/5_mcmc_scripts/poptre.quealb.fraame.short.1node.sh

@@ -1,0 +1,31 @@
+#!/bin/sh
+#PBS -q default
+#PBS -l walltime=120:00:00
+#PBS -l nodes=3:ppn=40
+#PBS -r n
+#PBS -N poptre.quealb.fraame.1
+
+module load gcc/4.9.2
+module load openmpi/1.8.3
+
+# written for version 1.2
+SRC=~/STModel-MCMC/bin
+
+
+SPECIES=195773-POP-TRE
+DIR=~/STModel-Two-State/species/$SPECIES
+cd $DIR; mkdir -p res/mcmc1
+cd $DIR; $SRC/stm2_mcmc -r res/mcmc1/resumeData.txt -t dat/mcmc_data.txt -i 20000 2>mcmc_log1.txt &
+
+SPECIES=19290-QUE-ALB
+DIR=~/STModel-Two-State/species/$SPECIES
+cd $DIR; mkdir -p res/mcmc1
+cd $DIR; $SRC/stm2_mcmc -r res/mcmc1/resumeData.txt -t dat/mcmc_data.txt -i 20000 2>mcmc_log1.txt &
+
+SPECIES=32931-FRA-AME
+DIR=~/STModel-Two-State/species/$SPECIES
+cd $DIR; mkdir -p res/mcmc1
+cd $DIR; $SRC/stm2_mcmc -r res/mcmc1/resumeData.txt -t dat/mcmc_data.txt -i 20000 2>mcmc_log1.txt &
+
+
+wait

### 2015-07-01 scr/5_mcmc_scripts/poptre.sh

@@ -1,22 +1,0 @@
-#!/bin/sh
-#PBS -q default
-#PBS -l walltime=336:00:00
-#PBS -l nodes=3:ppn=40
-#PBS -r n
-#PBS -N poptre
-SPECIES=195773-POP-TRE
-
-module load gcc/4.9.2
-module load openmpi/1.8.3
-
-# written for version 1.2
-SRC=~/STModel-MCMC/bin
-DIR=~/STModel-Two-State/species/$SPECIES
-
-cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
-
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
-
-wait

### 2015-07-01 scr/5_mcmc_scripts/quealb.sh

@@ -1,22 +1,0 @@
-#!/bin/sh
-#PBS -q default
-#PBS -l walltime=336:00:00
-#PBS -l nodes=3:ppn=40
-#PBS -r n
-#PBS -N quealb
-SPECIES=19290-QUE-ALB
-
-module load gcc/4.9.2
-module load openmpi/1.8.3
-
-# written for version 1.2
-SRC=~/STModel-MCMC/bin
-DIR=~/STModel-Two-State/species/$SPECIES
-
-cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
-
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
-
-wait

### 2015-07-01 scr/post_mcmc_check.r

@@ -2,16 +2,43 @@ library(coda)
 
 # temporary
 setwd("~/Dropbox/work/projects/STModel-Two-State_git/")
-spName = '28731-ACE-SAC'
+spName = '32931-FRA-AME'
 
 p1.raw = read.csv(file.path('species', spName, 'res', 'mcmc1', 'posterior.csv'))
-p2.raw = read.csv(file.path('species', spName, 'res', 'mcmc2', 'posterior.csv'))
-p3.raw = read.csv(file.path('species', spName, 'res', 'mcmc3', 'posterior.csv'))
-
-p1 = mcmc(p1.raw)
-p2 = mcmc(p2.raw)
-p3 = mcmc(p3.raw)
-
-posterior = mcmc.list(p1, p2, p3)
+posterior = mcmc(p1.raw)
 
 plot(posterior, ask=T)
+summary(posterior)
+
+pquant = summary(posterior)$quantiles
+minpar = which(pquant[,1] == min(pquant[,1]))
+maxpar = which(pquant[,5] == max(pquant[,5]))
+
+minsamp = posterior[posterior[,minpar] == min(posterior[,minpar]),]
+maxsamp = posterior[posterior[,maxpar] == min(posterior[,maxpar]),]
+
+findsum = function(x, fun)
+{
+	if(length(dim(x)) == 2)
+		return(x[which(apply(x, 1, sum) == fun(apply(x, 1, sum))),])
+	if(is.vector(x))
+		return(x)
+	stop("Need a vector or matrix-like object")
+}
+	
+
+initMax = findsum(maxsamp, max)
+initMin = findsum(minsamp, min)
+plot(initMax, col='blue', ylim=c(min(initMin), max(initMax)), pch=16)
+points(initMin, col='red', pch=16)
+abline(h=0)
+
+i2path = file.path('species', spName, 'dat', 'mcmc_inits2.txt')
+i2 = read.csv(i2path)
+i2$initialValue = initMax
+write.csv(i2, i2path, row.names=FALSE)
+
+i2path = file.path('species', spName, 'dat', 'mcmc_inits3.txt')
+i2 = read.csv(i2path)
+i2$initialValue = initMin
+write.csv(i2, i2path, row.names=FALSE)

### 2015-07-01 scr/5_mcmc_scripts/acerub.sh
Updated remaining mcmc scripts to only start the first chain (for getting better starts)

@@ -1,7 +1,7 @@
 #!/bin/sh
 #PBS -q default
-#PBS -l walltime=336:00:00
-#PBS -l nodes=3:ppn=40
+#PBS -l walltime=24:00:00
+#PBS -l nodes=1:ppn=40
 #PBS -r n
 #PBS -N acerub
 SPECIES=28728-ACE-RUB
@@ -16,7 +16,7 @@ DIR=~/STModel-Two-State/species/$SPECIES
 cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
 
 cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+#cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+#cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
 
 wait

### 2015-07-01 scr/5_mcmc_scripts/betall.sh

@@ -1,7 +1,7 @@
 #!/bin/sh
 #PBS -q default
-#PBS -l walltime=336:00:00
-#PBS -l nodes=3:ppn=40
+#PBS -l walltime=24:00:00
+#PBS -l nodes=1:ppn=40
 #PBS -r n
 #PBS -N betall
 SPECIES=19481-BET-ALL
@@ -16,7 +16,7 @@ DIR=~/STModel-Two-State/species/$SPECIES
 cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
 
 cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+#cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+#cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
 
 wait

### 2015-07-01 scr/5_mcmc_scripts/betpap.sh

@@ -1,7 +1,7 @@
 #!/bin/sh
 #PBS -q default
-#PBS -l walltime=336:00:00
-#PBS -l nodes=3:ppn=40
+#PBS -l walltime=24:00:00
+#PBS -l nodes=1:ppn=40
 #PBS -r n
 #PBS -N betpap
 SPECIES=19489-BET-PAP
@@ -16,7 +16,7 @@ DIR=~/STModel-Two-State/species/$SPECIES
 cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
 
 cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+#cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+#cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
 
 wait

### 2015-07-01 scr/5_mcmc_scripts/faggra.sh

@@ -1,7 +1,7 @@
 #!/bin/sh
 #PBS -q default
-#PBS -l walltime=336:00:00
-#PBS -l nodes=3:ppn=40
+#PBS -l walltime=24:00:00
+#PBS -l nodes=1:ppn=40
 #PBS -r n
 #PBS -N faggra
 SPECIES=19462-FAG-GRA
@@ -16,7 +16,7 @@ DIR=~/STModel-Two-State/species/$SPECIES
 cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
 
 cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+#cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+#cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
 
 wait

### 2015-07-01 scr/5_mcmc_scripts/frapen.sh

@@ -1,7 +1,7 @@
 #!/bin/sh
 #PBS -q default
-#PBS -l walltime=336:00:00
-#PBS -l nodes=3:ppn=40
+#PBS -l walltime=24:00:00
+#PBS -l nodes=1:ppn=40
 #PBS -r n
 #PBS -N frapen
 SPECIES=32929-FRA-PEN
@@ -16,7 +16,7 @@ DIR=~/STModel-Two-State/species/$SPECIES
 cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
 
 cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+#cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+#cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
 
 wait

### 2015-07-01 scr/5_mcmc_scripts/lirtul.sh

@@ -1,7 +1,7 @@
 #!/bin/sh
 #PBS -q default
-#PBS -l walltime=336:00:00
-#PBS -l nodes=3:ppn=40
+#PBS -l walltime=24:00:00
+#PBS -l nodes=1:ppn=40
 #PBS -r n
 #PBS -N lirtul
 SPECIES=18086-LIR-TUL
@@ -16,7 +16,7 @@ DIR=~/STModel-Two-State/species/$SPECIES
 cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
 
 cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+#cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+#cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
 
 wait

### 2015-07-01 scr/5_mcmc_scripts/nyssyl.sh

@@ -1,7 +1,7 @@
 #!/bin/sh
 #PBS -q default
-#PBS -l walltime=336:00:00
-#PBS -l nodes=3:ppn=40
+#PBS -l walltime=24:00:00
+#PBS -l nodes=1:ppn=40
 #PBS -r n
 #PBS -N nyssyl
 SPECIES=27821-NYS-SYL
@@ -16,7 +16,7 @@ DIR=~/STModel-Two-State/species/$SPECIES
 cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
 
 cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+#cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+#cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
 
 wait

### 2015-07-01 scr/5_mcmc_scripts/picgla.sh

@@ -1,7 +1,7 @@
 #!/bin/sh
 #PBS -q default
-#PBS -l walltime=336:00:00
-#PBS -l nodes=3:ppn=40
+#PBS -l walltime=24:00:00
+#PBS -l nodes=1:ppn=40
 #PBS -r n
 #PBS -N picgla
 SPECIES=183295-PIC-GLA
@@ -16,7 +16,7 @@ DIR=~/STModel-Two-State/species/$SPECIES
 cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
 
 cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+#cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+#cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
 
 wait

### 2015-07-01 scr/5_mcmc_scripts/picmar.sh

@@ -1,7 +1,7 @@
 #!/bin/sh
 #PBS -q default
-#PBS -l walltime=336:00:00
-#PBS -l nodes=3:ppn=40
+#PBS -l walltime=24:00:00
+#PBS -l nodes=1:ppn=40
 #PBS -r n
 #PBS -N picmar
 SPECIES=183302-PIC-MAR
@@ -16,7 +16,7 @@ DIR=~/STModel-Two-State/species/$SPECIES
 cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
 
 cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+#cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+#cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
 
 wait

### 2015-07-01 scr/5_mcmc_scripts/pinban.sh

@@ -1,7 +1,7 @@
 #!/bin/sh
 #PBS -q default
-#PBS -l walltime=336:00:00
-#PBS -l nodes=3:ppn=40
+#PBS -l walltime=24:00:00
+#PBS -l nodes=1:ppn=40
 #PBS -r n
 #PBS -N pinban
 SPECIES=183319-PIN-BAN
@@ -16,7 +16,7 @@ DIR=~/STModel-Two-State/species/$SPECIES
 cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
 
 cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+#cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+#cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
 
 wait

### 2015-07-01 scr/5_mcmc_scripts/pinstr.sh

@@ -1,7 +1,7 @@
 #!/bin/sh
 #PBS -q default
-#PBS -l walltime=336:00:00
-#PBS -l nodes=3:ppn=40
+#PBS -l walltime=24:00:00
+#PBS -l nodes=1:ppn=40
 #PBS -r n
 #PBS -N pinstr
 SPECIES=183385-PIN-STR
@@ -16,7 +16,7 @@ DIR=~/STModel-Two-State/species/$SPECIES
 cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
 
 cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+#cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+#cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
 
 wait

### 2015-07-01 scr/5_mcmc_scripts/pintae.sh

@@ -1,7 +1,7 @@
 #!/bin/sh
 #PBS -q default
-#PBS -l walltime=336:00:00
-#PBS -l nodes=3:ppn=40
+#PBS -l walltime=24:00:00
+#PBS -l nodes=1:ppn=40
 #PBS -r n
 #PBS -N pintae
 SPECIES=18037-PIN-TAE
@@ -16,7 +16,7 @@ DIR=~/STModel-Two-State/species/$SPECIES
 cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
 
 cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+#cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+#cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
 
 wait

### 2015-07-01 scr/5_mcmc_scripts/quenig.sh

@@ -1,7 +1,7 @@
 #!/bin/sh
 #PBS -q default
-#PBS -l walltime=336:00:00
-#PBS -l nodes=3:ppn=40
+#PBS -l walltime=48:00:00
+#PBS -l nodes=1:ppn=40
 #PBS -r n
 #PBS -N quenig
 SPECIES=19280-QUE-NIG
@@ -16,7 +16,7 @@ DIR=~/STModel-Two-State/species/$SPECIES
 cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
 
 cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+#cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+#cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
 
 wait

### 2015-07-01 scr/5_mcmc_scripts/querub.sh

@@ -1,7 +1,7 @@
 #!/bin/sh
 #PBS -q default
-#PBS -l walltime=336:00:00
-#PBS -l nodes=3:ppn=40
+#PBS -l walltime=48:00:00
+#PBS -l nodes=1:ppn=40
 #PBS -r n
 #PBS -N querub
 SPECIES=19408-QUE-RUB
@@ -16,7 +16,7 @@ DIR=~/STModel-Two-State/species/$SPECIES
 cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
 
 cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+#cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+#cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
 
 wait

### 2015-07-01 scr/5_mcmc_scripts/quevel.sh

@@ -1,7 +1,7 @@
 #!/bin/sh
 #PBS -q default
-#PBS -l walltime=336:00:00
-#PBS -l nodes=3:ppn=40
+#PBS -l walltime=48:00:00
+#PBS -l nodes=1:ppn=40
 #PBS -r n
 #PBS -N quevel
 SPECIES=19447-QUE-VEL
@@ -16,7 +16,7 @@ DIR=~/STModel-Two-State/species/$SPECIES
 cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
 
 cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+#cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+#cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
 
 wait

### 2015-07-01 scr/5_mcmc_scripts/ulmame.sh

@@ -1,7 +1,7 @@
 #!/bin/sh
 #PBS -q default
-#PBS -l walltime=336:00:00
-#PBS -l nodes=3:ppn=40
+#PBS -l walltime=48:00:00
+#PBS -l nodes=1:ppn=40
 #PBS -r n
 #PBS -N ulmame
 SPECIES=19049-ULM-AME
@@ -15,8 +15,8 @@ DIR=~/STModel-Two-State/species/$SPECIES
 
 cd $DIR; mkdir -p res/mcmc1 res/mcmc2 res/mcmc3
 
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
-cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
+cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits1.txt -t dat/mcmc_data.txt -o res/mcmc1 -n 1 -b 0 -i 10000 -c 40 -l 5 -v 2 2>mcmc_log1.txt &
+# #cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits2.txt -t dat/mcmc_data.txt -o res/mcmc2 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log2.txt &
+# #cd $DIR; $SRC/stm2_mcmc -p dat/mcmc_inits3.txt -t dat/mcmc_data.txt -o res/mcmc3 -n 1 -b 0 -i 20000 -c 40 -l 5 -v 2 2>mcmc_log3.txt &
 
 wait

### 2015-07-03 tex/manuscript.tex
Added ms support files and a bibliography

@@
 \usepackage{setspace}
-\usepackage{natbib}
+%\usepackage{natbib}
 
@@
 
+Fisichelli and colleagues reported recruitement of temperate trees into boreal ecosystems, supporting our conclusion that colonization of temperature trees varies as a function of temperature \cite{Fisichelli2014}.
+
 This model enables us to more precisely define the niche in terms of the fundamental processes of colonization and extinction (fig..). 
@@
 % this deserves a paragraph here
+Our model is particularly appropriate for forest trees, which are sessile, disperse relatively short distances, and are long-lived. 
+Unique challenges confront applying such a model without these characteristics, such as most animals. 
+Animals in highly patchy environments, with local populations linked by moderate dispersal rates, are likely to be the most appropriate species for use with our model.
 
@@
 
+\renewcommand\refname{Literature Cited}
+\bibliography{2state}{}
+\bibliographystyle{pnas}
+
 \end{document}

### 2015-07-03 tex/tab_si1_col_pars.tex
Added ms support files and a bibliography

@@
 
-\usepackage[margin=1in]{geometry}
+\usepackage[margin=0.5in]{geometry}
 \usepackage{pdflscape}
@@
 
-\multirow{2}{*}{\it Abies balsamea} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx\tnote{\textsection} & x.xx & x.xx & x.xx & x.xx \\
- &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} \\
+\multirow{2}{*}{\it Abies balsamea} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & -2.53\tnote{\textsection} & -2.46 & 0.07 & -1.26 & 0.03 \\
+ &  &  & {\ts (-2.86, -2.18)} & {\ts (-3.08, -1.81)} & {\ts (-0.01, 0.16)} & {\ts (-1.55, -0.98)} & {\ts (-0.03, 0.09)} \\
 
@@
 
-\multirow{2}{*}{\it Acer saccharum} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx & x.xx &  \\
- &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  \\
+\multirow{2}{*}{\it Acer saccharum} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & -2.10 & 0.47 & 0.09 & -0.34 &  \\
+ &  &  & {\ts (-2.22, -1.99)} & {\ts (0.34, 0.59)} & {\ts (0.01, 0.17)} & {\ts (-0.55, -0.12)} &  \\
 
@@
 
-\multirow{2}{*}{\it Populus tremuloides} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx & x.xx &  \\
- &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  \\
+\multirow{2}{*}{\it Populus tremuloides} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & -2.15 & -0.75 & -0.11 & -0.86 &  \\
+ &  &  & {\ts (-2.44, -1.86)} & {\ts (-1.48, -0.05)} & {\ts (-0.20, -0.03)} & {\ts (-1.32, -0.43)} &  \\
 
@@
 		respectively, and where a blank cell indicates a parameter that was not included 
-		in the best-supported model}
-\item [\textsection] {\ts Values are the median values from the posterior distribution with 5 and 95\% quantiles}
+		in the best-supported model. 
+		All parameters were fit to data scaled to 0 mean and unit variance.}
+\item [\textsection] {\ts Values are the mean values from the posterior distribution with 5 and 95\% quantiles}
 \end{tablenotes}

### 2015-07-03 tex/tab_si2_ext_pars.tex
Added ms support files and a bibliography

@@
 
-\multirow{2}{*}{\it Abies balsamea} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx\tnote{\textsection} & x.xx & x.xx & x.xx & x.xx \\
-&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} \\
+\multirow{2}{*}{\it Abies balsamea} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & -3.21\tnote{\textsection} & -1.55 & -0.45 & -0.98 & -0.13 \\
+&  &  & {\ts (-3.56, -2.88)} & {\ts (-2.18, -0.94)} & {\ts (-0.54, -0.37)} & {\ts (-1.26, -0.70)} & {\ts (-0.19, -0.07)} \\
 
@@
 
-\multirow{2}{*}{\it Acer saccharum} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx & x.xx &  \\
-&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  \\
+\multirow{2}{*}{\it Acer saccharum} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & -3.18 & 0.73 & 0.47 & 0.32 &  \\
+&  &  & {\ts (-3.32, -3.04)} & {\ts (0.61, 0.85)} & {\ts (0.39, 0.55)} & {\ts (0.08, 0.56)} &  \\
 
@@
 
-\multirow{2}{*}{\it Populus tremuloides} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx &  &  \\
-&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  &  \\
+\multirow{2}{*}{\it Populus tremuloides} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & -1.40 & 1.20 & -0.06 &  &  \\
+&  &  & {\ts (-1.56, -1.25)} & {\ts (1.02, 1.36)} & {\ts (-0.13, 0.01)} &  &  \\
 
@@
 		respectively, and where a blank cell indicates a parameter that was not included 
-		in the best-supported model}
-\item [\textsection] {\ts Values are the median values from the posterior distribution with 5 and 95\% quantiles}
+		in the best-supported model.
+		All parameters were fit to data scaled to 0 mean and unit variance.}
+\item [\textsection] {\ts Values are the mean values from the posterior distribution with 5 and 95\% quantiles}
 \end{tablenotes}

### 2015-07-03 tex/table.tex
Added ms support files and a bibliography

@@
 \end{table}
-
-\begin{table}[tb]
-\label{tab:parameterized_equations}
-\caption{Parameterized equations for colonization and extinction rate}
-\begin{tabular}{lll}
-\toprule
-Species                       & Colonization                               & Extinction                                  \\ \midrule
-{\it Abies balsamea}          & $x.xx + x.xxT + x.xxT^2 + x.xxP + x.xxP^2$ & $x.xx + x.xxT + x.xxT^2 + x.xxP + x.xxP^2$  \\
-{\it Acer rubrum}             & $x.xx + x.xxT + x.xxP + x.xxP^2$           & $x.xx + x.xxT + x.xxT^2 + x.xxP$            \\
-{\it Acer saccharum}          & $x.xx + x.xxT + x.xxT^2 + x.xxP$           & $x.xx + x.xxT + x.xxT^2 + x.xxP$            \\
-{\it Betula alleghaniensis}   & $x.xx + x.xxT + x.xxT^2 + x.xxP$           & $x.xx + x.xxT + x.xxP$                      \\
-{\it Betula papyrifera}       & $x.xx + x.xxT + x.xxT^2 + x.xxP + x.xxP^2$ & $x.xx + x.xxT + x.xxP$                      \\
-{\it Fagus grandifolia}       & $x.xx + x.xxT + x.xxT^2 + x.xxP$           & $x.xx + x.xxT + x.xxT^2 + x.xxP + x.xxP^2$  \\
-{\it Fraxinus americana}      & $x.xx + x.xxT + x.xxT^2 + x.xxP$           & $x.xx + x.xxT + x.xxT^2 + x.xxP + x.xxP^2$  \\
-{\it Fraxinus pennsylvanica}  & $x.xx + x.xxT + x.xxT^2 + x.xxP + x.xxP^2$ & $x.xx + x.xxT + x.xxT^2 + x.xxP + x.xxP^2$  \\
-{\it Liriodendron tulipifera} & $x.xx + x.xxT + x.xxT^2 + x.xxP$           & $x.xx + x.xxT + x.xxT^2 + x.xxP$            \\
-{\it Nyssa sylvatica}         & $x.xx + x.xxT + x.xxT^2 + x.xxP + x.xxP^2$ & $x.xx + x.xxT + x.xxP + x.xxP^2$            \\
-{\it Picea glauca}            & $x.xx + x.xxT + x.xxT^2 + x.xxP + x.xxP^2$ & $x.xx + x.xxT + x.xxP + x.xxP^2$            \\
-{\it Picea mariana}           & $x.xx + x.xxT + x.xxP$                     & $x.xx + x.xxT + x.xxT^2 + x.xxP$            \\
-{\it Pinus banksiana}         & $x.xx + x.xxT + x.xxP$                     & $x.xx + x.xxT + x.xxT^2 + x.xxP$            \\
-{\it Pinus strobus}           & $x.xx + x.xxT + x.xxP$                     & $x.xx + x.xxT + x.xxT^2 + x.xxP + x.xxP^2$  \\
-{\it Pinus taeda}             & $x.xx + x.xxT + x.xxT^2 + x.xxP$           & $x.xx + x.xxT + x.xxT^2 + x.xxP$            \\
-{\it Populus tremuloides}     & $x.xx + x.xxT + x.xxT^2 + x.xxP$           & $x.xx + x.xxT + x.xxP$                      \\ 
-{\it Quercus alba}            & $x.xx + x.xxT + x.xxT^2 + x.xxP$           & $x.xx + x.xxT + x.xxT^2 + x.xxP$            \\
-{\it Quercus nigra}           & $x.xx + x.xxT + x.xxP$                     & $x.xx + x.xxT + x.xxT^2 + x.xxP$            \\
-{\it Quercus rubra}           & $x.xx + x.xxT + x.xxT^2 + x.xxP + x.xxP^2$ & $x.xx + x.xxT + x.xxT^2 + x.xxP + x.xxP^2$  \\
-{\it Quercus velutina}        & $x.xx + x.xxT + x.xxT^2 + x.xxP + x.xxP^2$ & $x.xx + x.xxT + x.xxT^2 + x.xxP$            \\
-{\it Ulmus americana}         & $x.xx + x.xxT + x.xxP$                     & $x.xx + x.xxT + x.xxT^2 + x.xxP$            \\ 
-\bottomrule
-\end{tabular}
-\end{table}

### 2015-07-08 tex/manuscript.tex
Updated resume script for pintae

@@
 
+We built a dynamic state-and-transition model (STM) using classic metapopulation theory \citep{Levins1969, Nee1992} and the foundation.
+The classical formulation of this model is inappropriate for modeling range dynamics at large scales, as it assumes colonization and extinction rates are constant in space and time and treats prevalence as a global variable.
+Holt and Keitt extended the Levins model to a scale appropriate for range dynamics by varying the parameters along an environmental gradient, showing that, under metapopulation dynamics, range limits of a theoretical species arise when an environmental gradient causes an increase in the extinction rate, a decrease in the colonization rate, or a decrease in available habitat \cite{Holt2000}.
+
+We applied a model similar to the Holt model \cite{Holt2000} to 21 forest tree species that are abundant in eastern North American temperate and boreal forests.
+Our model, which is discrete in space and time, treats the landscape as a series of patches, where the state $Y_{x,t}$ of patch $x$ at time $t$ is a function of the colonization ($c_{x,t}$) and extinction ($e_{x,t}$) rates, as well as the prevalence of the species ($phi_{x,t}$).
+Both the colonization and extinction rates were parameterized as a function of two environmental variables, one related to temperature and one precipitation, thus $c_x = f(T_x, P_x); e_x = g(T_X, P_x)$ (see \emph{Methods} for descriptions of the specific variables used).
+As our objectives concern equilibrial dynamics and the role of CE dynamics in determining species ranges, for the present we ignore temporal dynamics and focus instead on spatial variation in the model parameters.
+
+Our model is a first order Markov chain that is dynamic in space and time (though we fix the time dimension for this paper).
+The state of a patch depends on the state of local patches at the previous time step and the transition rates.
+Range boundaries emerge at equilibrium due to the effect of environmental gradients on the transition rates
+
+We apply our model exclusively to forest trees, and because our model is parameterized using forest plot data, we treat all patches as potential habitat, thus setting $k=1$ from the Levins model.
+Following this simplification, the equilibrium condition for the presence of a species on the environmental gradient is any point $x$ such that $c_x > e_x$ \citep{Holt2000}.
+It follows that the range limit of the species will occur where $c_x = e_x$.
+
 We designed a colonization-extinction model with a dynamic parameterization that allowed the colonization and extinction rates to vary as a function of climate.
@@
 The consequence of these two processes, when overlaid, is a range covering intermediate temperatures, bounded by low colonization in the north and high extinction in the south (fig - map).
+A further advantage of the model, particularly when compared to static correlative SDMs, is the ability to investigate the resilience of the system at the range margin.
+At the range boundary, the dominant eigenvalue, representing the response of the system to small perturbations (i.e., recovery after disturbance or the acquisition of the new equilibrium following climate change) approaches zero \cite{Scheffer2009}.
+The rate at which this value approaches zero (i.e., the relative slopes of $c$ and $e$ along the environmental gradient -- ref response curves) indicates the degree to which the range margin will respond to changes in the gradient.
+Thus, "fast" species with steep slopes along the gradient (e.g., list some species) will respond quickly to climate change, while "slow" species may lag behind climate change, producing instability and potentially promoting catastrophic changes (e.g., widespread and rapid local extinctions along the range margin) \cite{Scheffer2009, Scheffer2012}.
+%% NOTE - this may be a paper in itself - worth considering.
+It could be cool to actually measure this and then investigate clim change scenarios via simulating this model.
+could easily measure both the variance and the autocorrelation in such sims
 
@@
 Second, the use of a Bayesian framework allows the inclusion of prior information.
-As our parameters were all estimated on the logit scale, we used the weakly informative Cauchy priors suggested by (cite Gelman paper), with a mean of 0 and (variance?) of 10 for intercept parameters or 2.5 for slope parameters.
+As our parameters were all estimated on the logit scale, we used the weakly informative Cauchy priors suggested by Gelman and colleagues \cite{Gelman2008}, with a center of 0 and scale of 10 for intercept parameters or 2.5 for slope parameters.
 Finally, having posterior distributions of the predictors allows for the information produced by our model to be incorporated in future meta-modeling efforts that can further improve our characterization of species' ranges and our understanding of model uncertainty (cite GEB paper).
@@
 For each species, we obtained xxx posterior samples...
-We assessed convergence...
+
+We assessed convergence using the multivariate potential scale reduction factor (MPSR) \cite{Brooks1998}.
+This method compares the variance within and among multiple parallel MCMC chains, where a value of 1 indicates perfect convergence (i.e., the chains are from indistinguishable distributions), and values approximately \textgreater1.1 indicate a lack of convergence.
+
+For climate variables, we used regional climate models at xx resolution \cite{McKenney2011}.
 

### 2015-07-08 scr/5_mcmc_scripts/pintae1.sh

@@ -14,6 +14,6 @@ SRC=~/STModel-MCMC/bin
 
 SPECIES=18037-PIN-TAE
 DIR=~/STModel-Two-State/species/$SPECIES
-cd $DIR; $SRC/stm2_mcmc -r res/mcmc1/resumeData.txt -t dat/mcmc_data.txt -i 20000 2>mcmc_log1.txt &
+cd $DIR; $SRC/stm2_mcmc -r res/mcmc1/resumeData.txt -t dat/mcmc_data.txt -i 40000 2>mcmc_log1.txt &
 
 wait

### 2015-07-08 scr/5_mcmc_scripts/pintae2.sh

@@ -14,6 +14,6 @@ SRC=~/STModel-MCMC/bin
 
 SPECIES=18037-PIN-TAE
 DIR=~/STModel-Two-State/species/$SPECIES
-cd $DIR; $SRC/stm2_mcmc -r res/mcmc2/resumeData.txt -t dat/mcmc_data.txt -i 20000 2>mcmc_log2.txt &
+cd $DIR; $SRC/stm2_mcmc -r res/mcmc2/resumeData.txt -t dat/mcmc_data.txt -i 40000 2>mcmc_log2.txt &
 
 wait

### 2015-07-08 scr/5_mcmc_scripts/pintae3.sh

@@ -14,6 +14,6 @@ SRC=~/STModel-MCMC/bin
 
 SPECIES=18037-PIN-TAE
 DIR=~/STModel-Two-State/species/$SPECIES
-cd $DIR; $SRC/stm2_mcmc -r res/mcmc3/resumeData.txt -t dat/mcmc_data.txt -i 20000 2>mcmc_log3.txt &
+cd $DIR; $SRC/stm2_mcmc -r res/mcmc3/resumeData.txt -t dat/mcmc_data.txt -i 40000 2>mcmc_log3.txt &
 
 wait

### 2015-07-08 scr/6_post_mcmc_check.r

@@ -2,9 +2,9 @@ library(coda)
 
 # temporary
 setwd("~/Dropbox/work/projects/STModel-Two-State_git/")
-spName = '19489-BET-PAP'
+spName = '18037-PIN-TAE'
 
-design = '11111001110000'
+design = '11110001111000'
 design = sapply(1:nchar(design), function(i) as.integer(substr(design, i, i)))
 constCols = which(design == 0)
 

### 2015-07-09 tex/manuscript.tex
Lots of updates for new results

@@
 An alternative more mechanistic approach is to examine the colonization rate $c$ of unoccupied patches and the extinction rate $e$ of occupied patches.
+Theoretical studies have suggested that range limits can arise due to spatial variation in these two rates \cite{Holt2000, Holt2005}.
 (here maybe briefly review the literature on the subject)
@@
 
-We built a dynamic state-and-transition model (STM) using classic metapopulation theory \citep{Levins1969, Nee1992} and the foundation.
+We built a dynamic state-and-transition model (STM) using classic metapopulation theory \citep{Levins1969} as the foundation.
 The classical formulation of this model is inappropriate for modeling range dynamics at large scales, as it assumes colonization and extinction rates are constant in space and time and treats prevalence as a global variable.

### 2015-07-09 scr/5_mcmc_scripts/nyssyl2.sh

@@ -14,6 +14,6 @@ SRC=~/STModel-MCMC/bin
 
 SPECIES=27821-NYS-SYL
 DIR=~/STModel-Two-State/species/$SPECIES
-cd $DIR; $SRC/stm2_mcmc -r res/mcmc2/resumeData.txt -t dat/mcmc_data.txt -i 40000 2>mcmc_log2.txt &
+cd $DIR; $SRC/stm2_mcmc -r res/mcmc2/resumeData.txt -t dat/mcmc_data.txt -i 17000 2>mcmc_log2.txt &
 
 wait

### 2015-07-09 scr/5_mcmc_scripts/quenig2.sh

@@ -14,6 +14,6 @@ SRC=~/STModel-MCMC/bin
 
 SPECIES=19280-QUE-NIG
 DIR=~/STModel-Two-State/species/$SPECIES
-cd $DIR; $SRC/stm2_mcmc -r res/mcmc2/resumeData.txt -t dat/mcmc_data.txt -i 40000 2>mcmc_log2.txt &
+cd $DIR; $SRC/stm2_mcmc -r res/mcmc2/resumeData.txt -t dat/mcmc_data.txt -i 14000 2>mcmc_log2.txt &
 
 wait

### 2015-07-09 scr/5_mcmc_scripts/quenig3.sh

@@ -14,6 +14,6 @@ SRC=~/STModel-MCMC/bin
 
 SPECIES=19280-QUE-NIG
 DIR=~/STModel-Two-State/species/$SPECIES
-cd $DIR; $SRC/stm2_mcmc -r res/mcmc3/resumeData.txt -t dat/mcmc_data.txt -i 40000 2>mcmc_log3.txt &
+cd $DIR; $SRC/stm2_mcmc -r res/mcmc3/resumeData.txt -t dat/mcmc_data.txt -i 19000 2>mcmc_log3.txt &
 
 wait

### 2015-07-09 scr/5_mcmc_scripts/querub2.sh

@@ -14,6 +14,6 @@ SRC=~/STModel-MCMC/bin
 
 SPECIES=19408-QUE-RUB
 DIR=~/STModel-Two-State/species/$SPECIES
-cd $DIR; $SRC/stm2_mcmc -r res/mcmc2/resumeData.txt -t dat/mcmc_data.txt -i 20000 2>mcmc_log2.txt &
+cd $DIR; $SRC/stm2_mcmc -r res/mcmc2/resumeData.txt -t dat/mcmc_data.txt -i 14000 2>mcmc_log2.txt &
 
 wait

### 2015-07-09 scr/6_post_mcmc_check.r

@@ -1,17 +1,27 @@
 library(coda)
 
-# temporary
 setwd("~/Dropbox/work/projects/STModel-Two-State_git/")
-spName = '18037-PIN-TAE'
+spName = '19481-BET-ALL'
 
-design = '11110001111000'
+design = '11110001110000'
 design = sapply(1:nchar(design), function(i) as.integer(substr(design, i, i)))
 constCols = which(design == 0)
 
+st1 = st2 = st3 = 1
+# special instructions for individual species due to poor starts
+if(spName == '19408-QUE-RUB')
+	st2 = 14000
+if(spName == '27821-NYS-SYL')
+	st2 = 17000
+if(spName == '19280-QUE-NIG')
+	st2 = 19000
+
 p1 = read.csv(file.path('species', spName, 'res', 'mcmc1', 'posterior.csv'))[,-constCols]
 p2 = read.csv(file.path('species', spName, 'res', 'mcmc2', 'posterior.csv'))[,-constCols]
+p2 = p2[st2:nrow(p2),]
 p3 = read.csv(file.path('species', spName, 'res', 'mcmc3', 'posterior.csv'))[,-constCols]
 posterior = mcmc.list(mcmc(p1), mcmc(p2), mcmc(p3))
+lapply(posterior, nrow)
 
 plot(posterior, ask=T)
 summary(posterior)

### 2015-07-09 scr/fig/fig1_response_curves.r

@@ -1,8 +1,9 @@
 #!/usr/bin/Rscript
 varNames = readRDS("dat/climVariableNames.rds")
-spNames = c('18032-ABI-BAL', '28728-ACE-RUB', '28731-ACE-SAC', '19462-FAG-GRA', 
-		'32931-FRA-AME', '183295-PIC-GLA', '183385-PIN-STR', '195773-POP-TRE', 
-		'19290-QUE-ALB', '19447-QUE-VEL')
+spNames = c('18032-ABI-BAL', '28728-ACE-RUB', '28731-ACE-SAC', '19489-BET-PAP',
+		'19462-FAG-GRA', '32931-FRA-AME', '32929-FRA-PEN', '183295-PIC-GLA', 
+		'183319-PIN-BAN', '183385-PIN-STR', '195773-POP-TRE', '19290-QUE-ALB', 
+		'19447-QUE-VEL', '19049-ULM-AME')
 spInfoAll = read.csv('dat/speciesInfo.csv', stringsAsFactors=FALSE, colClasses='character')
 
 pdf(w=6.5, h=8, file="img/response_curves.pdf")

### 2015-07-09 scr/fig/fig2_posterior_maps.r

@@ -4,9 +4,10 @@ library(sp)
 library(raster)
 library(rgdal)
 
-spNames = c('18032-ABI-BAL', '28728-ACE-RUB', '28731-ACE-SAC', '19462-FAG-GRA', 
-		'32931-FRA-AME', '183295-PIC-GLA', '183385-PIN-STR', '195773-POP-TRE', 
-		'19290-QUE-ALB', '19447-QUE-VEL')
+spNames = c('18032-ABI-BAL', '28728-ACE-RUB', '28731-ACE-SAC', '19489-BET-PAP',
+		'19462-FAG-GRA', '32931-FRA-AME', '32929-FRA-PEN', '183295-PIC-GLA', 
+		'183319-PIN-BAN', '183385-PIN-STR', '195773-POP-TRE', '19290-QUE-ALB', 
+		'19447-QUE-VEL', '19049-ULM-AME')
 spInfoAll = read.csv('dat/speciesInfo.csv', stringsAsFactors=FALSE, colClasses='character')
 
 load('dat/map_projections.rdata')

### 2015-07-10 tex/manuscript.tex
ms changes and resume scri[ts for pintae

@@
 Species distribution models (SDMs) are commonly-applied to understand the latter two options.
-Classical SDMs often use a bioclimatic modeling approach that statistically relates presence or absence of the species to climate.
+Classical SDMs often use a bioclimatic modeling approach that statistically relates presence or absence of the species to climate \cite{Guisan2005}.
 Such models are quite powerful when describing the present realized niche of a species, however they suffer from several drawbacks when predicting the future distribution of species, and they provide no insight into the mechanisms underlying a species' response to the environment.
-An alternative more mechanistic approach is to examine the colonization rate $c$ of unoccupied patches and the extinction rate $e$ of occupied patches.
+An alternative more mechanistic approach is to examine the colonization rate $c$ of unoccupied patches and the extinction rate $e$ of occupied patches \cite{Yackulic2015}.
 Theoretical studies have suggested that range limits can arise due to spatial variation in these two rates \cite{Holt2000, Holt2005}.
@@
 
-We built a dynamic state-and-transition model (STM) using classic metapopulation theory \citep{Levins1969} as the foundation.
+We built a dynamic state-and-transition model (STM) using classic metapopulation theory \cite{Levins1969} as the foundation.
 The classical formulation of this model is inappropriate for modeling range dynamics at large scales, as it assumes colonization and extinction rates are constant in space and time and treats prevalence as a global variable.
 Holt and Keitt extended the Levins model to a scale appropriate for range dynamics by varying the parameters along an environmental gradient, showing that, under metapopulation dynamics, range limits of a theoretical species arise when an environmental gradient causes an increase in the extinction rate, a decrease in the colonization rate, or a decrease in available habitat \cite{Holt2000}.
+Similar models have been applied to investigate whether species distributions are at equilibrium with climate \cite{Garcia2013}.
 
@@
 We apply our model exclusively to forest trees, and because our model is parameterized using forest plot data, we treat all patches as potential habitat, thus setting $k=1$ from the Levins model.
-Following this simplification, the equilibrium condition for the presence of a species on the environmental gradient is any point $x$ such that $c_x > e_x$ \citep{Holt2000}.
+Following this simplification, the equilibrium condition for the presence of a species on the environmental gradient is any point $x$ such that $c_x > e_x$ \cite{Holt2000}.
 It follows that the range limit of the species will occur where $c_x = e_x$.
@@
 
+Non-equilibrium dynamics are likely to become increasingly important in understanding the response of many species to global change \cite{Araujo2005, Svenning2013}, and many systems will likely experience significant time lags \cite{Garcia2013, Wu2015}.
+Thus, it is crucial to develop models capable of relaxing the equilibrium assumption. 
+Models of colonization and extinction allow species distribution to emerge from the underlying processes, thereby relaxing the equilibrium assumption \cite{Holt2000, Yackulic2015}.
+
 This model enables us to more precisely define the niche in terms of the fundamental processes of colonization and extinction (fig..). 
-For example, the response of \emph(Acer saccharum) to temperature suggests that the southern range limit is a consequence of elevated extinction rates, whereas the northern limit is likely determined by low rates of colonization.
+For example, the response of \emph{Acer saccharum} to temperature suggests that the southern range limit is a consequence of elevated extinction rates, whereas the northern limit is likely determined by low rates of colonization.
 The consequence of these two processes, when overlaid, is a range covering intermediate temperatures, bounded by low colonization in the north and high extinction in the south (fig - map).
@@
 The rate at which this value approaches zero (i.e., the relative slopes of $c$ and $e$ along the environmental gradient -- ref response curves) indicates the degree to which the range margin will respond to changes in the gradient.
-Thus, "fast" species with steep slopes along the gradient (e.g., list some species) will respond quickly to climate change, while "slow" species may lag behind climate change, producing instability and potentially promoting catastrophic changes (e.g., widespread and rapid local extinctions along the range margin) \cite{Scheffer2009, Scheffer2012}.
+Thus, “fast” species with steep slopes along the gradient (e.g., list some species) will respond quickly to climate change, while “slow” species may lag behind climate change, producing instability and potentially promoting catastrophic changes (e.g., widespread and rapid local extinctions along the range margin) \cite{Scheffer2009, Scheffer2012}.
 %% NOTE - this may be a paper in itself - worth considering.
@@
 
-% can this apply to other organisms? animals prob not - dispersal is just too high
-% this deserves a paragraph here
 Our model is particularly appropriate for forest trees, which are sessile, disperse relatively short distances, and are long-lived.
+For these organisms, the model can easily be applied in relatively continuous landscapes, as we have done, as well as in more heterogeneous landscapes \cite{Garcia2013}.
 Unique challenges confront applying such a model without these characteristics, such as most animals.
-Animals in highly patchy environments, with local populations linked by moderate dispersal rates, are likely to be the most appropriate species for use with our model.
+However, similar models have been successful in describing dynamics of animal introductions \cite{Yackulic2015} and in describing occupancy in highly patchy environments \cite{Ovaskainen2004, Harrison2011}.
 

### 2015-07-10 scr/5_mcmc_scripts/pintae1.sh

@@ -1,6 +1,6 @@
 #!/bin/sh
 #PBS -q default
-#PBS -l walltime=24:00:00
+#PBS -l walltime=72:00:00
 #PBS -l nodes=1:ppn=40
 #PBS -r n
 #PBS -N pintae1
@@ -14,6 +14,6 @@ SRC=~/STModel-MCMC/bin
 
 SPECIES=18037-PIN-TAE
 DIR=~/STModel-Two-State/species/$SPECIES
-cd $DIR; $SRC/stm2_mcmc -r res/mcmc1/resumeData.txt -t dat/mcmc_data.txt -i 40000 2>mcmc_log1.txt &
+cd $DIR; $SRC/stm2_mcmc -r res/mcmc1/resumeData.txt -t dat/mcmc_data.txt -i 80000 2>mcmc_log1.txt &
 
 wait

### 2015-07-10 scr/5_mcmc_scripts/pintae2.sh

@@ -1,6 +1,6 @@
 #!/bin/sh
 #PBS -q default
-#PBS -l walltime=24:00:00
+#PBS -l walltime=72:00:00
 #PBS -l nodes=1:ppn=40
 #PBS -r n
 #PBS -N pintae2
@@ -14,6 +14,6 @@ SRC=~/STModel-MCMC/bin
 
 SPECIES=18037-PIN-TAE
 DIR=~/STModel-Two-State/species/$SPECIES
-cd $DIR; $SRC/stm2_mcmc -r res/mcmc2/resumeData.txt -t dat/mcmc_data.txt -i 40000 2>mcmc_log2.txt &
+cd $DIR; $SRC/stm2_mcmc -r res/mcmc2/resumeData.txt -t dat/mcmc_data.txt -i 80000 2>mcmc_log2.txt &
 
 wait

### 2015-07-10 scr/5_mcmc_scripts/pintae3.sh

@@ -1,6 +1,6 @@
 #!/bin/sh
 #PBS -q default
-#PBS -l walltime=24:00:00
+#PBS -l walltime=72:00:00
 #PBS -l nodes=1:ppn=40
 #PBS -r n
 #PBS -N pintae3
@@ -14,6 +14,6 @@ SRC=~/STModel-MCMC/bin
 
 SPECIES=18037-PIN-TAE
 DIR=~/STModel-Two-State/species/$SPECIES
-cd $DIR; $SRC/stm2_mcmc -r res/mcmc3/resumeData.txt -t dat/mcmc_data.txt -i 40000 2>mcmc_log3.txt &
+cd $DIR; $SRC/stm2_mcmc -r res/mcmc3/resumeData.txt -t dat/mcmc_data.txt -i 80000 2>mcmc_log3.txt &
 
 wait

### 2015-07-10 scr/6_post_mcmc_check.r

@@ -1,25 +1,26 @@
 library(coda)
 
 setwd("~/Dropbox/work/projects/STModel-Two-State_git/")
-spName = '19481-BET-ALL'
+spName = '18037-PIN-TAE'
 
-design = '11110001110000'
+design = '11110001111000'
 design = sapply(1:nchar(design), function(i) as.integer(substr(design, i, i)))
 constCols = which(design == 0)
 
 st1 = st2 = st3 = 1
 # special instructions for individual species due to poor starts
 if(spName == '19408-QUE-RUB')
-	st2 = 14000
+	st2 = 14001
 if(spName == '27821-NYS-SYL')
-	st2 = 17000
+	st2 = 17001
 if(spName == '19280-QUE-NIG')
-	st2 = 19000
+	st3 = 19001
 
 p1 = read.csv(file.path('species', spName, 'res', 'mcmc1', 'posterior.csv'))[,-constCols]
 p2 = read.csv(file.path('species', spName, 'res', 'mcmc2', 'posterior.csv'))[,-constCols]
 p2 = p2[st2:nrow(p2),]
 p3 = read.csv(file.path('species', spName, 'res', 'mcmc3', 'posterior.csv'))[,-constCols]
+p3 = p3[st3:nrow(p3),]
 posterior = mcmc.list(mcmc(p1), mcmc(p2), mcmc(p3))
 lapply(posterior, nrow)
 

### 2015-07-16 tex/manuscript.tex
Lots of changes to text, and revized figure 2 substantially

@@
 \subsection*{Objectives}
-\begin{itemize}
-	\item Describe the range limits at equilibrium of ecologically important trees in eastern North America as predicted by metapopulation theory
-	\item Evaluate the relationship between macroclimatic variables and colonization and extinction rates
-	\item Test quantitatively whether the equilibrium prediction matches the contemporary observed range
-	\item Compare among species the predicted resilience to environmental change 
-\end{itemize}
+%%% STRUCTURE AHEAD
+% P1: general outline of the story, including the basics of the theory so the reader can grok the main point
+% P2: first sentence is the headline - the big result. Follow with a general description of the dataset, some of the cool bits of our study, and end with the main objectives/questions
+% P3: what is the state of the art with resepct to north american forests?
+why do we care about this?
+this is completely lacking from what I have below.
+Can probably glean something from the NSERC grant to start
+% P4: some details about the model and how it adds to what has been done
+% P5 (or end of P4): A summary of how we went about our study and how the pieces addressed the questions
+%% NOTE - keep ALL of the specifics about the methods out of the intro. Put just as much as is needed into the results, and rest in methods or appendix
 
-Present and future rapidly changing global environmental conditions are likely to push many species to the limits of their ecological niches.
-Biogeographical theory and empirical work suggests that species can respond to these changes by adapting \emph{in situ} to the new conditions, geographically tracking ideal conditions (resulting in range shifts), and/or experiencing local extinction where conditions no longer support presence.
-Species distribution models (SDMs) are commonly-applied to understand the latter two options.
-Classical SDMs often use a bioclimatic modeling approach that statistically relates presence or absence of the species to climate \cite{Guisan2005}.
-Such models are quite powerful when describing the present realized niche of a species, however they suffer from several drawbacks when predicting the future distribution of species, and they provide no insight into the mechanisms underlying a species' response to the environment.
-An alternative more mechanistic approach is to examine the colonization rate $c$ of unoccupied patches and the extinction rate $e$ of occupied patches \cite{Yackulic2015}.
-Theoretical studies have suggested that range limits can arise due to spatial variation in these two rates \cite{Holt2000, Holt2005}.
-(here maybe briefly review the literature on the subject)
-In such a stochastic system, the long-term expectation is that presence will only be possible where $c > e$.
-It follows that, if these rates vary along an environmental gradient, the equilibrium position of the range limit on the gradient will be the point at which $c = e$.
+% this paragraph might be a bit too modeling focused - find and focus on the STORY
+Present and future rapidly changing global environmental conditions are likely to push many species to the limits of their ecological niches and lead to mismatches between species ranges and local environmental conditions (cite).
+A comprehensive understanding of the mechanisms producing species range boundaries is thus a crucial aspect of predicting the response to global change.
+A prediction of niche theory (cite) is that species will occur where environmental conditions permit positive population growth rates.
+Species distribution models (SDMs) are a common application of niche theory; often these models use a bioclimatic modeling approach that statistically relates presence or absence of the species to climate \cite{Guisan2005}.
+Such models are quite powerful when describing the present realized niche of a species, however they are largely phenomenological and thus may fail when projecting into the future due to nonlinear responses to environmental conditions and non-equilibrial dynamics.
+A related drawback is that such models offer little insight into the mechanisms driving the response to the environment.
+Metapopulation theory \cite{Levins1969} offers an appealing more mechanistic alternative, where subpopulations in a landscape are linked by dispersal and occupancy is governed by stochastic colonizations and extinctions.
+When the rate of colonization ($c$) and extinction ($e$) vary with environmental gradients at large scales, metapopulation theory predicts that colonization-extinction (herefter C-E) dynamics can generate range limits \cite{Holt2000, Holt2005}.
+Because the fundamental rates $c$ and $e$ vary with climate, this framework is coherent with the predictions of niche theory, but it offers a compromise between highly detailed individual-based models and phenomenological SDMs \cite{Harrison2011, Yackulic2015}.
 
-We built a dynamic state-and-transition model (STM) using classic metapopulation theory \cite{Levins1969} as the foundation.
-The classical formulation of this model is inappropriate for modeling range dynamics at large scales, as it assumes colonization and extinction rates are constant in space and time and treats prevalence as a global variable.
-Holt and Keitt extended the Levins model to a scale appropriate for range dynamics by varying the parameters along an environmental gradient, showing that, under metapopulation dynamics, range limits of a theoretical species arise when an environmental gradient causes an increase in the extinction rate, a decrease in the colonization rate, or a decrease in available habitat \cite{Holt2000}.
-Similar models have been applied to investigate whether species distributions are at equilibrium with climate \cite{Garcia2013}.
+We applied a spatially heterogenous state-and-transition model (STM) based on classic metapopulation theory \cite{Levins1969, Holt2000} to show that variation in colonization and extinction rates along climatic gradients produces range limits for several abundant North American tree species.
 
-We applied a model similar to the Holt model \cite{Holt2000} to 21 forest tree species that are abundant in eastern North American temperate and boreal forests.
-Our model, which is discrete in space and time, treats the landscape as a series of patches, where the state $Y_{x,t}$ of patch $x$ at time $t$ is a function of the colonization ($c_{x,t}$) and extinction ($e_{x,t}$) rates, as well as the prevalence of the species ($phi_{x,t}$).
-Both the colonization and extinction rates were parameterized as a function of two environmental variables, one related to temperature and one precipitation, thus $c_x = f(T_x, P_x); e_x = g(T_X, P_x)$ (see \emph{Methods} for descriptions of the specific variables used).
-As our objectives concern equilibrial dynamics and the role of CE dynamics in determining species ranges, for the present we ignore temporal dynamics and focus instead on spatial variation in the model parameters.
-
-Our model is a first order Markov chain that is dynamic in space and time (though we fix the time dimension for this paper).
-The state of a patch depends on the state of local patches at the previous time step and the transition rates.
-Range boundaries emerge at equilibrium due to the effect of environmental gradients on the transition rates
-
-We apply our model exclusively to forest trees, and because our model is parameterized using forest plot data, we treat all patches as potential habitat, thus setting $k=1$ from the Levins model.
-Following this simplification, the equilibrium condition for the presence of a species on the environmental gradient is any point $x$ such that $c_x > e_x$ \cite{Holt2000}.
-It follows that the range limit of the species will occur where $c_x = e_x$.
-
-We designed a colonization-extinction model with a dynamic parameterization that allowed the colonization and extinction rates to vary as a function of climate.
 A major challenge to employing these models is the lack of time series data needed to estimate the parameters.
 We used a large database of permanent forest plots, containing repeated observations of xxx plots in eastern North America (n=xxx paired observations). 
-We selected the 21 most abundant species (based on the number of observations in the database) as the focus of this study.
-We performed model selection, using 6 macroclimatic variables, then parameterized the model using Markov Chain Monte Carlo (MCMC) sampling and addressed the following questions for each selected species:
+Our main objectives were to address the following questions:
 \begin{enumerate}
-	\item Are range limits driven by colonization-extinction dynamics?
-	\item To what extent are the species studied presently out of equilibrium with the environment?
-	\item How resilient are range limits to perturbations in the environment?
-	\item Which climatic variables best describe colonization and extinction rates?
+	\item Do colonization and extinction rates vary as a function of climate?
+	\item Are range limits of the species studied governed by C-E dynamics?
+	\item Are presently observed species ranges at equilibrium with current climatic conditions?
 \end{enumerate}
+
+The classical Levins model is inappropriate for modeling range dynamics at large scales, as it assumes colonization and extinction rates are constant in space and time and allows global dispersal from anywhere in the metapopulation.
+We built a model similar to Holt's extension of the Levins model \cite{Holt2000}, where colonization and extinction rates vary along an environmental gradient, with the addition of local, rather than global dispersal.
+Our model, which is discrete in space and time, treats the landscape as a series of patches, where the state $Y_{x,t}$ of patch $x$ at time $t$ can take one of two states: 1 (species is present) or 0 (absence).
+Empty patches are colonized based on the spatially heterogeneous colonization rate $c_x$ and the local prevalence $\phi_{x}$ (i.e., the proportion of patches in the local neighborhood that are occupied):
+\begin{equation}
+	Pr(Y_{x,t+1} = 1 \given Y_{x,t} = 0) = c_{x}\phi_{x,t}
+\end{equation}
+Occupied patches go extinct stochastically based on the extinction rate $e_x$:
+\begin{equation}
+	Pr(Y_{x,t+1} = 0 \given Y_{x,t} = 1) = e_{x}
+\end{equation}
+In this system, the equilibrium condition for the presence of a species on the environmental gradient is any point $x$ such that $c_x > e_x$ \cite{Holt2000}.
+It follows that the range limit of the species will occur where $c_x = e_x$.
+We parameterized the colonization and extinction rates as a function of two environmental variables, one related to temperature and one precipitation, thus $c_x = f(T_x, P_x); e_x = g(T_X, P_x)$, where $T_x$ and $P_x$ are the temperature and precipitation at location $x$ (see \emph{Methods} for descriptions of the specific climate variables used).
 
@@
 \section*{Results}
-\begin{itemize}
-	\item response of C-E to env variables
-	\item validation of models w.r.t. reserved data (how good is the model)
-	\item ability of the model to predict the range (equilibrium)
-	\item rate of response to disequilibrium
-	\item ability of model to predict response to environmental change
-\end{itemize}
+
+\subsection*{Model Validation}
+We first obtained maximum likelihood estimates for all model parameters using the best-supported model, estimated via simulated annealing and fit using half of the data.
+We then used the parameters to predict the transition probabilities for the reserved half of the data.
+To determine how well the model predicted observed colonizations and extinctions, we computed two common validation statistics, xxxx (TSS) and the area under the receiver operating curve (RSS) (CITE).
+For all species, the models performed well, with ROC > 0.96 and TSS > 0.81 (Table \ref{tab:validation}).
+
+\subsection*{Response of colonization and extinction rate to climate}
+Colonization and/or extinction rates of most species (xx of 21) showed strong responses to either temperature or precipitation gradients (Fig. \ref{fig:response_curves}, Appendix xx).
+The major exceptions to this pattern were the conifers \emph{Abies balsamea}, \emph{Picea glauca}, \emph{Pinus banksiana}, and (others??).
+Extinction rates in these species were flat with respect to climate, and colonization rates showed weak responses with high posterior variance.
+
+Temperature was a significant limiting factor for many species, as demonstrated by colonization and extinction curves that crossed, indicating the range limit as determined by CE dynamics.
+xx of xx broadleaf trees (\emph{Acer rubrum}, \emph{A.
+saccharum}, \emph{Fagus grandifolia}, \emph{Populous tremuloides}, \emph{Quercus alba}, \emph{Q.
+velutina}, and \emph{Ulmus americana}) and xx of xx conifers(\emph{Pinus strobus}) (and others??) were limited by CE dynamics at both the warm and cold end of their ranges.
+An additional xx broadleaf trees (\emph{Betuyla papyrifera}, \emph{Fraxinus americana} were limited by CE dynamics at the warm but not the cold end of the range, and \emph{F.
+pennsylvanica} was limited at the cold end of the range.
+xx conifers (\emph{Abies balsamea}, \emph{Picea glauca}, \emph{Pinus banksiana}) were not limited by CE dynamics with respect to temperature.
+
+In addition to temperature dependence, xx of xx species showed range limits with respect to precipitation, consisting of \emph{A.
+saccharum} (limited by greater seasonality), \emph{Fagus grandifolia} (limited by wet and dry summers), and \emph{Quercus alba} and \emph{Ulmus americana} (limited by low seasonality).
+
+\subsection*{Model predictions of range limits}
+Here we compare the STM to the SDM in its ability to predict presence/absence at equilibrium in the temporary plots.
+Waiting for the temp plot data from Steve before I do this.
+
+\subsection*{Resilience/Rate of response to perturbation}
+
 \begin{figure}
@@
 	\caption{
-		Response of colonization (blue) and extinction (red) rates of selected species to the environment.
-		Solid lines are the median predictions of xx posterior samples, and shaded regions show 95\% Bayesian credible intervals.
+		Response of colonization (blue) and extinction (red) rates of two broadleaf trees (top two rows) and two conifers (bottom two rows) to temperature (left column) and precipitation (right column).
+		Solid lines are the median posterior predictions, and shaded regions show 95\% credible intervals.
 		Species not shown here can be found in supporting information.
+		NOTE: so far, I like Acer saccharum, Betula papyrifera, Abies balsamea, and Quercus alba for the 4 species to use here and in the second figure
 	}
@@
 
+\subsection*{Colonization-extinction dynamics as the determinant of range limits}
+
+\subsection*{disorganized stuff}
+Holt and Keitt extended the Levins model to a scale appropriate for range dynamics by varying the parameters along an environmental gradient, showing that, under metapopulation dynamics, range limits of a theoretical species arise when an environmental gradient causes an increase in the extinction rate, a decrease in the colonization rate, or a decrease in available habitat \cite{Holt2000}.
+Similar models have been applied to investigate whether species distributions are at equilibrium with climate \cite{Garcia2013}.
+
 Fisichelli and colleagues reported recruitement of temperate trees into boreal ecosystems, supporting our conclusion that colonization of temperature trees varies as a function of temperature \cite{Fisichelli2014}.
@@
 This model enables us to more precisely define the niche in terms of the fundamental processes of colonization and extinction (fig..). 
-For example, the response of \emph{Acer saccharum} to temperature suggests that the southern range limit is a consequence of elevated extinction rates, whereas the northern limit is likely determined by low rates of colonization.
+For example, the response of \emph{Acer saccharum} to temperature suggests that the northern range limit is likely determined by the dispersal and establishment of seedlings, whereas the southern range limit is a consequence of elevated extinction rates despite much higher colonization rates compared to the northern limit (Fig.
+\ref{fig:response_curves}).
 The consequence of these two processes, when overlaid, is a range covering intermediate temperatures, bounded by low colonization in the north and high extinction in the south (fig - map).

### 2015-07-16 tex/tab_si1_col_pars.tex
Lots of changes to text, and revized figure 2 substantially

@@
 
-\multirow{2}{*}{\it Acer rubrum} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx &  & x.xx \\
- &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  & {\ts (x.xx, x.xx)} \\
+\multirow{2}{*}{\it Acer rubrum} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & -1.68 & 0.13 & 0.00 &  & -0.06 \\
+ &  &  & {\ts (-1.74, -1.63)} & {\ts (0.08, 0.18)} & {\ts (-0.07, 0.06)} &  & {\ts (-0.09, -0.02)} \\
 
@@
 
-\multirow{2}{*}{\it Betula papyrifera} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx & x.xx & x.xx \\
- &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} \\
+\multirow{2}{*}{\it Betula papyrifera} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & -2.70 & -1.60 & -0.20 & -1.06 & 0.09 \\
+ &  &  & {\ts (-3.04 -2.37)} & {\ts (-2.33 -0.89)} & {\ts (-0.28 -0.11)} & {\ts (-1.42 -0.72)} & {\ts (0.03  0.15)} \\
 
-\multirow{2}{*}{\it Fagus grandifolia} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx & x.xx &  \\
- &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  \\
+\multirow{2}{*}{\it Fagus grandifolia} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & -1.93 & 0.41 & -0.05 & -0.29 &  \\
+ &  &  & {\ts (-2.07 -1.81)} & {\ts (0.30  0.52)} & {\ts (-0.17  0.08)} & {\ts (-0.46 -0.11)} &  \\
 
-\multirow{2}{*}{\it Fraxinus americana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx & x.xx &  \\
- &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  \\
+\multirow{2}{*}{\it Fraxinus americana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & -1.97 & 0.06 & -0.28 & 0.32 &  \\
+ &  &  & {\ts (-2.08 -1.87)} & {\ts (-0.08  0.21)} & {\ts (-0.40 -0.16)} & {\ts (0.14  0.51)} &  \\
 
-\multirow{2}{*}{\it Fraxinus pennsylvanica} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx & x.xx & x.xx \\
- &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} \\
+\multirow{2}{*}{\it Fraxinus pennsylvanica} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & -1.64 & 0.38 & -0.28 & -0.24 & 0.18 \\
+ &  &  & {\ts (-1.85 -1.45)} & {\ts (0.05  0.73)} & {\ts (-0.45 -0.12)} & {\ts (-0.44 -0.05)} & {\ts (0.10  0.27)} \\
 
@@
 
-\multirow{2}{*}{\it Picea glauca} & \multirow{2}{*}{MDR} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx & x.xx & x.xx \\
- &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} \\
+\multirow{2}{*}{\it Picea glauca} & \multirow{2}{*}{MDR} & \multirow{2}{*}{PS} & -1.79 & 0.08 & 0.08 & 0.04 & 0.09 \\
+ &  &  & {\ts (-1.93 -1.64)} & {\ts (-0.05  0.20)} & {\ts (-0.02  0.18)} & {\ts (-0.02  0.09)} & {\ts (0.02  0.17)} \\
 
@@
 
-\multirow{2}{*}{\it Pinus banksiana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx &  &  \\
- &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  &  \\
+\multirow{2}{*}{\it Pinus banksiana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & -2.07 & 0.81 & 0.28 &  &  \\
+ &  &  & {\ts (-2.77 -1.41)} & {\ts (0.35  1.26)} & {\ts (-0.06  0.63)} &  &  \\
 
-\multirow{2}{*}{\it Pinus strobus} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx &  &  \\
- &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  &  \\
+\multirow{2}{*}{\it Pinus strobus} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & -2.12 & 0.01 & -0.05 &  &  \\
+ &  &  & {\ts (-2.25 -1.98)} & {\ts (-0.17  0.19)} & {\ts (-0.19  0.09)} &  &  \\
 
@@
 
-\multirow{2}{*}{\it Quercus alba} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx & x.xx &  \\
- &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  \\
+\multirow{2}{*}{\it Quercus alba} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & -3.00 & 0.30 & -0.06 & 0.40 &  \\
+ &  &  & {\ts (-3.15 -2.85)} & {\ts (-0.04  0.65)} & {\ts (-0.21  0.09)} & {\ts (0.15  0.65)} &  \\
 
@@
 
-\multirow{2}{*}{\it Quercus velutina} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx & x.xx & x.xx \\
- &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} \\
+\multirow{2}{*}{\it Quercus velutina} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & -1.95 & -0.49 & -0.15 & 0.45 & 0.21 \\
+ &  &  & {\ts (-2.09 -1.81)} & {\ts (-0.84 -0.14)} & {\ts (-0.30  0.00)} & {\ts (0.11  0.80)} & {\ts (0.06  0.35)} \\
 
-\multirow{2}{*}{\it Ulmus americana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx &  & \\
- &  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & & \\
+\multirow{2}{*}{\it Ulmus americana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & -1.15 & -0.07 & 0.06 &  & \\
+ &  &  & {\ts (-1.24 -1.07)} & {\ts (-0.17  0.03)} & {\ts (0.00  0.13)} & & \\
 \bottomrule

### 2015-07-16 tex/tab_si2_ext_pars.tex
Lots of changes to text, and revized figure 2 substantially

@@
 
-\multirow{2}{*}{\it Acer rubrum} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx & x.xx & \\
-&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  \\
+\multirow{2}{*}{\it Acer rubrum} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & -3.42 & 0.34 & -0.20 & 0.44 & \\
+&  &  & {\ts (3.51, -3.33)} & {\ts (0.28, 0.41)} & {\ts (-0.26, -0.14)} & {\ts (0.36, 0.53)} &  \\
 
@@
 
-\multirow{2}{*}{\it Betula papyrifera} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx &  &  \\
-&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  &  \\
+\multirow{2}{*}{\it Betula papyrifera} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & -1.65 & 1.06 & 0.19 &  &  \\
+&  &  & {\ts (-1.80 -1.49)} & {\ts (0.89  1.22)} & {\ts (0.14  0.23)} &  &  \\
 
-\multirow{2}{*}{\it Fagus grandifolia} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx & x.xx & x.xx \\
-&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} \\
+\multirow{2}{*}{\it Fagus grandifolia} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & -3.50 & 0.00 & -0.20 & 0.64 & 0.23 \\
+&  &  & {\ts (-3.68 -3.32)} & {\ts (-0.13  0.12)} & {\ts (-0.33 -0.08)} & {\ts (0.46  0.82)} & {\ts (0.13  0.33)} \\
 
-\multirow{2}{*}{\it Fraxinus americana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx & x.xx & x.xx \\
-&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)}  \\
+\multirow{2}{*}{\it Fraxinus americana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & -2.58 & 0.47 & -0.45 & 0.55 & 0.09 \\
+&  &  & {\ts (-2.69 -2.47)} & {\ts (0.32  0.63)} & {\ts (-0.57 -0.34)} & {\ts (0.38  0.73)} & {\ts (-0.01  0.17)}  \\
 
-\multirow{2}{*}{\it Fraxinus pennsylvanica} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx & x.xx & x.xx \\
-&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} \\
+\multirow{2}{*}{\it Fraxinus pennsylvanica} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & -1.68 & -0.69 & 0.16 & 0.43 & -0.16 \\
+&  &  & {\ts (-1.86 -1.51)} & {\ts (-0.98 -0.42)} & {\ts (0.02  0.31)} & {\ts (0.26  0.62)} & {\ts (-0.24 -0.08)} \\
 
@@
 
-\multirow{2}{*}{\it Picea glauca} & \multirow{2}{*}{MDR} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx &  & x.xx \\
-&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  & {\ts (x.xx, x.xx)} \\
+\multirow{2}{*}{\it Picea glauca} & \multirow{2}{*}{MDR} & \multirow{2}{*}{PS} & -2.45 & 0.13 & 0.09 &  & 0.10 \\
+&  &  & {\ts (-2.58 -2.31)} & {\ts (0.07  0.19)} & {\ts (-0.01  0.18)} &  & {\ts (0.03  0.16)} \\
 
@@
 
-\multirow{2}{*}{\it Pinus banksiana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & x.xx & x.xx & x.xx & x.xx &  \\
-&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  \\
+\multirow{2}{*}{\it Pinus banksiana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MAP} & -3.30 & -1.32 & -0.62 & -1.17 &  \\
+&  &  & {\ts (-4.09 -2.58)} & {\ts (-2.72 -0.02)} & {\ts (-0.86 -0.37)} & {\ts (-1.87 -0.50)} &  \\
 
-\multirow{2}{*}{\it Pinus strobus} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx & x.xx & x.xx \\
-&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} \\
+\multirow{2}{*}{\it Pinus strobus} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & -3.29 & 0.16 & 0.15 & 0.81 & -0.19 \\
+&  &  & {\ts (-3.50 -3.08)} & {\ts (-0.15  0.44)} & {\ts (-0.06  0.36)} & {\ts (0.46  1.16)} & {\ts (-0.39  0.00)} \\
 
@@
 
-\multirow{2}{*}{\it Quercus alba} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx & x.xx & x.xx \\
-&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} \\
+\multirow{2}{*}{\it Quercus alba} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & -3.09 & -0.61 & 0.05 & 1.00 & 0.13 \\
+&  &  & {\ts (-3.25 -2.94)} & {\ts (-0.87 -0.32)} & {\ts (-0.10  0.21)} & {\ts (0.81  1.17)} & {\ts (0.02  0.25)} \\
 
@@
 
-\multirow{2}{*}{\it Quercus velutina} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & x.xx & x.xx & x.xx & x.xx &  \\
-&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} &  \\
+\multirow{2}{*}{\it Quercus velutina} & \multirow{2}{*}{MAT} & \multirow{2}{*}{MPWQ} & -2.16 & -1.06 & -0.08 & 1.52 &  \\
+&  &  & {\ts (-2.27 -2.05)} & {\ts (-1.32 -0.79)} & {\ts (-0.18  0.02)} & {\ts (1.28  1.76)} &  \\
 
-\multirow{2}{*}{\it Ulmus americana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & x.xx & x.xx & x.xx & x.xx & \\
-&  &  & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & {\ts (x.xx, x.xx)} & \\
+\multirow{2}{*}{\it Ulmus americana} & \multirow{2}{*}{MAT} & \multirow{2}{*}{PS} & -1.59 & -0.16 & -0.24 & 0.26 & \\
+&  &  & {\ts (-1.67 -1.51)} & {\ts (-0.27 -0.05)} & {\ts (-0.29 -0.19)} & {\ts (0.15  0.37)} & \\
 \bottomrule

### 2015-07-16 scr/6_post_mcmc_check.r

@@ -28,11 +28,25 @@ plot(posterior, ask=T)
 summary(posterior)
 gelman.diag(posterior)
 
+
+# throw out first half of the samples
+# everyone gets thinned down to 10,000 samples for the 'full' posterior and 1000 for maps
 st = floor(nrow(p1)/2) + 1
 en = nrow(p1)
-th.int = 10
-th = seq(st, en, th.int)
-posterior.burnin = mcmc.list(mcmc(p1[st:en,], start=st), mcmc(p2[st:en,], start=st), mcmc(p3[st:en,], start=st))
-posterior.thinned = mcmc.list(mcmc(p1[th,], start=st, thin=th.int), mcmc(p2[th,], start=st, thin=th.int), mcmc(p3[th,], start=st, thin=th.int))
+
+full.len = 10000
+full.th = floor((en - st)/(full.len - 1))
+full.seq = seq(st, en, by = full.th)
+
+thin.len = 1000
+thin.th = floor((en - st)/(thin.len - 1))
+thin.seq = seq(st, en, by = thin.th)
+
+posterior.burnin = list(p1[full.seq,], p2[full.seq,], p3[full.seq,])
+posterior.burnin = as.mcmc.list(lapply(posterior.burnin, mcmc, start=st, thin=full.th))
+
+posterior.thinned = list(p1[thin.seq,], p2[thin.seq,], p3[thin.seq,])
+posterior.thinned = as.mcmc.list(lapply(posterior.thinned, mcmc, start=st, thin=thin.th))
+
 saveRDS(posterior.burnin, file.path('species', spName, 'res', paste(spName, 'posterior.rds', sep='_')))
 saveRDS(posterior.thinned, file.path('species', spName, 'res', paste(spName, 'posterior_thinned.rds', sep='_')))

### 2015-07-16 scr/fig/fig1_response_curves.r

@@ -1,9 +1,17 @@
 #!/usr/bin/Rscript
 varNames = readRDS("dat/climVariableNames.rds")
-spNames = c('18032-ABI-BAL', '28728-ACE-RUB', '28731-ACE-SAC', '19489-BET-PAP',
-		'19462-FAG-GRA', '32931-FRA-AME', '32929-FRA-PEN', '183295-PIC-GLA', 
-		'183319-PIN-BAN', '183385-PIN-STR', '195773-POP-TRE', '19290-QUE-ALB', 
-		'19447-QUE-VEL', '19049-ULM-AME')
+
+# ugly ones
+spNames = c('18032-ABI-BAL', '28728-ACE-RUB', '183295-PIC-GLA', '183302-PIC-MAR', '183319-PIN-BAN')
+
+# nice ones
+spNames = c('28731-ACE-SAC', '19481-BET-ALL', '19489-BET-PAP', '195773-POP-TRE', '19290-QUE-ALB')
+
+## spNames = c('18032-ABI-BAL', '28728-ACE-RUB', '28731-ACE-SAC', '19481-BET-ALL', 
+## 		'19489-BET-PAP', '19462-FAG-GRA', '32931-FRA-AME', '32929-FRA-PEN', 
+## 		'18086-LIR-TUL', '27821-NYS-SYL', '183295-PIC-GLA', '183302-PIC-MAR',  
+## 		'183319-PIN-BAN', '183385-PIN-STR', '195773-POP-TRE', '19290-QUE-ALB', 
+## 		'19280-QUE-NIG', '19408-QUE-RUB', '19447-QUE-VEL', '19049-ULM-AME')
 spInfoAll = read.csv('dat/speciesInfo.csv', stringsAsFactors=FALSE, colClasses='character')
 
 pdf(w=6.5, h=8, file="img/response_curves.pdf")

### 2015-07-16 scr/fig/fig2_posterior_maps.r

@@ -4,10 +4,13 @@ library(sp)
 library(raster)
 library(rgdal)
 
-spNames = c('18032-ABI-BAL', '28728-ACE-RUB', '28731-ACE-SAC', '19489-BET-PAP',
-		'19462-FAG-GRA', '32931-FRA-AME', '32929-FRA-PEN', '183295-PIC-GLA', 
-		'183319-PIN-BAN', '183385-PIN-STR', '195773-POP-TRE', '19290-QUE-ALB', 
-		'19447-QUE-VEL', '19049-ULM-AME')
+spNames = c('18032-ABI-BAL', '28728-ACE-RUB', '28731-ACE-SAC', '19481-BET-ALL', 
+		'19489-BET-PAP', '19462-FAG-GRA', '32931-FRA-AME', '32929-FRA-PEN', 
+		'18086-LIR-TUL', '27821-NYS-SYL', '183295-PIC-GLA', '183302-PIC-MAR',  
+		'183385-PIN-STR', '195773-POP-TRE', '19290-QUE-ALB', 
+		'19280-QUE-NIG', '19408-QUE-RUB', '19447-QUE-VEL', '19049-ULM-AME')
+		
+spNames = c('28731-ACE-SAC', '19481-BET-ALL', '19489-BET-PAP', '195773-POP-TRE', '19290-QUE-ALB')
 spInfoAll = read.csv('dat/speciesInfo.csv', stringsAsFactors=FALSE, colClasses='character')
 
 load('dat/map_projections.rdata')
@@ -26,26 +29,7 @@ plotbg = function(txt="", rangeMap=T)
 	plot(grLakes, col="white", add=T)
 	mtext(txt, adj=0, cex = 1)
 	
-	# plot grid lines and labels
-	llgridlines(ocean, easts=seq(-90, -40, 10), norths=seq(20,50,10), ndiscr=100, 
-			side="ES", offset=1e6, col='#44444444')
-	par(xpd=TRUE)
-#	westText = c("30°N", "40°N", "50°N")
-	eastText = c("20°N", "30°N", "40°N")
-#	westCoords = list(rep(-3e5,3), c(0.75e6, 1.9e6, 3e6))
-	eastCoords = list(rep(3.5e6,3), c(0.55e6, 1.7e6, 2.9e6))
-	southText = c("90°W", "80°W", "70°W")
[…]

### 2015-07-17 tex/manuscript.tex
Lots of work on the ms

--- 2015-07-09 tex/manuscript.tex
@@
 \usepackage[flushleft]{threeparttable}
+
+\usepackage{gensymb}
 
@@
 
-	\textbf{Keywords:}
+	\textbf{Keywords:} climate change, ecological resilience, forest inventory data, Levins metapopulation model, species ranges, species distribution modeling
 	\end{flushleft}
@@
 \section*{Introduction}
-\subsection*{Objectives}
+% this paragraph might be a bit too modeling focused - find and focus on the STORY
+Present and future rapidly changing global environmental conditions are likely to push many species to the limits of their ecological niches and lead to mismatches between species ranges and local environmental conditions \cite{Walther2002}.
+When species are unable to move as quickly as the climatic niche, the increasing distance between the present distribution and the climatic optimum can drive the system toward a tipping point, where large and catastrophic changes occur rapidly \cite{Scheffer2009}.
+In such a situation, phenomenological species distribution models (SDMs) that predict species occurrence as a function of climate using present occurrence data will likely fail to correctly predict the response of the system, necessitating the use of more mechanistic models \cite{Guisan2005, Lavergne2010}.
+Metapopulation theory \cite{Levins1969} offers an appealing and more mechanistic alternative; the original Levins model described linked subpopulations in a landscape that interacted with global dispersal and stochastic local colonizations and extinctions.
+More recently, the Levins model has been extended to allow the colonization and extinction rates ($c$ and $e$) to vary along an environmental gradient \cite{Holt2000, Holt2005}.
+When applied at large scales, the prediction of the theory is that the portion of the environmental gradient where $c > e$ comprises the species range \cite{Holt2000}.
+Similar approaches have demonstrated in both theoretical and empirical systems the importance of colonization-extinction (hereafter C-E) dynamics in determining range margins \cite{Harrison2011, Yackulic2015}.
+
+We applied a spatially heterogenous state-and-transition model (STM) based on classic metapopulation theory \cite{Levins1969, Holt2000} to a very large database of repeated observations of eastern North American trees, spanning 30 years of observations and ranging from the southern extent of temperate forests to the northern extent of the boreal forest.
+The main objectives of our research were to determine whether colonization and extinction rates of abundant species in temperate and boreal forests varied with climate, whether C-E dynamics determine range limits, and the extent to which the present range of these species is out of equilibrium with the present climate.
+A major challenge to employing these models is the lack of time series data needed to estimate the parameters.
+Our study is unique in the extent of high-quality time-series data applied to a mechanistic study of range dynamics, consisting of 92,724 repeated observations of 60,661 plots ranging in latitude from 25\degree N -- 53\degree N and in latitude from 57\degree W to 97\degree W.
+We selected the 20 most commonly occurring tree species in the database and found that for xx species, colonization and extinction rates varied with climate and that C-E dynamics likely determined range limits.
+
+Forest ecosystems worldwide have already responded to recent climate warming with both altitudinal and latitudinal shifts in distribution \cite{Parmesan2006, Beckage2008, Lenoir2008, Rosenzweig2008}.
+At the scale of individual trees, climate change has likely disrupted phenological cycles and increased mortality rates for many North American trees \cite{Chuine2001, vanMantgem2009, Menzel2006}.
+For many species, the aggregate effect of these individual-level changes may be changes in patch-scale local colonizations and extinctions.
+Furthermore, the rates at which these processes occur may vary substantially at the boundaries of major bioclimatic domains.
+Potentially slow responses to changing climate at range margins, particularly at the temperate-boreal and boreal-tundra transitions, may signal the potential for catastrophic shifts to new stable equilibria \cite{Scheffer2009,Scheffer2013}, with consequently dramatic changes to biodiversity and ecosystem function \cite{Hooper2005}.
+The challenge of understanding and responding to climate change at the scale of species ranges calls for a new suite of more mechanistic models \cite{Yackulic2015}.
+
+The classical Levins model is inappropriate for modeling range dynamics at large scales, as it assumes colonization and extinction rates are constant in space and time and allows global dispersal from anywhere in the metapopulation.
+Holt's model \cite{Holt2000} extends the classic framework by varying colonization and extinction rates along an environmental gradient.
+However, in real ecosystems, dispersal limitation may impose a significant constraint on the colonization of empty patches, thus the classic assumption of local dispersal is inappropriate and would lead to biased estimates of colonization rates during parameterization.
+Our model, which is discrete in space and time, treats the landscape as a series of patches constrained, where the state $Y_{x,t}$ of patch $x$ at time $t$ can take one of two states: 1 (species is present) or 0 (absence).
+Empty patches are colonized based on the spatially heterogeneous colonization rate $c_x$ and the local prevalence $\phi_{x}$ (i.e., the proportion of patches in the local neighborhood that are occupied):
+\begin{equation}
+Pr(Y_{x,t+1} = 1 \mid Y_{x,t} = 0) = c_{x}\phi_{x,t}
+\end{equation}
+Occupied patches go extinct stochastically based on the extinction rate $e_x$:
+\begin{equation}
+Pr(Y_{x,t+1} = 0 \mid Y_{x,t} = 1) = e_{x}
+\end{equation}
+In this system, the equilibrium condition for the presence of a species on the environmental gradient is any point $x$ such that $c_x > e_x$ \cite{Holt2000}.
+It follows that the range limit of the species will occur where $c_x = e_x$.
+We parameterized the colonization and extinction rates as a function of two environmental variables, one related to temperature and one precipitation, thus $c_x = f(T_x, P_x); e_x = g(T_x, P_x)$, where $T_x$ and $P_x$ are the temperature and precipitation at location $x$ (see \emph{Methods} for descriptions of the specific climate variables used).
+We used standard model selection and evaluation methods to determine if and how $c$ and $e$ varied with climate for each species and to evaluate whether range limits for these species are determined by C-E dynamics.
+
+\section*{Results}
+
+\subsection*{Key results}
 \begin{itemize}
-	\item Describe the range limits at equilibrium of ecologically important trees in eastern North America as predicted by metapopulation theory
-	\item Evaluate the relationship between macroclimatic variables and colonization and extinction rates
-	\item Test quantitatively whether the equilibrium prediction matches the contemporary observed range
-	\item Compare among species the predicted resilience to environmental change 
+	\item All species showed some climate dependance for colonization and extinction rates, though for 2 species responses were weak and variance was high.
+	\item For all but 4 species, latitudinal range limits are likely determined in part by C-E dynamics. For the other 4, the model predicts no constraints to distribution based on C-E dynamics.
+	\item For xx species, the current distribution of the species is out of equilibrium with the environment.
+	\item xx species show a clear signal of climate warming, with the equilibrium range at the current climate projecting the loss of populations at the southern range limit and expansion at the northern range limit.
 \end{itemize}
 
-Present and future rapidly changing global environmental conditions are likely to push many species to the limits of their ecological niches.
-Biogeographical theory and empirical work suggests that species can respond to these changes by adapting \emph{in situ} to the new conditions, geographically tracking ideal conditions (resulting in range shifts), and/or experiencing local extinction where conditions no longer support presence.
-Species distribution models (SDMs) are commonly-applied to understand the latter two options.
-Classical SDMs often use a bioclimatic modeling approach that statistically relates presence or absence of the species to climate.
-Such models are quite powerful when describing the present realized niche of a species, however they suffer from several drawbacks when predicting the future distribution of species, and they provide no insight into the mechanisms underlying a species' response to the environment.
-An alternative more mechanistic approach is to examine the colonization rate $c$ of unoccupied patches and the extinction rate $e$ of occupied patches.
-Theoretical studies have suggested that range limits can arise due to spatial variation in these two rates \cite{Holt2000, Holt2005}.
-(here maybe briefly review the literature on the subject)
-In such a stochastic system, the long-term expectation is that presence will only be possible where $c > e$.
-It follows that, if these rates vary along an environmental gradient, the equilibrium position of the range limit on the gradient will be the point at which $c = e$.
+%% here expand a bit on each question above and perhaps provide a testable hypothesis.
+%We evaluated the model by reserving half of the observed sample pairs.
+%We parameterized the model and performed model selection on the calibration half, then used the parameters to predict the second state as a function of the climate and the first state in the evaluation dataset.
+%We then compared the predicted to the observed second state and computed TSS and ROC (explanation and citation?).
+%This evaluation can be interpreted as a test of the model iteself, and thus of the hypothesis that range limits are determined by colonization-extinction dynamics.
+%
+%To test to what extent species were out of equilibrium with the environment, we employed an independent dataset from the same database.
+%This dataset consisted of plots that were only sampled once, and thus were not useful in parameterizing the CE model.
+%We used the posterior predictions from the CE model to compute the probability of presence at equilibrium; we then compared this with observed presences from the dataset and computed TSS and ROC.
+%For comparison with a highly fit model, we also built an SDM using the Random Forests algorithm and computed the same statistics.
+%Species that are in equilibrium with their environment will show high coherence between the two models, and thus should have high validation scores in for both the CE and SDM.
+%Species out of equilibrium should perform poorly for the CE model and better for the SDM.
 
-We built a dynamic state-and-transition model (STM) using classic metapopulation theory \citep{Levins1969} as the foundation.
-The classical formulation of this model is inappropriate for modeling range dynamics at large scales, as it assumes colonization and extinction rates are constant in space and time and treats prevalence as a global variable.
-Holt and Keitt extended the Levins model to a scale appropriate for range dynamics by varying the parameters along an environmental gradient, showing that, under metapopulation dynamics, range limits of a theoretical species arise when an environmental gradient causes an increase in the extinction rate, a decrease in the colonization rate, or a decrease in available habitat \cite{Holt2000}.
+%\subsection*{Model Validation}
+%We first obtained maximum likelihood estimates for all model parameters using the best-supported model, estimated via simulated annealing and fit using half of the data.
+%We then used the parameters to predict the transition probabilities for the reserved half of the data.
+%To determine how well the model predicted observed colonizations and extinctions, we computed two common validation statistics, xxxx (TSS) and the area under the receiver operating curve (RSS) (CITE).
+%For all species, the models performed well, with ROC > 0.96 and TSS > 0.81 (Table \ref{tab:validation}).
 
-We applied a model similar to the Holt model \cite{Holt2000} to 21 forest tree species that are abundant in eastern North American temperate and boreal forests.
-Our model, which is discrete in space and time, treats the landscape as a series of patches, where the state $Y_{x,t}$ of patch $x$ at time $t$ is a function of the colonization ($c_{x,t}$) and extinction ($e_{x,t}$) rates, as well as the prevalence of the species ($phi_{x,t}$).
-Both the colonization and extinction rates were parameterized as a function of two environmental variables, one related to temperature and one precipitation, thus $c_x = f(T_x, P_x); e_x = g(T_X, P_x)$ (see \emph{Methods} for descriptions of the specific variables used).
-As our objectives concern equilibrial dynamics and the role of CE dynamics in determining species ranges, for the present we ignore temporal dynamics and focus instead on spatial variation in the model parameters.
+%\subsection*{Response of colonization and extinction rate to climate}
+%Colonization and/or extinction rates of most species (xx of 21) showed strong responses to either temperature or precipitation gradients (Fig. \ref{fig:response_curves}, Appendix xx).
+%The major exceptions to this pattern were the conifers \emph{Abies balsamea}, \emph{Picea glauca}, \emph{Pinus banksiana}, and (others??).
+%Extinction rates in these species were flat with respect to climate, and colonization rates showed weak responses with high posterior variance.
+%
+%Temperature was a significant limiting factor for many species, as demonstrated by colonization and extinction curves that crossed, indicating the range limit as determined by CE dynamics.
+%xx of xx broadleaf trees (\emph{Acer rubrum}, \emph{A.
+saccharum}, \emph{Fagus grandifolia}, \emph{Populous tremuloides}, \emph{Quercus alba}, \emph{Q.
+velutina}, and \emph{Ulmus americana}) and xx of xx conifers(\emph{Pinus strobus}) (and others??) were limited by CE dynamics at both the warm and cold end of their ranges.
+%An additional xx broadleaf trees (\emph{Betuyla papyrifera}, \emph{Fraxinus americana} were limited by CE dynamics at the warm but not the cold end of the range, and \emph{F.
+pennsylvanica} was limited at the cold end of the range.
+%xx conifers (\emph{Abies balsamea}, \emph{Picea glauca}, \emph{Pinus banksiana}) were not limited by CE dynamics with respect to temperature.
+%
+%In addition to temperature dependence, xx of xx species showed range limits with respect to precipitation, consisting of \emph{A.
+saccharum} (limited by greater seasonality), \emph{Fagus grandifolia} (limited by wet and dry summers), and \emph{Quercus alba} and \emph{Ulmus americana} (limited by low seasonality).
+%
+%\subsection*{Model predictions of range limits}
+%Here we compare the STM to the SDM in its ability to predict presence/absence at equilibrium in the temporary plots.
+%Waiting for the temp plot data from Steve before I do this.
 
-Our model is a first order Markov chain that is dynamic in space and time (though we fix the time dimension for this paper).
-The state of a patch depends on the state of local patches at the previous time step and the transition rates.
-Range boundaries emerge at equilibrium due to the effect of environmental gradients on the transition rates
-
-We apply our model exclusively to forest trees, and because our model is parameterized using forest plot data, we treat all patches as potential habitat, thus setting $k=1$ from the Levins model.
-Following this simplification, the equilibrium condition for the presence of a species on the environmental gradient is any point $x$ such that $c_x > e_x$ \citep{Holt2000}.
-It follows that the range limit of the species will occur where $c_x = e_x$.
-
-We designed a colonization-extinction model with a dynamic parameterization that allowed the colonization and extinction rates to vary as a function of climate.
-A major challenge to employing these models is the lack of time series data needed to estimate the parameters.
-We used a large database of permanent forest plots, containing repeated observations of xxx plots in eastern North America (n=xxx paired observations). 
-We selected the 21 most abundant species (based on the number of observations in the database) as the focus of this study.
-We performed model selection, using 6 macroclimatic variables, then parameterized the model using Markov Chain Monte Carlo (MCMC) sampling and addressed the following questions for each selected species:
-\begin{enumerate}
-	\item Are range limits driven by colonization-extinction dynamics?
-	\item To what extent are the species studied presently out of equilibrium with the environment?
-	\item How resilient are range limits to perturbations in the environment?
-	\item Which climatic variables best describe colonization and extinction rates?
-\end{enumerate}
-
-%% here expand a bit on each question above and perhaps provide a testable hypothesis.
-We evaluated the model by reserving half of the observed sample pairs.
-We parameterized the model and performed model selection on the calibration half, then used the parameters to predict the second state as a function of the climate and the first state in the evaluation dataset.
-We then compared the predicted to the observed second state and computed TSS and ROC (explanation and citation?).
-This evaluation can be interpreted as a test of the model iteself, and thus of the hypothesis that range limits are determined by colonization-extinction dynamics.
-
-To test to what extent species were out of equilibrium with the environment, we employed an independent dataset from the same database.
-This dataset consisted of plots that were only sampled once, and thus were not useful in parameterizing the CE model.
-We used the posterior predictions from the CE model to compute the probability of presence at equilibrium; we then compared this with observed presences from the dataset and computed TSS and ROC.
-For comparison with a highly fit model, we also built an SDM using the Random Forests algorithm and computed the same statistics.
-Species that are in equilibrium with their environment will show high coherence between the two models, and thus should have high validation scores in for both the CE and SDM.
-Species out of equilibrium should perform poorly for the CE model and better for the SDM.
-
-\section*{Results}
-\begin{itemize}
-	\item response of C-E to env variables
-	\item validation of models w.r.t. reserved data (how good is the model)
-	\item ability of the model to predict the range (equilibrium)
-	\item rate of response to disequilibrium
-	\item ability of model to predict response to environmental change
-\end{itemize}
 \begin{figure}
-	\includegraphics[width=5in]{../img/response_curves.pdf}
+\includegraphics[width=5in]{../img/response_curves.png}
 \caption{
-		Response of colonization (blue) and extinction (red) rates of selected species to the environment.
-		Solid lines are the median predictions of xx posterior samples, and shaded regions show 95\% Bayesian credible intervals.
-		Species not shown here can be found in supporting information.
+	Response of colonization (blue) and extinction (red) rates to temperature (left column) and precipitation (right column) for selected species (see SI for the remaining 16 species).
+	All species except \emph{Picea glauca} show clear responses to climate with low uncertainty and clear optima where $c > e$, the condition predicted for presence under the model.
+	The flat response to climate, high uncertainty, and lack of a region where $c < e$ for \emph{P. glauca} suggests that other processes likely limit the distribution of this species.
+	Solid lines are the median predictions, and shaded regions show 95\% credible intervals from $n \ge 10000$ MCMC samples.
 }
@@
 \begin{figure}
-	\includegraphics[width=6.5in]{../img/posterior_maps.pdf}
+\includegraphics[width=6.5in]{../img/posterior_maps.png}
 \caption{
-		Maps showing the predicted colonization rate (blue, left column), extinction rate(red, second column), posterior probability of presence from the CE model (third column), and predicted probability of presence from a conventional SDM (right column).
-		Posterior probability of presence is computed as the proportion of posterior samples where colonization exceeds extinction for a given location.
-		Species not shown here can be found in supporting information.
+	Predicted ranges and regions of disequilibrium for selected species (see SI for maps of the remaining 16 species).
+	Presence from the STM (left column) is computed as the probability that the colonization rate exceeds the extinction rate, while SDM predictions (middle column) show the probability of presence determined by relating occurrences to climate using a Random Forest model.
+	Range disequilibrium (right column) shows the areas where both models predict presence (blue), where the species is currently present according to the SDM but predicted to be absent at equilibrium under the STM (red), and where the species is currently absent in the SDM but expected to be present at equilibrium (green).
+	Three species (\emph{B.
+alleghaniensis}, \emph{P.
+tremuloides}, and \emph{Q.
+rubra}) have clear range limits under the STM, suggesting that C-E dynamics contribute to range boundaries, while \emph{P.
+glauca} shows no such constraints.
+	All three of the species with C-E range limits are at disequilibrium with current climate, including areas of absence despite suitable climate in the north and presence despite unsuitable climate in the south.
 }
-	\label{fig:response_curves}
+\label{fig:maps}
 \end{figure}
@@
 
-Fisichelli and colleagues reported recruitement of temperate trees into boreal ecosystems, supporting our conclusion that colonization of temperature trees varies as a function of temperature \cite{Fisichelli2014}.
+\subsection*{Colonization-extinction dynamics as the determinant of range limits}
 
-This model enables us to more precisely define the niche in terms of the fundamental processes of colonization and extinction (fig..). 
-For example, the response of \emph(Acer saccharum) to temperature suggests that the southern range limit is a consequence of elevated extinction rates, whereas the northern limit is likely determined by low rates of colonization.
-The consequence of these two processes, when overlaid, is a range covering intermediate temperatures, bounded by low colonization in the north and high extinction in the south (fig - map).
-A further advantage of the model, particularly when compared to static correlative SDMs, is the ability to investigate the resilience of the system at the range margin.
-At the range boundary, the dominant eigenvalue, representing the response of the system to small perturbations (i.e., recovery after disturbance or the acquisition of the new equilibrium following climate change) approaches zero \cite{Scheffer2009}.
-The rate at which this value approaches zero (i.e., the relative slopes of $c$ and $e$ along the environmental gradient -- ref response curves) indicates the degree to which the range margin will respond to changes in the gradient.
-Thus, "fast" species with steep slopes along the gradient (e.g., list some species) will respond quickly to climate change, while "slow" species may lag behind climate change, producing instability and potentially promoting catastrophic changes (e.g., widespread and rapid local extinctions along the range margin) \cite{Scheffer2009, Scheffer2012}.
+\subsection*{The equilibrium assumption in species distribution modeling}
+
+\subsection*{Resilience and the response to climate change}
+
+\subsection*{Applications of STMs to non-forest systems}
+
+%\subsection*{disorganized stuff}
+%Holt and Keitt extended the Levins model to a scale appropriate for range dynamics by varying the parameters along an environmental gradient, showing that, under metapopulation dynamics, range limits of a theoretical species arise when an environmental gradient causes an increase in the extinction rate, a decrease in the colonization rate, or a decrease in available habitat \cite{Holt2000}.
+%Similar models have been applied to investigate whether species distributions are at equilibrium with climate \cite{Garcia2013}.
+%
+%
+%Fisichelli and colleagues reported recruitement of temperate trees into boreal ecosystems, supporting our conclusion that colonization of temperature trees varies as a function of temperature \cite{Fisichelli2014}.
+%
+%Non-equilibrium dynamics are likely to become increasingly important in understanding the response of many species to global change \cite{Araujo2005, Svenning2013}, and many systems will likely experience significant time lags \cite{Garcia2013, Wu2015}.
+%Thus, it is crucial to develop models capable of relaxing the equilibrium assumption. 
+%Models of colonization and extinction allow species distribution to emerge from the underlying processes, thereby relaxing the equilibrium assumption \cite{Holt2000, Yackulic2015}.
+%
+%This model enables us to more precisely define the niche in terms of the fundamental processes of colonization and extinction (fig..). 
+%For example, the response of \emph{Acer saccharum} to temperature suggests that the northern range limit is likely determined by the dispersal and establishment of seedlings, whereas the southern range limit is a consequence of elevated extinction rates despite much higher colonization rates compared to the northern limit (Fig.
+\ref{fig:response_curves}).
+%The consequence of these two processes, when overlaid, is a range covering intermediate temperatures, bounded by low colonization in the north and high extinction in the south (fig - map).
+%A further advantage of the model, particularly when compared to static correlative SDMs, is the ability to investigate the resilience of the system at the range margin.
+%At the range boundary, the dominant eigenvalue, representing the response of the system to small perturbations (i.e., recovery after disturbance or the acquisition of the new equilibrium following climate change) approaches zero \cite{Scheffer2009}.
+%The rate at which this value approaches zero (i.e., the relative slopes of $c$ and $e$ along the environmental gradient -- ref response curves) indicates the degree to which the range margin will respond to changes in the gradient.
+%Thus, “fast” species with steep slopes along the gradient (e.g., list some species) will respond quickly to climate change, while “slow” species may lag behind climate change, producing instability and potentially promoting catastrophic changes (e.g., widespread and rapid local extinctions along the range margin) \cite{Scheffer2009, Scheffer2012}.
 %% NOTE - this may be a paper in itself - worth considering.
@@
 could easily measure both the variance and the autocorrelation in such sims
-
-% can this apply to other organisms? animals prob not - dispersal is just too high
-% this deserves a paragraph here
-Our model is particularly appropriate for forest trees, which are sessile, disperse relatively short distances, and are long-lived. 
-Unique challenges confront applying such a model without these characteristics, such as most animals. 
-Animals in highly patchy environments, with local populations linked by moderate dispersal rates, are likely to be the most appropriate species for use with our model.
+%
+%Our model is particularly appropriate for forest trees, which are sessile, disperse relatively short distances, and are long-lived.
+%For these organisms, the model can easily be applied in relatively continuous landscapes, as we have done, as well as in more heterogeneous landscapes \cite{Garcia2013}.
+%Unique challenges confront applying such a model without these characteristics, such as most animals.
+%However, similar models have been successful in describing dynamics of animal introductions \cite{Yackulic2015} and in describing occupancy in highly patchy environments \cite{Ovaskainen2004, Harrison2011}.
 
 \subsection*{Conclusions}
-\begin{itemize}
-	\item Range limits of [some/many/most/all] [temperate/boreal] tree species in northeastern north America are determined by colonization-extinction dynamics
-	\item Moisture and energy (i.e., mean annual temperature and precipitation) are associated with colonization and extinction rates for many species
-	\item For some species, the equilibrium state of the model did not match the predicted distribution of the species, suggesting that these species are either out of equilibrium with the environment, or that other processes (e.g., large-scale disturbances) contribute to their range dynamics.
-	\item Resilience (represented by the dominant eigenvalue (c - e) near the transition point) varies widely among species, suggesting greatly differential rates of response to environmental change.
-	\item Validation? Model Selection?
-\end{itemize}
+%\begin{itemize}
+%	\item Range limits of [some/many/most/all] [temperate/boreal] tree species in northeastern north America are determined by colonization-extinction dynamics
+%	\item Moisture and energy (i.e., mean annual temperature and precipitation) are associated with colonization and extinction rates for many species
+%	\item For some species, the equilibrium state of the model did not match the predicted distribution of the species, suggesting that these species are either out of equilibrium with the environment, or that other processes (e.g., large-scale disturbances) contribute to their range dynamics.
+%	\item Resilience (represented by the dominant eigenvalue (c - e) near the transition point) varies widely among species, suggesting greatly differential rates of response to environmental change.
+%	\item Validation? Model Selection?
+%\end{itemize}
 
-\subsection*{List of figures to make}
-\begin{itemize}
-	\item Fig 3: maps for selected species (with remaining in appendix) - 4 panels - c; e; predicted range; SDM
-	\item Fig 3: eigenvalue maps
-\end{itemize}
-
-\subsection*{Tables}
-\input{table}
-
+%\subsection*{Tables}
+%\input{table}
+%
+%
 \section*{Methods}
@@
 \subsection*{Prevalence model}
-We used the RF package in R to build models of regional prevalence for each species. 
-For predictors, we initially evaluated a set of xx environmental variables. 
-From these, we selected xx variables that were relatively uncorrelated to each other (pairwise Pearson correlation < xx) and that had high loadings in a principal components analysis (PCA) of all variables.
-In this way, we reduced the risk of estimation problems due to correlations among the predictors while retaining variables that were representative of the available climatic space.
+%We used the RF package in R to build models of regional prevalence for each species. 
+%For predictors, we initially evaluated a set of xx environmental variables. 
+%From these, we selected xx variables that were relatively uncorrelated to each other (pairwise Pearson correlation < xx) and that had high loadings in a principal components analysis (PCA) of all variables.
+%In this way, we reduced the risk of estimation problems due to correlations among the predictors while retaining variables that were representative of the available climatic space.
 
 \subsection*{Model selection}
-For each species, we evaluated a set of 145 colonization-extinction models. 
-Each model consisted of at most two environmental variables, one associated with temperature and the other precipitation\. 
-The model set was constructed from eight possible model forms (Table?) applied to all possible two-variable combinations of the four energy and four moisture variables (i.e., 16 possible variable combinations), along with two additional single-variable forms for each variable and an intercept-only model.
-Because of the large number of models, fitting all models using MCMC was infeasible; therefore, we performed model selection by computing the Akaike Information Criterion (AIC) on maximum likelihood estimates for each model, fit using simulated annealing (package GenSA in R).
-We then selected the model that minimized AIC for final parameterization.
+%For each species, we evaluated a set of 145 colonization-extinction models. 
+%Each model consisted of at most two environmental variables, one associated with temperature and the other precipitation\. 
+%The model set was constructed from eight possible model forms (Table?) applied to all possible two-variable combinations of the four energy and four moisture variables (i.e., 16 possible variable combinations), along with two additional single-variable forms for each variable and an intercept-only model.
+%Because of the large number of models, fitting all models using MCMC was infeasible; therefore, we performed model selection by computing the Akaike Information Criterion (AIC) on maximum likelihood estimates for each model, fit using simulated annealing (package GenSA in R).
+%We then selected the model that minimized AIC for final parameterization.
 
 \subsection*{Parameterization}
-We used a Markov Chain Monte Carlo (MCMC) sampler employing a Metropolis-Hastings scheme to estimate posterior distributions for all parameters.
-This method has a number of advantages over the simulated annealing approach used for preliminary model selection.
-First, MCMC results in full joint posterior distributions of the parameters, allowing for a robust understanding of the uncertainty both in model parameters and predictions.
-Second, the use of a Bayesian framework allows the inclusion of prior information.
-As our parameters were all estimated on the logit scale, we used the weakly informative Cauchy priors suggested by Gelman and colleagues \cite{Gelman2008}, with a center of 0 and scale of 10 for intercept parameters or 2.5 for slope parameters.
-Finally, having posterior distributions of the predictors allows for the information produced by our model to be incorporated in future meta-modeling efforts that can further improve our characterization of species' ranges and our understanding of model uncertainty (cite GEB paper).
+%We used a Markov Chain Monte Carlo (MCMC) sampler employing a Metropolis-Hastings scheme to estimate posterior distributions for all parameters.
+%This method has a number of advantages over the simulated annealing approach used for preliminary model selection.
+%First, MCMC results in full joint posterior distributions of the parameters, allowing for a robust understanding of the uncertainty both in model parameters and predictions.
+%Second, the use of a Bayesian framework allows the inclusion of prior information.
+%As our parameters were all estimated on the logit scale, we used the weakly informative Cauchy priors suggested by Gelman and colleagues \cite{Gelman2008}, with a center of 0 and scale of 10 for intercept parameters or 2.5 for slope parameters.
+%Finally, having posterior distributions of the predictors allows for the information produced by our model to be incorporated in future meta-modeling efforts that can further improve our characterization of species' ranges and our understanding of model uncertainty (cite GEB paper).
+%
+%For each species, we obtained xxx posterior samples...
+%
+%We assessed convergence using the multivariate potential scale reduction factor (MPSR) \cite{Brooks1998}.
+%This method compares the variance within and among multiple parallel MCMC chains, where a value of 1 indicates perfect convergence (i.e., the chains are from indistinguishable distributions), and values approximately \textgreater1.1 indicate a lack of convergence.
+%
+%For climate variables, we used regional climate models at xx resolution \cite{McKenney2011}.
 
-For each species, we obtained xxx posterior samples...
-
-We assessed convergence using the multivariate potential scale reduction factor (MPSR) \cite{Brooks1998}.
-This method compares the variance within and among multiple parallel MCMC chains, where a value of 1 indicates perfect convergence (i.e., the chains are from indistinguishable distributions), and values approximately \textgreater1.1 indicate a lack of convergence.
-
-For climate variables, we used regional climate models at xx resolution \cite{McKenney2011}.
-
-\renewcommand\refname{Literature Cited}
 \bibliography{2state}{}

### 2015-07-17 tex/si_tab_model_selection.tex
Lots of work on the ms

\begin{table}[tb]
\begin{threeparttable}
\label{tab:model_selection}
\caption{Model selection statistics for the best model selected for each species}
\begin{tabular}{llllll}
\toprule
Scientific name    & Number of parameters & $\Delta$AIC\tnote{1} & AIC$_w$\tnote{2} & TSS\tnote{2} \\ 
\midrule
{\it Abies balsamea}          &   & 2.15      & 0.74   & 0.8545  \\
{\it Acer rubrum}             &   & 0.36      & 0.27   & 0.8854  \\
{\it Acer saccharum}          &   & 0.82      & 0.36   & 0.9377  \\
{\it Betula alleghaniensis}   &   & 1.45      & 0.35   & 0.9208  \\
{\it Betula papyrifera}       &   & 0.90      & 0.32   & 0.8856  \\
{\it Fagus grandifolia}       &   & 0.52      & 0.41   & 0.9135  \\
{\it Fraxinus americana}      &   & 0.28      & 0.23   & 0.8741  \\
{\it Fraxinus pennsylvanica}  &   & 1.80      & 0.40   & 0.8379  \\
{\it Liriodendron tulipifera} &   & 0.75      & 0.30   & 0.9162  \\
{\it Nyssa sylvatica}         &   & 0.35      & 0.18   & 0.8622  \\
{\it Picea glauca}            &   & 0.64      & 0.12   & 0.8405  \\
{\it Picea mariana}           &   & 1.39      & 0.34   & 0.9148  \\
{\it Pinus strobus}           &   & 0.032     & 0.14   & 0.8978  \\
{\it Pinus taeda}             &   & 0.99      & 0.30   & 0.9278  \\
{\it Populus tremuloides}     &   & 0.65      & 0.12   & 0.8830  \\ 
{\it Quercus alba}            &   & 0.15      & 0.27   & 0.9454  \\
{\it Quercus nigra}           &   & 0.34      & 0.092  & 0.8927  \\
{\it Quercus rubra}           &   & 1.81      & 0.50   & 0.8980  \\
{\it Quercus velutina}        &   & 1.11      & 0.41   & 0.8940  \\
{\it Ulmus americana}         &   & 0.79      & 0.13   & 0.8168  \\ 
\bottomrule
\end{tabular}
\begin{tablenotes}
\item [1] $\Delta$AIC refers to absolute AIC difference between the selected model and the next best model in the set.
\item [2] AIC$_w$ can be interpreted as the probability that this model is the best model in the set (cite Burnham and Anderson).
\item [3] TSS is a measure of the ability of the model to predict observed colonizations and extinctions that were reserved (i.e., not used in parameter estimation). A value of 1 indicates perfect prediction.
\end{tablenotes}
\end{threeparttable}
\end{table}

### 2015-07-17 scr/fig/fig1_response_curves.r

@@ -1,20 +1,23 @@
 #!/usr/bin/Rscript
 varNames = readRDS("dat/climVariableNames.rds")
 
-# ugly ones
-spNames = c('18032-ABI-BAL', '28728-ACE-RUB', '183295-PIC-GLA', '183302-PIC-MAR', '183319-PIN-BAN')
-
-# nice ones
-spNames = c('28731-ACE-SAC', '19481-BET-ALL', '19489-BET-PAP', '195773-POP-TRE', '19290-QUE-ALB')
+## species chosen as the 4 example species
+spNames = c('19481-BET-ALL', '183295-PIC-GLA', '195773-POP-TRE', '19408-QUE-RUB')
 
 ## spNames = c('18032-ABI-BAL', '28728-ACE-RUB', '28731-ACE-SAC', '19481-BET-ALL', 
 ## 		'19489-BET-PAP', '19462-FAG-GRA', '32931-FRA-AME', '32929-FRA-PEN', 
 ## 		'18086-LIR-TUL', '27821-NYS-SYL', '183295-PIC-GLA', '183302-PIC-MAR',  
-## 		'183319-PIN-BAN', '183385-PIN-STR', '195773-POP-TRE', '19290-QUE-ALB', 
+## 		'183385-PIN-STR', '195773-POP-TRE', '19290-QUE-ALB', 
 ## 		'19280-QUE-NIG', '19408-QUE-RUB', '19447-QUE-VEL', '19049-ULM-AME')
 spInfoAll = read.csv('dat/speciesInfo.csv', stringsAsFactors=FALSE, colClasses='character')
 
-pdf(w=6.5, h=8, file="img/response_curves.pdf")
+paperwidth = 6.5
+dpi = 600
+hToWRatio = 1.5
+width = as.integer(dpi*paperwidth)
+height = as.integer(width * hToWRatio)
+fontsize = 15
+png(w=width, h=height, file="img/response_curves.png", pointsize=fontsize, res = dpi)
 par(mfrow=c(4,2), bty='n', mar=c(3,3,0.5,0.5), oma=c(0,0,0,2), mgp=c(2,0.5,0), tck=-0.03)
 for(spName in spNames)
 {

### 2015-07-17 scr/fig/fig2_posterior_maps.r

@@ -4,13 +4,16 @@ library(sp)
 library(raster)
 library(rgdal)
 
-spNames = c('18032-ABI-BAL', '28728-ACE-RUB', '28731-ACE-SAC', '19481-BET-ALL', 
-		'19489-BET-PAP', '19462-FAG-GRA', '32931-FRA-AME', '32929-FRA-PEN', 
-		'18086-LIR-TUL', '27821-NYS-SYL', '183295-PIC-GLA', '183302-PIC-MAR',  
-		'183385-PIN-STR', '195773-POP-TRE', '19290-QUE-ALB', 
-		'19280-QUE-NIG', '19408-QUE-RUB', '19447-QUE-VEL', '19049-ULM-AME')
+## species chosen as the 4 example species
+spNames = c('19481-BET-ALL', '183295-PIC-GLA', '195773-POP-TRE', '19408-QUE-RUB')
+
+## spNames = c('18032-ABI-BAL', '28728-ACE-RUB', '28731-ACE-SAC', '19481-BET-ALL', 
+## 		'19489-BET-PAP', '19462-FAG-GRA', '32931-FRA-AME', '32929-FRA-PEN', 
+## 		'18086-LIR-TUL', '27821-NYS-SYL', '183295-PIC-GLA', '183302-PIC-MAR',  
+## 		'183385-PIN-STR', '195773-POP-TRE', '19290-QUE-ALB', 
+## 		'19280-QUE-NIG', '19408-QUE-RUB', '19447-QUE-VEL', '19049-ULM-AME')
 		
-spNames = c('28731-ACE-SAC', '19481-BET-ALL', '19489-BET-PAP', '195773-POP-TRE', '19290-QUE-ALB')
+
 spInfoAll = read.csv('dat/speciesInfo.csv', stringsAsFactors=FALSE, colClasses='character')
 
 load('dat/map_projections.rdata')
@@ -42,13 +45,22 @@ sdm.colors = pres.colors
 exp.colors = colorRampPalette(c('#ffffff', '#eff3ff', '#6baed6', '#08519c'))
 con.colors = colorRampPalette(c('#ffffff', '#fee5d9', '#fb6a4a', '#a50f15'))
 stay.colors = colorRampPalette(c('#ffffff', '#efedf5', '#bcbddc', '#756bb1'))
-cat.colors = c('#1f78b4', '#b2df8a', '#fb9a99')
+cat.3colors = c('#1f78b4', '#b2df8a', '#fb9a99')
[…]
